### 2024-07-15 manuscript/manuscript.qmd
first draft

---
title: "Statistical Inconsistencies in Experimental Linguistics"
shorttitle: "Statistical Inconsistencies"
author:
  - name: Timo B. Roettger
    corresponding: true
    orcid: 0000-0000-0000-0001
    email: timo.roettger@gmail.com
    affiliations:
      - name: University of Oslo
        department: Department of Linguistics & Scandinavian Studies
        city: Oslo
        country: Norway
  - name: Dara Leonard Jenssen Etemady
    corresponding: false
    email: daraetemady@gmail.com
    affiliations:
      - name: University of Oslo
        department: Department of Linguistics & Scandinavian Studies
        city: Oslo
        country: Norway
author-note:
  status-changes: 
    affiliation-change: null
    deceased: null
  disclosures:
    study-registration: null
    data-sharing: null
    related-report: null
    conflict-of-interest: null
    financial-support: null
    gratitude: null
    authorship-agreements: null
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum."
keywords: [statistics, statcheck, reproducibility]
bibliography: bibliography.bib
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: doc
    keep-tex: true

execute:
  echo: false
editor: 
  markdown: 
    wrap: sentence
---

# 1. Introduction

What we know about human language and its cognitive underpinnings is often informed by experimental data, data based on which researchers draw theoretically relevant inferences using common statistical frameworks.
This inference should be transparent in order for other researchers to critically evaluate the inference process and potentially detect and correct human errors.
In the recent years, the quantitative sciences have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share], sharing of statistical protocols is still rare across the language sciences [@bochynska2023reproducible].
If statistical procedures cannot be critically evaluated, human errors might be left undetected and thus remain uncorrected in the publication record.
If these errors affect the decision procedure of the analysis, i.e. whether a hypothesis is rejected or accepted, these errors might lead to -at best- overconfident, -at worst- false theoretical conclusions.
The present paper will present evidence that the published literature in experimental linguistics contains a concerning amount of statistical errors, a state of affairs which warrants more rigorous data sharing practices.

# 2. Statistal reporting inconsistencies
The null-hypothesis significance testing (NHST) framework is, to date, the most dominant statistical framework that researchers use to test hypotheses in the language sciences [@sondereggera2024advancements].
These statistical tests are reported in specific formats which usually contain a test statistic, the degrees of freedom of that test (if applicable), and the p-value, representing the probability of observing the data (or more extreme data) given the null hypothesis (i.e. given that the test statistic is zero) (see example 1):
\
(1) F(1, 66) = 3.88, p \< .05
\
Since data and statistical protocol sharing remains rare across the language sciences [@bochynska2023reproducible], interested readers are left with trusting the authors that the statistical analysis is run and reported correctly.
Whether that trust is warranted can be assessed, however.
The three sets of indices in (1) have a clearly defined mathematical relationship and can thus be checked for consistency.
An F test with the specified degrees of freedom and a test-statistic of 3.88 should result in a p-value of 0.053 which is larger, not smaller, than 0.05.
Thus, something is wrong with the statistical report in (1).
Possible reasons for this inconsistency are manifold: It could be a typo of the comparison sign, i.e. the authors meant to use `=` or `>` rather than `<`. 
Alternatively, any of the numbers could be a typo.
Sometimes an error might indicate erroneous rounding (e.g. 0.057 being rounded down to 0.05).
Other times, human error along the data analytical pipeline might have caused an error down the line.
Without access to data and scripts, it remains unclear to the reader, what has caused the inconsistency. 
Such inconsistencies can be particularly concerning if the calculated p-value and the reported p-values are not on the same side of the alpha threshold.
In NHST, p-values below a conventionalized alpha threshold, most commonly 0.05, are interpreted as evidence that the data are sufficiently inconsistent with the null hypothesis (significant).
p-values above that threshold are considered consistent with the null hypothesis and practically lead to rejecting the alternative hypothesis (not significant).
In (1) above, the reported p-value suggest a significant result, the p-value derived from the degrees of freedom and the test statistic suggest a non-significant result.
In the following we refer to these inconsistencies as "decision inconsistency".

This form of inconsistency assessment can be automatically assessed if statistical tests are reported in an unambiguous format.
Recently, a series of studies used such automatic assessments to evaluate the prevalence of inconsistent statistical reporting across different disciplines [e.g. @nuijten2016prevalence; @nuijten2020statcheck; @green2018statcheck; @gross2021fidelity; @buckley2023estimating].
For example, looking at over 250000 p-values published in major psychology journals, @nuijten2016prevalence found that around 50% of the articles with statistical results contained at least one inconsistencies and around 12.5% contained at least one “decision inconsistency”.

To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses over 39000 p-values reported in eight experimental linguistic journals published between 2000 and 2023.
We further assess whether the rate of inconsistency differs across journals, whether that rate has changed over the course of the last 20 years and whether there is evidence for bias in these statistical-reporting inconsistencies. 
We discuss the results and offer concrete recommendations for authors, reviewers, and editors to tackle this problem.

```{r source-script}
#| echo: false
#| output: false
#| warning: false
#| message: false
#| include: false

suppressWarnings(suppressMessages(suppressPackageStartupMessages({
source("../scripts/01_Analysis.R")
})))

knitr::opts_chunk$set(dpi = 300)

## to do:
# add options to share original articles

```

# 3. Method
All quantitative analyses were conducted using R [] and the r packages [LIST ALL PACKAGES from 01_Analysis.R and add to references]

## 3.1. Statcheck
We used the R package statcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
Statcheck works as follows: After converting pdf or html articles to plain text, statcheck searches for specific strings that correspond to a NHST result, using “regular expressions”.
That way, statcheck can detect results of t tests, F tests, Z tests, χ2 tests, correlation tests, and Q tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e. it usually misses results reported in tables; and (c) the test result is reported in American Psychological Association style [@APA2020].
Given these constraints, statcheck is estimated to detect roughly 60% of all reported NHST results [@nuijten2016prevalence] Statcheck uses the reported test statistic and degrees of freedom to recalculate the p-value, compares the reported and recalculated p-values and, if there is a mismatch, reports a comparison as an “inconsistency.” The algorithm takes into account that tests might have been performed as one-tailed by identifying the search strings “one-tailed,” “one-sided,” or “directional”.
Moreover, statcheck counts p = .000 and p \< .000 as inconsistent as p-values of exactly zero are mathematically impossible and the APA manual [@APA2020] advises to report very small p-values as p \< .001.
Validity checks suggest that inter rater reliability between manual coding and statcheck is high, i.e. 0.76 for inconsistencies and 0.89 for decision inconsistencies [@nuijten2016prevalence].
The overall accuracy of statcheck is estimated between 96.2% to 99.9% [@nuijten2017validity].
We thus consider statcheck a valid, but rough proxy of statistical reporting inconsistencies.

## 3.2. Sample
Focusing on experimental linguistic research, we started with @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the term “experiment* in title, abstract or keywords. 
Out of these 100 journals, we selected all journals with at least 10% of articles containing the search string "experiment\*".
Out of the remaining 37 journals, we selected those journals that urged APA formatting either in the main body of the text or specifically regarding statistics in the author guidelines, resulting in nine remaining journals.
Moreover, to download the pdfs, the articles had to be either accessible to us through our library license, or open access, resulting in a final list of eight journals: Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA).

We included only original research articles within the publication years of 2000-2023, excluding any book reviews, response articles, commentaries, editorials, corrigenda, errata, advertisements, etc.
Articles from LAB spanned 2011-2023, while the rest spanned 2000-2023.
This procedure resulted in `{r} length(files)` research articles.
All `{r} length(files)` articles were submitted to analysis, using Statcheck v1.5.0.
157 articles were not analyzable by statcheck, and thus had to be removed from the pool, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt.

## 3.3. Data availability statement

All derived data and corresponding R scripts are publicly available here: LINK.
The original journal articles cannot be bulk-shared due to distribution restrictions by the publishers.
NEED TO FIGURE OUT A WAY

# 4. Results

## 4.1 Prevalence of inconsistencies
The results are summarized in @tbl-summary.
Out of `{r} journal_summary_counts[9,2][[1]]` articles, `{r} journal_summary_counts[9,3][[1]]` articles contained statistical tests that statcheck could assess (`{r} 100 * round(journal_summary_counts[9,3][[1]] / journal_summary_counts[9,2][[1]],2)`%), amounting to `{r} journal_summary_counts[9,4][[1]]` assessable p-values.
`{r} journal_summary_counts[9,5][[1]]` p-values were flagged as inconsistent (`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`%) and `{r} journal_summary_counts[9,6][[1]]` of which were considered decisions inconsistencies (`{r} 100* round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`%) (see @tbl-summary).

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary. Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA)."

knitr::kable(journal_summary_counts)

```

The proportion of inconsistencies ranged from `{r} round(min(journal_summary_prop[,2]),0)` to `{r} round(max(journal_summary_prop[,2]),0)`% across journals (`{r} round(min(journal_summary_prop[,3]),0)` to `{r} round(max(journal_summary_prop[,3]),0)`% for decisions inconsistencies) (see @fig-stacked-bar). These rates appear to be stable across year of publication (see @fig-year-prop).
On average, `{r} 100*round(gross_per_article[[1]],2)`% of assessable articles contained one or more inconsistencies (journals range from `{r} 100*round(min(per_article[,2]),2)` to `{r} 100*round(max(per_article[,2]),2)`%) and `{r} 100*round(gross_per_article[[2]],2)`% contained one or more decisions inconsistencies (journals range from `{r} 100*round(min(per_article[,3]),2)` to `{r} 100*round(max(per_article[,3]),2)`%).

```{r}
#| label: fig-stacked-bar
#| out-width: 100%
#| fig-pos: 'H'
#| fig-cap: "Proportion of articles containing at least one inconsistency / decision inconsistency. Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA)."

knitr::include_graphics("../plots/figure1.png")

```

```{r}
#| label: fig-year-prop
#| out-width: 100%
#| fig-cap: "Proportion of inconsistencies / decision inconsistencies from 2000 to 2023. Left: global rates with all journals pooled; Right: rates per journal. Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA)."

knitr::include_graphics("../plots/figure2.png")

```

When examining reported against recalculated p-values (see @fig-scatter) for all inconsistencies, we can identify certain spatial patterns (only inconsistencies that are based on '=' comparisons were included to make them easier to interpret, n = 
`{r} nrow(xdata[xdata$error == TRUE & xdata$test_comp == "=",])`
out of 
`{r} nrow(xdata[xdata$error == TRUE,])`).
First, the center of density of points is located in the bottom left corner with more values reported closer to the alpha level.
This is not surprising, since there is a documented publication bias in published quantitative articles, with hypotheses being much more often confirmed (and thus p < .05) than not [@franco2014publication; @sterling1959publication].
Second, there is a linear density of points along the diagonal axis which corresponds to numerically small inconsistencies, some of which might be related to simple rounding errors.
However, comparing the diagonal to the black line, which represents a linear model predicting recalculated by reported p-values, we can see a clear divergence of what is expected if inconsistencies were equally likely in both directions (i.e. recalculated p-values were as likely to be smaller than the reported p-value as larger).
The slope of the regression line is flatter than the diagonal axis which means that, on average, reported p-values are lower than their recalculated counter parts. 
In other words, inconsistencies have a tendency to lead to smaller p-values.
This potential bias can also be observed in decision inconsistencies. 
Of all decisions inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant value (p \> 0.05).

```{r}
#| label: fig-scatter
#| out-width: 100%
#| fig-cap: "Reported vs. recalculated p-values for inconsistencies that are based on '=' comparions. Black line descriptively indicates the  linear relationship, indicating a bias towards lower reported p-values."

knitr::include_graphics("../plots/figure3.png")

```

## 4.2 Manual inspection of decision inconsistencies
To further understand the nature of decision inconsistencies, we manually inspected all of them (n=655). 
We (a) evaluated whether statcheck has extracted the information correctly, and (b) assessed whether the text actually suggest an erroneous inferential decision, i.e. whether the paper uses a reported significant result to claim a significant effect or a reported non-significant result to claim a null result.

# 5. Discussion and Recommendations

## 5.1. Statistical reporting inconsistencies are prevalent
The present study found a concerning amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[9,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[9,3][[1]]` assessable p-values. 
`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`% of all p-values were flagged as inconsistent and `{r} 100 * round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`% were flagged as decision inconsistencies, i.e. the reported p-value is on the opposite side of the alpha threshold than the recalculated p-value. 
On average, `{r} 100*round(gross_per_article[[1]],2)`% of assessable articles contained at least one inconsistency and `{r} 100*round(gross_per_article[[2]],2)`% contained at least one decision inconsistency.

The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency in psychology [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness; @nuijten2016prevalence; @nuijten2020statcheck; @green2018statcheck; @claesen2023data], medical sciences [@garcia2004incongruence,; @van2023comparing], psychiatry [@berle2007inconsistencies], cyber security studies[@gross2021fidelity], technological education research [@buckley2023estimating], and experimental philosophy [@colombo2018statistical]. 
These studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% decision consistencies.

Even if the prevalence of these inconsistencies could be largely attributed to inconsequential typos or rounding errors (an assumption we cannot test without access to the data), the sheer amount of these errors that have made it through peer-review should concern us. 
They are human errors.
If such a substantial amount of errors is found in plain site, the question naturally arises how many errors that happen during the data analysis itself remain undetected. 
If the tip of the iceberg above water is so large, how large is the iceberg underneath the surface?

The present examination also looked at whether the rates of inconsistency changed over the last 23 years of publications. 
Descriptive assessment did not indicate any noticeable trend.
Moreover, the rates of both inconsistencies and decision inconsistencies seem to remain stable across journals.

Observed inconsistencies were characterized by reported p-values being on average lower than their recalculated counterparts and the prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant. 
This could indicate a systematic bias in favor of lower p-values in general and a bias towards significant results in particular. 
Our data do not speak to the causes of these biases, but possible reasons include the following:
Researchers might intentionally round down p-values because they think a lower p-value is more convincing to reviewers and/or readers.
This practice has been admitted to by 1 in 5 surveyed psychological researchers [@john2012measuring]. 
Given that quantitative linguists have been reported to commit questionable research practices (and even fraud) more often than we would like to [@isbell2022misconduct], we cannot exclude the possibility that some of the inconsistencies were intentional. 
It is our strong belief, however, that the majority of inconsistencies are unintentional.

Researchers might scrutinize non-significant results more than significant results or fail to double check significant results more often because results that confirm their hypothesis feed into their confirmation bias [@nickerson1998confirmation].
For example, @fugelsang2004theory let researchers evaluate data that are either consistent or inconsistent with their prior expectations. 
They showed that when researchers encounter results that disconfirm their expectations, they are likely to blame the methodology while results that confirmed their expectations were rarely critically scrutinized.
Alternatively, the observed bias might be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than non-significant ones.

## 5.2. Limitations of our study
While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, there are, of course, a number of limitations.
The present assessment and the conclusions we can draw from them are limited.
First, our sample is limited to only a subset of experimental linguistic journals.
Our sample is based on a crude criterion of what constitutes an experimental linguistic journal [see @kobrock2023assessing] and we restricted ourselves further to (for us) accessible journals which explicitly require APA statistical formatting in the author guidelines.
Thus it is possible that a different selection of journals would have resulted in different results.
However, given that the inconsistency rates of our study are comparable to similar studies from other disciplines and that the inconsistency rates are both comparable across the eight journals and stable across the most recent 23 years, suggests that our findings are relevant for experimental linguistics at large.

Second, given the constraints on automatically detecting test statistics, statcheck misses reported values that either diverge from APA reporting standards or are reported in tables.
However, inconsistency rates have been shown to be similar for results in APA format vs. results that diverge from APA formatting [@bakker2011mis; @nuijten2016prevalence]. 
Third, statcheck slightly overestimates inconsistency rates, because it might not accurately detect corrections for multiple comparisons [@schmidt2017statcheck].
@nuijten2017validity, however, show that not only were there only a small proportion of flagged inconsistencies related to multiple comparisons, but also that these multiple comparisons themselves were often erroneously reported.
They conclude that "[a]ny reporting inconsistencies associated with these tests and corrections could not explain the high prevalence of reporting inconsistencies" [@nuijten2017validity, p. 27].
More elaborate automatic tools for the extraction of statistical information might allow a more detailed and more accurate assessment of statistical reporting in the future [e.g. @kalmbach2023rule].
But even statcheck is limited and only provides a rough proxy of true inconsistency rate in the published literature, we hope the reader agrees that it is a state of affairs that should be tackled.

## 5.3. Recommendations for the field
There are concrete actionable steps the field of experimental linguistics can make to reduce statistical reporting inconsistencies.
In order to avoid simple copy and paste errors related to working in two separate programs for writing the manuscript and conducting the statistical analysis, authors should consider 'literate programming', i.e. an integration of analysis code and prose into a single, dynamic document [@knuth1984literate; @casillas2023opening]. 
Several implementations of literate programming are freely available to researchers including common R markdown files (Rmd) and Quarto markdown files (qmd).
Literate programming can ensure that values derived by the statistical analysis is automatically integrated into the manuscript document, avoiding errors that might happen during a transfer from one program to the other.

Authors should also consider sharing their derived data (i.e. the anonymized data table that was analyzed) as well as a detailed description of their statistical protocol, ideally in form of reproducible scripts. 
Sharing reproducible analyses with reviewers allows the reviewers the reproduce the authors' analyses, possibly detect errors or even inappropriate statistical choices before publication, thus improving the quality of their work. 
Moreover, publicly sharing their analyses has numerous benefits to the authors themselves beyond error detection: 
Open data and materials can facilitate collaboration [@boland2017ten], increase efficiency and sustainability [@lowndes2017our], and are cited more often [@colavizza2020citation].

If authors do not share data and scripts, reviewers can check at least the statistical reporting consistency in the manuscript by using tools such as statcheck [@nuijten2023statcheck, http://statcheck.io] or p-checker [@schonbrodt2015p, http://shinyapps.org/apps/p-checker/].
Reviewers could consider consider requesting data and scripts during peer review.
Such requests might be particularly justified when inconsistencies are apparent. 
Explicitly requesting to share data might already instill additional care and quality checks when authors prepare their materials, but also allows the reviewers to carefully reproduce the results, and critically evaluate all choices made in the statistical analysis [@sakaluk2014analytic]. 
Recent evidence suggest experimental linguistics are still characterized by a pluralism of statistical approaches, even when trying to answer the same research question [@coretta_multidimensional_2023]. 
Some of these approaches are might be more appropriate than others [@sondereggera2024advancements; @vasishth2023some], so more thorough evaluations of how researchers arrive at their statistical conclusions might elevate their analytical robustness.

Journal editors could explicitly recommend consistency checks with apps such as statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines [Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous]. 
Editors could also demand, recommend or at least encourage data sharing for publication in their journal. 
Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g. @laurinavichyute2022share; @hardwicke2018data].

[^1]: http://www.psychologicalscience.org/publications/psychological_science/ps-submissions; accessed on July 15, 2024.
[^2]: https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines; accessed on on July 15, 2024.

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Appendix

# Title for Appendix

### 2024-07-15 manuscript/manuscript.tex
first draft

\documentclass[
  doc,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{Statistical Inconsistencies in Experimental Linguistics}

\shorttitle{Statistical Inconsistencies}

\usepackage{etoolbox}

\authorsnames{Timo B. Roettger,Dara Leonard Jenssen Etemady}

\affiliation{
{Department of Linguistics \& Scandinavian Studies, University of Oslo}}

\leftheader{Roettger and Etemady}

\abstract{Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed
do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit
in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
Excepteur sint occaecat cupidatat non proident, sunt in culpa qui
officia deserunt mollit anim id est laborum.}

\keywords{statistics, statcheck, reproducibility}

\authornote{\par{\addORCIDlink{Timo B. Roettger}{0000-0000-0000-0001}} 
\par{ }
\par{       }
\par{Correspondence concerning this article should be addressed to Timo
B. Roettger, Email: timo.roettger@gmail.com}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{1. Introduction}\label{introduction}

What we know about human language and its cognitive underpinnings is
often informed by experimental data, data based on which researchers
draw theoretically relevant inferences using common statistical
frameworks. This inference should be transparent in order for other
researchers to critically evaluate the inference process and potentially
detect and correct human errors. In the recent years, the quantitative
sciences have seen repeated calls to become more transparent and
reproducible by sharing data and statistical protocols
(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
\citeproc{ref-roettger2019researcher}{Roettger, 2019}), sharing of
statistical protocols is still rare across the language sciences
(\citeproc{ref-bochynska2023reproducible}{Bochynska et al., 2023}). If
statistical procedures cannot be critically evaluated, human errors
might be left undetected and thus remain uncorrected in the publication
record. If these errors affect the decision procedure of the analysis,
i.e.~whether a hypothesis is rejected or accepted, these errors might
lead to -at best- overconfident, -at worst- false theoretical
conclusions. The present paper will present evidence that the published
literature in experimental linguistics contains a concerning amount of
statistical errors, a state of affairs which warrants more rigorous data
sharing practices.

\section{2. Statistal reporting
inconsistencies}\label{statistal-reporting-inconsistencies}

The null-hypothesis significance testing (NHST) framework is, to date,
the most dominant statistical framework that researchers use to test
hypotheses in the language sciences
(\citeproc{ref-sondereggera2024advancements}{Sonderegger \& Sóskuthy,
2024}). These statistical tests are reported in specific formats which
usually contain a test statistic, the degrees of freedom of that test
(if applicable), and the p-value, representing the probability of
observing the data (or more extreme data) given the null hypothesis
(i.e.~given that the test statistic is zero) (see example 1):\\
(1) F(1, 66) = 3.88, p \textless{} .05\\
Since data and statistical protocol sharing remains rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}), interested readers are left with trusting the authors that
the statistical analysis is run and reported correctly. Whether that
trust is warranted can be assessed, however. The three sets of indices
in (1) have a clearly defined mathematical relationship and can thus be
checked for consistency. An F test with the specified degrees of freedom
and a test-statistic of 3.88 should result in a p-value of 0.053 which
is larger, not smaller, than 0.05. Thus, something is wrong with the
statistical report in (1). Possible reasons for this inconsistency are
manifold: It could be a typo of the comparison sign, i.e.~the authors
meant to use \texttt{=} or \texttt{\textgreater{}} rather than
\texttt{\textless{}}. Alternatively, any of the numbers could be a typo.
Sometimes an error might indicate erroneous rounding (e.g.~0.057 being
rounded down to 0.05). Other times, human error along the data
analytical pipeline might have caused an error down the line. Without
access to data and scripts, it remains unclear to the reader, what has
caused the inconsistency. Such inconsistencies can be particularly
concerning if the calculated p-value and the reported p-values are not
on the same side of the alpha threshold. In NHST, p-values below a
conventionalized alpha threshold, most commonly 0.05, are interpreted as
evidence that the data are sufficiently inconsistent with the null
hypothesis (significant). p-values above that threshold are considered
consistent with the null hypothesis and practically lead to rejecting
the alternative hypothesis (not significant). In (1) above, the reported
p-value suggest a significant result, the p-value derived from the
degrees of freedom and the test statistic suggest a non-significant
result. In the following we refer to these inconsistencies as ``decision
inconsistency''.

This form of inconsistency assessment can be automatically assessed if
statistical tests are reported in an unambiguous format. Recently, a
series of studies used such automatic assessments to evaluate the
prevalence of inconsistent statistical reporting across different
disciplines (\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-green2018statcheck}{Green et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021}; e.g.
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016};
\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020}). For
example, looking at over 250000 p-values published in major psychology
journals, Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016})
found that around 50\% of the articles with statistical results
contained at least one inconsistencies and around 12.5\% contained at
least one ``decision inconsistency''.

To assess the prevalence of statistical-reporting inconsistencies in
experimental linguistics, the present paper conceptually replicates
Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
over 39000 p-values reported in eight experimental linguistic journals
published between 2000 and 2023. We further assess whether the rate of
inconsistency differs across journals, whether that rate has changed
over the course of the last 20 years and whether there is evidence for
bias in these statistical-reporting inconsistencies. We discuss the
results and offer concrete recommendations for authors, reviewers, and
editors to tackle this problem.

\section{3. Method}\label{method}

All quantitative analyses were conducted using R {[}{]} and the r
packages {[}LIST ALL PACKAGES from 01\_Analysis.R and add to
references{]}

\subsection{3.1. Statcheck}\label{statcheck}

We used the R package statcheck (Version 1.4.1-beta.2,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
automatically detect statistical-reporting inconsistencies. Statcheck
works as follows: After converting pdf or html articles to plain text,
statcheck searches for specific strings that correspond to a NHST
result, using ``regular expressions''. That way, statcheck can detect
results of t tests, F tests, Z tests, χ2 tests, correlation tests, and Q
tests as long as the test result fulfills three conditions: (a) the test
result is reported completely including the test statistic, degrees of
freedom (if applicable), and the p-value; (b) the test result is in the
body of the text, i.e.~it usually misses results reported in tables; and
(c) the test result is reported in American Psychological Association
style (APA, 2019). Given these constraints, statcheck is estimated to
detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}) Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-values and, if
there is a mismatch, reports a comparison as an ``inconsistency.'' The
algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional''. Moreover, statcheck counts p = .000
and p \textless{} .000 as inconsistent as p-values of exactly zero are
mathematically impossible and the APA manual
(\citeproc{ref-APA2020}{American Psychological Association, 2020})
advises to report very small p-values as p \textless{} .001. Validity
checks suggest that inter rater reliability between manual coding and
statcheck is high, i.e.~0.76 for inconsistencies and 0.89 for decision
inconsistencies (\citeproc{ref-nuijten2016prevalence}{Nuijten et al.,
2016}). The overall accuracy of statcheck is estimated between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider statcheck a valid, but rough proxy of statistical
reporting inconsistencies.

\subsection{3.2. Sample}\label{sample}

Focusing on experimental linguistic research, we started with Kobrock
and Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the term ``experiment* in title, abstract or
keywords. Out of these 100 journals, we selected all journals with at
least 10\% of articles containing the search string''experiment*``. Out
of the remaining 37 journals, we selected those journals that urged APA
formatting either in the main body of the text or specifically regarding
statistics in the author guidelines, resulting in nine remaining
journals. Moreover, to download the pdfs, the articles had to be either
accessible to us through our library license, or open access, resulting
in a final list of eight journals: Applied Psycholinguistics (APS),
Bilingualism: Language and Cognition (BLC), Linguistic Approaches to
Bilingualism (LAB), Language and Speech (LaS), Language Learning and
Techology (LLT), Journal of Language and Social Psychology (LSP),
Journal of Child Language (JCL), and Studies in Second Language
Acquisition (SLA).

We included only original research articles within the publication years
of 2000-2023, excluding any book reviews, response articles,
commentaries, editorials, corrigenda, errata, advertisements, etc.
Articles from LAB spanned 2011-2023, while the rest spanned 2000-2023.
This procedure resulted in 5804 research articles. All 5804 articles
were submitted to analysis, using Statcheck v1.5.0. 157 articles were
not analyzable by statcheck, and thus had to be removed from the pool,
likely related to issues with rendering the Chi-Squared symbol being
erroneously converted from .pdf to .txt.

\subsection{3.3. Data availability
statement}\label{data-availability-statement}

All derived data and corresponding R scripts are publicly available
here: LINK. The original journal articles cannot be bulk-shared due to
distribution restrictions by the publishers. NEED TO FIGURE OUT A WAY

\section{4. Results}\label{results}

\subsection{4.1 Prevalence of
inconsistencies}\label{prevalence-of-inconsistencies}

The results are summarized in Table~\ref{tbl-summary}. Out of 5804
articles, 3059 articles contained statistical tests that statcheck could
assess (53\%), amounting to 39532 assessable p-values. 5166 p-values
were flagged as inconsistent (13.1\%) and 655 of which were considered
decisions inconsistencies (1.7\%) (see Table~\ref{tbl-summary}).

\begin{table}

{\caption{{Summary. Applied Psycholinguistics (APS), Bilingualism:
Language and Cognition (BLC), Linguistic Approaches to Bilingualism
(LAB), Language and Speech (LaS), Language Learning and Techology (LLT),
Journal of Language and Social Psychology (LSP), Journal of Child
Language (JCL), and Studies in Second Language Acquisition
(SLA).}{\label{tbl-summary}}}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0755}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1698}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1887}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1792}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1509}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2358}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Journal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
eligible articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessible articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessible results
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
inconsistencies
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
decision inconsistencies
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
APS & 953 & 690 & 9570 & 1368 & 170 \\
BLC & 964 & 610 & 9093 & 1161 & 120 \\
JCL & 1109 & 529 & 6240 & 750 & 69 \\
LAB & 471 & 133 & 1719 & 234 & 25 \\
LLT & 421 & 111 & 919 & 201 & 61 \\
LSP & 695 & 376 & 4320 & 429 & 60 \\
LaS & 598 & 363 & 4717 & 552 & 86 \\
SLA & 593 & 247 & 2954 & 471 & 64 \\
Total & 5804 & 3059 & 39532 & 5166 & 655 \\
\end{longtable}

\end{table}

The proportion of inconsistencies ranged from 10 to 22\% across journals
(1 to 7\% for decisions inconsistencies) (see
Figure~\ref{fig-stacked-bar}). These rates appear to be stable across
year of publication (see Figure~\ref{fig-year-prop}). On average, 50\%
of assessable articles contained one or more inconsistencies (journals
range from 43 to 56\%) and 13\% contained one or more decisions
inconsistencies (journals range from 11 to 15\%).

\begin{figure}[H]

\caption{\label{fig-stacked-bar}Proportion of articles containing at
least one inconsistency / decision inconsistency. Applied
Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
Language Learning and Techology (LLT), Journal of Language and Social
Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
Language Acquisition (SLA).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{../plots/figure1.png}

}

\end{figure}%

\begin{figure}

\caption{\label{fig-year-prop}Proportion of inconsistencies / decision
inconsistencies from 2000 to 2023. Left: global rates with all journals
pooled; Right: rates per journal. Applied Psycholinguistics (APS),
Bilingualism: Language and Cognition (BLC), Linguistic Approaches to
Bilingualism (LAB), Language and Speech (LaS), Language Learning and
Techology (LLT), Journal of Language and Social Psychology (LSP),
Journal of Child Language (JCL), and Studies in Second Language
Acquisition (SLA).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{../plots/figure2.png}

}

\end{figure}%

When examining reported against recalculated p-values (see
Figure~\ref{fig-scatter}) for all inconsistencies, we can identify
certain spatial patterns (only inconsistencies that are based on `='
comparisons were included to make them easier to interpret, n = 5154 out
of 5166). First, the center of density of points is located in the
bottom left corner with more values reported closer to the alpha level.
This is not surprising, since there is a documented publication bias in
published quantitative articles, with hypotheses being much more often
confirmed (and thus p \textless{} .05) than not
(\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}). Second, there
is a linear density of points along the diagonal axis which corresponds
to numerically small inconsistencies, some of which might be related to
simple rounding errors. However, comparing the diagonal to the black
line, which represents a linear model predicting recalculated by
reported p-values, we can see a clear divergence of what is expected if
inconsistencies were equally likely in both directions
(i.e.~recalculated p-values were as likely to be smaller than the
reported p-value as larger). The slope of the regression line is flatter
than the diagonal axis which means that, on average, reported p-values
are lower than their recalculated counter parts. In other words,
inconsistencies have a tendency to lead to smaller p-values. This
potential bias can also be observed in decision inconsistencies. Of all
decisions inconsistencies (n = 655), 71\% represent cases in which a
reported significant result (p \textless{} 0.05) is recalculated as
non-significant value (p \textgreater{} 0.05).

\begin{figure}

\caption{\label{fig-scatter}Reported vs.~recalculated p-values for
inconsistencies that are based on `=' comparions. Black line
descriptively indicates the linear relationship, indicating a bias
towards lower reported p-values.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{../plots/figure3.png}

}

\end{figure}%

\subsection{4.2 Manual inspection of decision
inconsistencies}\label{manual-inspection-of-decision-inconsistencies}

To further understand the nature of decision inconsistencies, we
manually inspected all of them (n=655). We (a) evaluated whether
statcheck has extracted the information correctly, and (b) assessed
whether the text actually suggest an erroneous inferential decision,
i.e.~whether the paper uses a reported significant result to claim a
significant effect or a reported non-significant result to claim a null
result.

\section{5. Discussion and
Recommendations}\label{discussion-and-recommendations}

\subsection{5.1. Statistical reporting inconsistencies are
prevalent}\label{statistical-reporting-inconsistencies-are-prevalent}

The present study found a concerning amount of statistical reporting
inconsistencies across a sample of 5804 experimental linguistic
articles, containing 3059 assessable p-values. 13.1\% of all p-values
were flagged as inconsistent and 1.7\% were flagged as decision
inconsistencies, i.e.~the reported p-value is on the opposite side of
the alpha threshold than the recalculated p-value. On average, 50\% of
assessable articles contained at least one inconsistency and 13\%
contained at least one decision inconsistency.

The present study can be considered a conceptual replication of previous
studies investigating statistical-reporting inconsistency in psychology
(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011},
\citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-claesen2023data}{Claesen et al., 2023};
\citeproc{ref-green2018statcheck}{Green et al., 2018};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016};
\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}), medical
sciences (\citeproc{ref-garcia2004incongruence}{Garcı́a-Berthou \&
Alcaraz, 2004},; \citeproc{ref-van2023comparing}{Van Aert et al.,
2023}), psychiatry (\citeproc{ref-berle2007inconsistencies}{Berle \&
Starcevic, 2007}), cyber security
studies(\citeproc{ref-gross2021fidelity}{Groß, 2021}), technological
education research (\citeproc{ref-buckley2023estimating}{Buckley et al.,
2023}), and experimental philosophy
(\citeproc{ref-colombo2018statistical}{Colombo et al., 2018}). These
studies report on inconsistency rates between 4\% and 14\%, with between
10\% and 63\% of articles containing at least one inconsistency and
between 3\% and 21\% decision consistencies.

Even if the prevalence of these inconsistencies could be largely
attributed to inconsequential typos or rounding errors (an assumption we
cannot test without access to the data), the sheer amount of these
errors that have made it through peer-review should concern us. They are
human errors. If such a substantial amount of errors is found in plain
site, the question naturally arises how many errors that happen during
the data analysis itself remain undetected. If the tip of the iceberg
above water is so large, how large is the iceberg underneath the
surface?

The present examination also looked at whether the rates of
inconsistency changed over the last 23 years of publications.
Descriptive assessment did not indicate any noticeable trend. Moreover,
the rates of both inconsistencies and decision inconsistencies seem to
remain stable across journals.

Observed inconsistencies were characterized by reported p-values being
on average lower than their recalculated counterparts and the prevalence
of decision inconsistencies was higher for p-values reported as
significant than for those reported as non-significant. This could
indicate a systematic bias in favor of lower p-values in general and a
bias towards significant results in particular. Our data do not speak to
the causes of these biases, but possible reasons include the following:
Researchers might intentionally round down p-values because they think a
lower p-value is more convincing to reviewers and/or readers. This
practice has been admitted to by 1 in 5 surveyed psychological
researchers (\citeproc{ref-john2012measuring}{John et al., 2012}). Given
that quantitative linguists have been reported to commit questionable
research practices (and even fraud) more often than we would like to
(\citeproc{ref-isbell2022misconduct}{Isbell et al., 2022}), we cannot
exclude the possibility that some of the inconsistencies were
intentional. It is our strong belief, however, that the majority of
inconsistencies are unintentional.

Researchers might scrutinize non-significant results more than
significant results or fail to double check significant results more
often because results that confirm their hypothesis feed into their
confirmation bias (\citeproc{ref-nickerson1998confirmation}{Nickerson,
1998}). For example, Fugelsang et al.
(\citeproc{ref-fugelsang2004theory}{2004}) let researchers evaluate data
that are either consistent or inconsistent with their prior
expectations. They showed that when researchers encounter results that
disconfirm their expectations, they are likely to blame the methodology
while results that confirmed their expectations were rarely critically
scrutinized. Alternatively, the observed bias might be a reflection of
publication bias (\citeproc{ref-franco2014publication}{Franco et al.,
2014}; \citeproc{ref-sterling1959publication}{Sterling, 1959}) with
(erroneously) reported significant p-values being more likely to be
published than non-significant ones.

\subsection{5.2. Limitations of our
study}\label{limitations-of-our-study}

While we believe our work offers an important contribution to improving
statistical reporting practices in experimental linguistics, there are,
of course, a number of limitations. The present assessment and the
conclusions we can draw from them are limited. First, our sample is
limited to only a subset of experimental linguistic journals. Our sample
is based on a crude criterion of what constitutes an experimental
linguistic journal (see \citeproc{ref-kobrock2023assessing}{Kobrock \&
Roettger, 2023}) and we restricted ourselves further to (for us)
accessible journals which explicitly require APA statistical formatting
in the author guidelines. Thus it is possible that a different selection
of journals would have resulted in different results. However, given
that the inconsistency rates of our study are comparable to similar
studies from other disciplines and that the inconsistency rates are both
comparable across the eight journals and stable across the most recent
23 years, suggests that our findings are relevant for experimental
linguistics at large.

Second, given the constraints on automatically detecting test
statistics, statcheck misses reported values that either diverge from
APA reporting standards or are reported in tables. However,
inconsistency rates have been shown to be similar for results in APA
format vs.~results that diverge from APA formatting
(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Third,
statcheck slightly overestimates inconsistency rates, because it might
not accurately detect corrections for multiple comparisons
(\citeproc{ref-schmidt2017statcheck}{Schmidt, 2017}). Nuijten et al.
(\citeproc{ref-nuijten2017validity}{2017}), however, show that not only
were there only a small proportion of flagged inconsistencies related to
multiple comparisons, but also that these multiple comparisons
themselves were often erroneously reported. They conclude that
``{[}a{]}ny reporting inconsistencies associated with these tests and
corrections could not explain the high prevalence of reporting
inconsistencies'' (\citeproc{ref-nuijten2017validity}{Nuijten et al.,
2017, p. 27}). More elaborate automatic tools for the extraction of
statistical information might allow a more detailed and more accurate
assessment of statistical reporting in the future (e.g.
\citeproc{ref-kalmbach2023rule}{Kalmbach et al., 2023}). But even
statcheck is limited and only provides a rough proxy of true
inconsistency rate in the published literature, we hope the reader
agrees that it is a state of affairs that should be tackled.

\subsection{5.3. Recommendations for the
field}\label{recommendations-for-the-field}

There are concrete actionable steps the field of experimental
linguistics can make to reduce statistical reporting inconsistencies. In
order to avoid simple copy and paste errors related to working in two
separate programs for writing the manuscript and conducting the
statistical analysis, authors should consider `literate programming',
i.e.~an integration of analysis code and prose into a single, dynamic
document (\citeproc{ref-casillas2023opening}{Casillas et al., 2023};
\citeproc{ref-knuth1984literate}{Knuth, 1984}). Several implementations
of literate programming are freely available to researchers including
common R markdown files (Rmd) and Quarto markdown files (qmd). Literate
programming can ensure that values derived by the statistical analysis
is automatically integrated into the manuscript document, avoiding
errors that might happen during a transfer from one program to the
other.

Authors should also consider sharing their derived data (i.e.~the
anonymized data table that was analyzed) as well as a detailed
description of their statistical protocol, ideally in form of
reproducible scripts. Sharing reproducible analyses with reviewers
allows the reviewers the reproduce the authors' analyses, possibly
detect errors or even inappropriate statistical choices before
publication, thus improving the quality of their work. Moreover,
publicly sharing their analyses has numerous benefits to the authors
themselves beyond error detection: Open data and materials can
facilitate collaboration (\citeproc{ref-boland2017ten}{Boland et al.,
2017}), increase efficiency and sustainability
(\citeproc{ref-lowndes2017our}{Lowndes et al., 2017}), and are cited
more often (\citeproc{ref-colavizza2020citation}{Colavizza et al.,
2020}).

If authors do not share data and scripts, reviewers can check at least
the statistical reporting consistency in the manuscript by using tools
such as statcheck (\citeproc{ref-nuijten2023statcheck}{Nuijten \&
Epskamp, 2023}, http://statcheck.io) or p-checker
(\citeproc{ref-schonbrodt2015p}{Schönbrodt, 2015},
http://shinyapps.org/apps/p-checker/). Reviewers could consider consider
requesting data and scripts during peer review. Such requests might be
particularly justified when inconsistencies are apparent. Explicitly
requesting to share data might already instill additional care and
quality checks when authors prepare their materials, but also allows the
reviewers to carefully reproduce the results, and critically evaluate
all choices made in the statistical analysis
(\citeproc{ref-sakaluk2014analytic}{Sakaluk et al., 2014}). Recent
evidence suggest experimental linguistics are still characterized by a
pluralism of statistical approaches, even when trying to answer the same
research question (\citeproc{ref-coretta_multidimensional_2023}{Coretta
et al., 2023}). Some of these approaches are might be more appropriate
than others (\citeproc{ref-sondereggera2024advancements}{Sonderegger \&
Sóskuthy, 2024}; \citeproc{ref-vasishth2023some}{Vasishth, 2023}), so
more thorough evaluations of how researchers arrive at their statistical
conclusions might elevate their analytical robustness.

Journal editors could explicitly recommend consistency checks with apps
such as statcheck during peer review, a practice that has been taken up
on by several journals from neighboring disciplines (Psychological
Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
  accessed on July 15, 2024.}, Advances in Methods and Practices in
Psychological Science\footnote{https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines;
  accessed on on July 15, 2024.}, Stress \& Health
\citeproc{ref-barber2017meticulous}{Barber, 2017}). Editors could also
demand, recommend or at least encourage data sharing for publication in
their journal. Data sharing policies have been shown to substantially
increase the reproducibility of analyses
(\citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018}; e.g.
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022}).

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-APA2020}
American Psychological Association. (2020). \emph{Publication manual of
the {American Psychological Association}} (7th ed.). Author.
\url{https://doi.org/10.1037/0000173-000}

\bibitem[\citeproctext]{ref-arvan2022reproducibility}
Arvan, M., Pina, L., \& Parde, N. (2022). Reproducibility in
computational linguistics: Is source code enough? \emph{Proceedings of
the 2022 Conference on Empirical Methods in Natural Language
Processing}, 2350--2361.

\bibitem[\citeproctext]{ref-bakker2011mis}
Bakker, M., \& Wicherts, J. M. (2011). The (mis) reporting of
statistical results in psychology journals. \emph{Behavior Research
Methods}, \emph{43}, 666--678.

\bibitem[\citeproctext]{ref-bakker2014outlier}
Bakker, M., \& Wicherts, J. M. (2014). Outlier removal and the relation
with reporting errors and quality of psychological research. \emph{PloS
One}, \emph{9}(7), e103360.

\bibitem[\citeproctext]{ref-barber2017meticulous}
Barber, L. K. (2017). Meticulous manuscripts, messy results: Working
together for robust science reporting. \emph{Stress \& Health},
\emph{33}(2), 89--91.

\bibitem[\citeproctext]{ref-berle2007inconsistencies}
Berle, D., \& Starcevic, V. (2007). Inconsistencies between reported
test statistics and p-values in two psychiatry journals.
\emph{International Journal of Methods in Psychiatric Research},
\emph{16}(4), 202--207.

\bibitem[\citeproctext]{ref-bochynska2023reproducible}
Bochynska, A., Keeble, L., Halfacre, C., Casillas, J. V., Champagne,
I.-A., Chen, K., Röthlisberger, M., Buchanan, E. M., \& Roettger, T.
(2023). Reproducible research practices and transparency across
linguistics. \emph{Glossa Psycholinguistics}, \emph{2}(1).

\bibitem[\citeproctext]{ref-boland2017ten}
Boland, M. R., Karczewski, K. J., \& Tatonetti, N. P. (2017). Ten simple
rules to enable multi-site collaborations through data sharing. In
\emph{PLoS computational biology} (1; Vol. 13, p. e1005278). Public
Library of Science San Francisco, CA USA.

\bibitem[\citeproctext]{ref-buckley2023estimating}
Buckley, J., Hyland, T., \& Seery, N. (2023). Estimating the
replicability of technology education research. \emph{International
Journal of Technology and Design Education}, \emph{33}(4), 1243--1264.

\bibitem[\citeproctext]{ref-caperos2013consistency}
Caperos, J. M., \& Pardo, A. (2013). Consistency errors in p-values
reported in spanish psychology journals. \emph{Psicothema},
\emph{25}(3), 408--414.

\bibitem[\citeproctext]{ref-casillas2023opening}
Casillas, J. V., Constantin-Dureci, G., Rascón, I. A., Shao, J.,
Rodrı́guez, S. A., Gadamsetty, A., Minetti, A., Laungani, K., Thatcher,
J., Gardere, R.-T., et al. (2023). \emph{Opening open science to all:
Demystifying reproducibility and transparency practices in linguistic
research}.

\bibitem[\citeproctext]{ref-claesen2023data}
Claesen, A., Vanpaemel, W., Maerten, A.-S., Verliefde, T., Tuerlinckx,
F., \& Heyman, T. (2023). Data sharing upon request and statistical
consistency errors in psychology: A replication of wicherts, bakker and
molenaar (2011). \emph{Plos One}, \emph{18}(4), e0284243.

\bibitem[\citeproctext]{ref-colavizza2020citation}
Colavizza, G., Hrynaszkiewicz, I., Staden, I., Whitaker, K., \&
McGillivray, B. (2020). The citation advantage of linking publications
to research data. \emph{PloS One}, \emph{15}(4), e0230416.

\bibitem[\citeproctext]{ref-colombo2018statistical}
Colombo, M., Duev, G., Nuijten, M. B., \& Sprenger, J. (2018).
Statistical reporting inconsistencies in experimental philosophy.
\emph{PloS One}, \emph{13}(4), e0194360.

\bibitem[\citeproctext]{ref-coretta_multidimensional_2023}
Coretta, S., Casillas, J. V., Roessig, S., Franke, M., Ahn, B.,
Al-Hoorie, A. H., Al-Tamimi, J., Alotaibi, N. E., AlShakhori, M. K.,
Altmiller, R. M., Arantes, P., Athanasopoulou, A., Baese-Berk, M. M.,
Bailey, G., Sangma, C. B. A., Beier, E. J., Benavides, G. M., Benker,
N., BensonMeyer, E. P., \ldots{} Roettger, T. B. (2023).
Multidimensional {Signals} and {Analytic} {Flexibility}: {Estimating}
{Degrees} of {Freedom} in {Human}-{Speech} {Analyses}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{6}(3),
25152459231162567. \url{https://doi.org/10.1177/25152459231162567}

\bibitem[\citeproctext]{ref-franco2014publication}
Franco, A., Malhotra, N., \& Simonovits, G. (2014). Publication bias in
the social sciences: Unlocking the file drawer. \emph{Science},
\emph{345}(6203), 1502--1505.

\bibitem[\citeproctext]{ref-fugelsang2004theory}
Fugelsang, J. A., Stein, C. B., Green, A. E., \& Dunbar, K. N. (2004).
Theory and data interactions of the scientific mind: Evidence from the
molecular and the cognitive laboratory. \emph{Canadian Journal of
Experimental Psychology/Revue Canadienne de Psychologie
Exp{é}rimentale}, \emph{58}(2), 86.

\bibitem[\citeproctext]{ref-garcia2004incongruence}
Garcı́a-Berthou, E., \& Alcaraz, C. (2004). Incongruence between test
statistics and p values in medical papers. \emph{BMC Medical Research
Methodology}, \emph{4}, 1--5.

\bibitem[\citeproctext]{ref-green2018statcheck}
Green, C. D., Abbas, S., Belliveau, A., Beribisky, N., Davidson, I. J.,
DiGiovanni, J., Heidari, C., Martin, S. M., Oosenbrug, E., \&
Wainewright, L. M. (2018). Statcheck in canada: What proportion of CPA
journal articles contain errors in the reporting of p-values?
\emph{Canadian Psychology/Psychologie Canadienne}, \emph{59}(3), 203.

\bibitem[\citeproctext]{ref-gross2021fidelity}
Groß, T. (2021). Fidelity of statistical reporting in 10 years of cyber
security user studies. \emph{Socio-Technical Aspects in Security and
Trust: 9th International Workshop, STAST 2019, Luxembourg City,
Luxembourg, September 26, 2019, Revised Selected Papers 9}, 3--26.

\bibitem[\citeproctext]{ref-hardwicke2018data}
Hardwicke, T. E., Mathur, M. B., MacDonald, K., Nilsonne, G., Banks, G.
C., Kidwell, M. C., Hofelich Mohr, A., Clayton, E., Yoon, E. J., Henry
Tessler, M., et al. (2018). Data availability, reusability, and analytic
reproducibility: Evaluating the impact of a mandatory open data policy
at the journal cognition. \emph{Royal Society Open Science},
\emph{5}(8), 180448.

\bibitem[\citeproctext]{ref-isbell2022misconduct}
Isbell, D. R., Brown, D., Chen, M., Derrick, D. J., Ghanem, R., Arvizu,
M. N. G., Schnur, E., Zhang, M., \& Plonsky, L. (2022). Misconduct and
questionable research practices: The ethics of quantitative data
handling and reporting in applied linguistics. \emph{The Modern Language
Journal}, \emph{106}(1), 172--195.

\bibitem[\citeproctext]{ref-john2012measuring}
John, L. K., Loewenstein, G., \& Prelec, D. (2012). Measuring the
prevalence of questionable research practices with incentives for truth
telling. \emph{Psychological Science}, \emph{23}(5), 524--532.

\bibitem[\citeproctext]{ref-kalmbach2023rule}
Kalmbach, T., Hoffmann, M., Lell, N., \& Scherp, A. (2023). On the
rule-based extraction of statistics reported in scientific papers.
\emph{International Conference on Applications of Natural Language to
Information Systems}, 326--338.

\bibitem[\citeproctext]{ref-knuth1984literate}
Knuth, D. E. (1984). Literate programming. \emph{The Computer Journal},
\emph{27}(2), 97--111.

\bibitem[\citeproctext]{ref-kobrock2023assessing}
Kobrock, K., \& Roettger, T. (2023). Assessing the replication landscape
in experimental linguistics. \emph{Glossa Psycholinguistics},
\emph{2}(1), 1--28.

\bibitem[\citeproctext]{ref-laurinavichyute2022share}
Laurinavichyute, A., Yadav, H., \& Vasishth, S. (2022). Share the code,
not just the data: A case study of the reproducibility of articles
published in the journal of memory and language under the open data
policy. \emph{Journal of Memory and Language}, \emph{125}, 104332.

\bibitem[\citeproctext]{ref-lowndes2017our}
Lowndes, J. S. S., Best, B. D., Scarborough, C., Afflerbach, J. C.,
Frazier, M. R., O'Hara, C. C., Jiang, N., \& Halpern, B. S. (2017). Our
path to better science in less time using open data science tools.
\emph{Nature Ecology \& Evolution}, \emph{1}(6), 0160.

\bibitem[\citeproctext]{ref-nickerson1998confirmation}
Nickerson, R. S. (1998). Confirmation bias: A ubiquitous phenomenon in
many guises. \emph{Review of General Psychology}, \emph{2}(2), 175--220.

\bibitem[\citeproctext]{ref-nuijten2017validity}
Nuijten, M. B., Assen, M. A. van, Hartgerink, C., Epskamp, S., \&
Wicherts, J. M. (2017). \emph{The validity of the tool {``statcheck''}
in discovering statistical reporting inconsistencies}.

\bibitem[\citeproctext]{ref-nuijten2023statcheck}
Nuijten, M. B., \& Epskamp, S. (2023). \emph{Statcheck: Extract
statistics from articles and recompute p-values(1.4. 1-beta. 2){[}r{]}}.

\bibitem[\citeproctext]{ref-nuijten2016prevalence}
Nuijten, M. B., Hartgerink, C. H., Van Assen, M. A., Epskamp, S., \&
Wicherts, J. M. (2016). The prevalence of statistical reporting errors
in psychology (1985--2013). \emph{Behavior Research Methods}, \emph{48},
1205--1226.

\bibitem[\citeproctext]{ref-nuijten2020statcheck}
Nuijten, M. B., \& Polanin, J. R. (2020). {``Statcheck''}: Automatically
detect statistical reporting inconsistencies to increase reproducibility
of meta-analyses. \emph{Research Synthesis Methods}, \emph{11}(5),
574--579.

\bibitem[\citeproctext]{ref-roettger2019researcher}
Roettger, T. B. (2019). Researcher degrees of freedom in phonetic
research. \emph{Laboratory Phonology}, \emph{10}(1).

\bibitem[\citeproctext]{ref-sakaluk2014analytic}
Sakaluk, J., Williams, A., \& Biernat, M. (2014). Analytic review as a
solution to the misreporting of statistical results in psychological
science. \emph{Perspectives on Psychological Science}, \emph{9}(6),
652--660.

\bibitem[\citeproctext]{ref-schmidt2017statcheck}
Schmidt, T. (2017). \emph{Statcheck does not work: All the numbers.
Reply to nuijten et al.(2017)}.

\bibitem[\citeproctext]{ref-schonbrodt2015p}
Schönbrodt, F. D. (2015). \emph{P-checker: One-for-all p-value
analyzer}. \url{http://shinyapps.org/apps/p-checker/}.

\bibitem[\citeproctext]{ref-sondereggera2024advancements}
Sonderegger, M., \& Sóskuthy, M. (2024). \emph{Advancements of phonetics
in the 21st century: Quantitative data analysis}.

\bibitem[\citeproctext]{ref-sterling1959publication}
Sterling, T. D. (1959). Publication decisions and their possible effects
on inferences drawn from tests of significance---or vice versa.
\emph{Journal of the American Statistical Association}, \emph{54}(285),
30--34.

\bibitem[\citeproctext]{ref-van2023comparing}
Van Aert, R. C., Nuijten, M. B., Olsson-Collentine, A., Stoevenbelt, A.
H., Van Den Akker, O. R., Klein, R. A., \& Wicherts, J. M. (2023).
Comparing the prevalence of statistical reporting inconsistencies in
COVID-19 preprints and matched controls: A registered report.
\emph{Royal Society Open Science}, \emph{10}(8), 202326.

\bibitem[\citeproctext]{ref-vasishth2023some}
Vasishth, S. (2023). Some right ways to analyze (psycho) linguistic
data. \emph{Annual Review of Linguistics}, \emph{9}(1), 273--291.

\bibitem[\citeproctext]{ref-veldkamp2014statistical}
Veldkamp, C. L., Nuijten, M. B., Dominguez-Alvarez, L., Van Assen, M.
A., \& Wicherts, J. M. (2014). Statistical reporting errors and
collaboration on statistical analyses in psychological science.
\emph{PloS One}, \emph{9}(12), e114876.

\bibitem[\citeproctext]{ref-wicherts2011willingness}
Wicherts, J. M., Bakker, M., \& Molenaar, D. (2011). Willingness to
share research data is related to the strength of the evidence and the
quality of reporting of statistical results. \emph{PloS One},
\emph{6}(11), e26828.

\end{CSLReferences}

\appendix

\section{Title for Appendix}\label{title-for-appendix}

\end{document}

### 2024-07-15 scripts/Error_Ling_Analysis.qmd
first draft

---
title: "Statistical Inconsistencies in Experimental Linguistics"
shorttitle: "Statistical Inconsistencies"
author:
  - name: Timo Roettger
    corresponding: true
    orcid: 0000-0000-0000-0001
    email: timo.roettger@gmail.com
    affiliations:
      - name: University of Oslo
        department: Department of Linguistics & Scandinavian Studies
        city: Oslo
        country: Norway
        postal-code: 0175
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "This document is a template."
keywords: [keyword1, keyword2, keyword3]
bibliography: bibliography.bib
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
---

```{r libraries}
#| echo: false
#| output: false

# load in relevant libraries, install if not installed
if (!require("pacman")) install.packages("pacman")
pacman::p_load(rstudioapi, 
               tidyverse, 
               ggpubr, 
               broom, 
               statcheck,
               janitor,
               # brms,
               # cmdstanr,
               # tidybayes,
               patchwork
               )

# install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos")))

```

```{r data}

#| echo: false
#| output: false

## load in data from location of script
#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
#setwd(current_working_dir)

## list all journal units
files <- list.files("../Journals", recursive = TRUE, pattern = "\\.pdf")

## how many?
length(files)
# 5804

## summary table
files_wrangled <- as.data.frame(files) %>%
  mutate(year = substring(files, 5, 8)) %>%
  filter(year != 2024) %>%
  mutate(journal = substring(files, 13, 15)) %>%
  group_by(journal) %>%
  summarise(n.art = length(unique(files)))

## view
files_wrangled

## double check
sum(files_wrangled$n.art) == length(files)

n_articles = length(files)

```

```{r statcheck_extraction}

#| echo: false
#| output: false

## read pdfs in /Journals with statcheck, checks for one tailed tests
## can only be run with access to original articles, bulk sharing is not prohibited by the journals
#stat <- checkPDFdir("Journals", OneTailedTxt = TRUE)

## write statcheck data to csv
#write.csv(stat, "../data/statcheck_data.csv", row.names = FALSE)

xdata <- read.csv("../data/statcheck_data.csv")

```

```{r wrangle}

#| echo: false
#| output: false

# split strings into year and journal, drop NAs
xdata <- xdata %>% 
  mutate(year = strtoi(substring(source, 1, 4)),
         journal = substring(source, 9, 11)) %>%
  drop_na(reported_p) %>%
  drop_na(computed_p) %>%
  filter(year != 2024)

```

```{r general_summary}

#| echo: false
#| output: false

## summary table counts
journal_summary_counts <- xdata %>% 
  group_by(journal) %>% 
  summarise( 
            checked_articles = length(unique(source)),
            checked_results = n(),
            errors = sum(error),
            gross_errors = sum(decision_error)
          ) %>% 
  full_join(files_wrangled) %>% 
  adorn_totals(name = "Total") %>% 
  select(journal, n.art, checked_articles, checked_results, errors, gross_errors)

colnames(journal_summary_counts) <- c("Journal", "eligible articles", "assessible articles",
                                      "assessible results", "inconsistencies", "decision inconsistencies")

#journal_summary_counts

```

```{r}

#| echo: false
#| output: true

knitr::kable(journal_summary_counts)

```

```{r}

#| echo: false
#| output: false

## summary table proportions
journal_summary_prop <- xdata %>% 
  group_by(journal) %>% 
  summarise( 
            checked_articles = length(unique(source)),
            checked_results = n(),
            errors = sum(error),
            gross_errors = sum(decision_error),
            prop_errors = errors / checked_results,
            prop_gross = gross_errors / checked_results
          ) %>% 
  full_join(files_wrangled) %>% 
  group_by(journal) %>% 
  mutate(prop_articles =  checked_articles / n.art) %>% 
  select(-checked_articles, -checked_results, -errors, -gross_errors, -n.art) %>% 
  adorn_totals(name = "Mean") %>%
    mutate(across(where(is.numeric), 
          ~ replace(., n(), .[n()]/(n()-1))))

## now proportion of checkable articles that contain 1 or more errors and 1 or more decision errors
 per_article <- xdata %>% 
  group_by(journal, source) %>% 
  summarise(checked_results = n(),
            errors = sum(error),
            gross_errors = sum(decision_error),
            has_error = ifelse(errors > 0, 1, 0),
            has_gross = ifelse(gross_errors > 0, 1, 0)
          ) %>% 
   ungroup() %>% 
   group_by(journal) %>% 
   summarise(prop_has_error = mean(has_error),
             prop_has_gross = mean(has_gross)
             )

```

```{r proportions_per_article}

#| echo: false
#| output: true

# into long format for plot
per_article_long <- per_article %>% 
  pivot_longer(!journal, names_to = "type", values_to = "proportion")

# plot as stacked barplot
ggplot(per_article_long) +
  geom_bar(data = per_article_long,
           aes(x = journal, 
               y = 50,
               fill = type),
           stat = "identity")  +
   geom_bar(data = per_article_long,
           aes(x = journal, 
               y = 50),
           fill = "grey",
           stat = "identity")  + 
   geom_bar(data = per_article_long %>% filter(type == "prop_has_error"),
           aes(x = journal, 
               y = proportion * 100,
               fill = type),
           fill = "#6002ee",
           stat = "identity") +  
  geom_text(data = per_article_long %>% filter(type == "prop_has_error"),
           aes(x = journal, 
               y = 30,
               label = paste0(round(proportion * 100, 0), "%")),
           color = "white") +
   geom_bar(data = per_article_long %>% filter(type == "prop_has_gross"),
           aes(x = journal, 
               y = proportion * 100),
           fill = "#ee6002",
           stat = "identity") +
  geom_text(data = per_article_long %>% filter(type == "prop_has_gross"),
           aes(x = journal, 
               y = 5,
               label = paste0(round(proportion * 100, 0), "%")),
           vjust = 0,
           color = "white") +
  scale_fill_manual(name = "",
                     values = c("prop_has_error" = "#6002ee", "prop_has_gross" = "#ee6002"),
                     labels = c("inconsistencies", "decision inconsistencies")) +
  labs(x = "\njournal",
       y = "proportion of articles \nwith at least one inconsistency\n") + 
  scale_y_continuous(limits = c(0,100)) +
  theme_minimal()

```

```{r proportions_time}

#| echo: false
#| output: true

# aggregate for proportion over year
year_summary_prop <- xdata %>% 
  group_by(year) %>% 
  summarise( 
            checked_articles = length(unique(source)),
            checked_results = n(),
            errors = sum(error),
            gross_errors = sum(decision_error),
            prop_errors = errors / checked_results,
            prop_gross = gross_errors / checked_results
          )

# aggregate for proportion over year per journal
year_journal_prop <- xdata %>% 
  group_by(year, journal) %>% 
  summarise( 
            checked_articles = length(unique(source)),
            checked_results = n(),
            errors = sum(error),
            gross_errors = sum(decision_error),
            prop_errors = errors / checked_results,
            prop_gross = gross_errors / checked_results
          )

# plot prop over time
year_plot <- 
  ggplot(year_summary_prop) + 
  geom_point(aes(x = year, 
                 y = prop_errors * 100),
             pch = 21,
             size = 2,
             color = "black",
             fill = "#6002ee") +
  geom_path(aes(x = year, 
                 y = prop_errors* 100),
             color = "#6002ee") +
  geom_point(aes(x = year, 
                 y = prop_gross * 100
                 ),
             pch = 21,
             size = 2,
             color = "black",
             fill = "#ee6002") +
  geom_path(aes(x = year, 
                 y = prop_gross* 100),
             color = "#ee6002") +
  annotate("segment", 
           x = c(2000, 2000), xend = c(2004, 2004),
           y = c(100,90), yend = c(100,90),
           color = c("#6002ee", "#ee6002"),
           cex = 2) +
  annotate("text", 
           x = c(2005, 2005), 
           y = c(100,90), 
           label = c("inconsistency", "decision inconsistency"),
           hjust = 0) +
  labs(x = "\nyear of publication",
       y = "proportion of inconsistencies per article in %\n") +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(breaks =  c(2000, 2005, 2010, 2015, 2020, 2023)) +
  theme_minimal() + 
  theme(legend.position = "Top right")

# plot prop over time per journal
year_journal_plot <- 
ggplot(year_journal_prop) + 
  geom_path(aes(x = year, 
                 y = prop_errors* 100),
             color = "#6002ee") +
  geom_path(aes(x = year, 
                 y = prop_gross* 100),
             color = "#ee6002") +
  facet_wrap(~ journal, ncol = 2) +
  labs(x = "\nyear of publication",
       y = "") +
  scale_y_continuous(limits = c(0, 100),
                     breaks = c(0,50,100)) +
  scale_x_continuous(breaks =  c(2000, 2023)) +
  theme_minimal() + 
  theme(legend.position = "Top right",
        panel.spacing.x = unit(1.5, "lines"))

year_p <- year_plot + year_journal_plot
year_p

```

```{r}
1 + 1
```

```{r}
1 + 1
```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2024-07-15 scripts/01_Analysis.R

@@ -1,0 +1,407 @@
+## title: Statistical reporting inconsistencies in Ling
+## author: Timo B. Roettger
+## contact: timo.b.roettger@gmail.com   
+## date created: 2024-07-11
+## date last edited:2024-07-11
+
+
+## packages ####################################################################                        
+ 
+
+# load in relevant libraries, install if not installed
+# if (!require("pacman")) install.packages("pacman")
+# pacman::p_load(rstudioapi, 
+#                tidyverse, 
+#                ggpubr, 
+#                broom, 
+#                statcheck,
+#                janitor,
+#                patchwork
+# )
+
+library(rstudioapi)
+library(tidyverse)
+library(ggpubr)
+library(broom)
+library(statcheck)
+library(janitor)
+library(patchwork)
+library(ggstream)
+
+
+## data ########################################################################                        
+
+
+## load in data from location of script (uncomment if you run script outside of manuscript.qmd)
+#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
+#setwd(current_working_dir)
+
+## list all journal units
+files <- list.files("../Journals", recursive = TRUE, pattern = "\\.pdf")
+
+## how many articles?
+length(files)
+# 5804
+
+## summary table
+files_wrangled <- as.data.frame(files) %>%
+  mutate(year = substring(files, 5, 8)) %>%
+  filter(year != 2024) %>%
+  mutate(journal = substring(files, 13, 15)) %>%
+  group_by(journal) %>%
+  summarise(n.art = length(unique(files)))
+
+## view
+files_wrangled
+
+## store for later
+n_articles = length(files)
+
+
+## statcheck ###################################################################                      
+
+
+## read pdfs in /Journals with statcheck, checks for one tailed tests
+## can only be run with access to original articles, bulk sharing is not prohibited by the journals
+#stat <- checkPDFdir("Journals", OneTailedTxt = TRUE)
+
+## write statcheck data to csv
+#write.csv(stat, "../data/statcheck_data.csv", row.names = FALSE)
+
+
+## data ########################################################################                        
+
+
+## load in data
+xdata <- read.csv("../data/statcheck_data.csv")
+
+## split strings into year and journal, drop NAs
+xdata <- xdata %>% 
[…]

### 2024-07-15 manuscript/manuscript.qmd
tided up

@@
 author:
-  - name: Timo B. Roettger
-    corresponding: true
-    orcid: 0000-0000-0000-0001
-    email: timo.roettger@gmail.com
+  - name: Dara Leonard Jenssen Etemady
+    corresponding: false
+    email: daraetemady@gmail.com
     affiliations:
@@
         country: Norway
-  - name: Dara Leonard Jenssen Etemady
-    corresponding: false
-    email: daraetemady@gmail.com
+  - name: Timo B. Roettger
+    corresponding: true
+    orcid: 0000-0000-0000-0001
+    email: timo.roettger@gmail.com
     affiliations:
@@
 # 1. Introduction
-
 What we know about human language and its cognitive underpinnings is often informed by experimental data, data based on which researchers draw theoretically relevant inferences using common statistical frameworks.

### 2024-07-15 manuscript/manuscript.tex
tided up

@@
 
-\authorsnames{Timo B. Roettger,Dara Leonard Jenssen Etemady}
+\authorsnames{Dara Leonard Jenssen Etemady,Timo B. Roettger}
 
@@
 
-\leftheader{Roettger and Etemady}
+\leftheader{Etemady and Roettger}
 
@@
 (c) the test result is reported in American Psychological Association
-style (APA, 2019). Given these constraints, statcheck is estimated to
-detect roughly 60\% of all reported NHST results
+style (\citeproc{ref-APA2020}{American Psychological Association,
+2020}). Given these constraints, statcheck is estimated to detect
+roughly 60\% of all reported NHST results
 (\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}) Statcheck

### 2024-07-18 manuscript/manuscript.qmd
edits and added conclusion

@@
 ---
-title: "Statistical Inconsistencies in Experimental Linguistics"
+title: "Statistical Reporting Inconsistencies in Experimental Linguistics"
 shorttitle: "Statistical Inconsistencies"
@@
     keep-tex: true
-
+header-includes: \usepackage{annotate-equations}
 execute:
@@
 # 1. Introduction
-What we know about human language and its cognitive underpinnings is often informed by experimental data, data based on which researchers draw theoretically relevant inferences using common statistical frameworks.
-This inference should be transparent in order for other researchers to critically evaluate the inference process and potentially detect and correct human errors.
-In the recent years, the quantitative sciences have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share], sharing of statistical protocols is still rare across the language sciences [@bochynska2023reproducible].
+What we know about human language and its cognitive underpinnings is often informed by experimental data.
+Researchers test theoretical predictions with these data using statistical tests.
+Depending on the results of such tests, researchers make claims for or against theoretical assumptions.\
+Since they play such an integral part in the reasoning process of experimentalists, both the data these tests are based on and the computational procedure of the tests itself should be transparent in order for other researchers to critically evaluate them.
+Moreover things can go wrong.
+To err is human.
+Transparent sharing allows others to detect and correct human error.
+In the recent years, the quantitative sciences have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
+Despite these calls, sharing of statistical protocols is still rather rare across the language sciences [@bochynska2023reproducible].
 If statistical procedures cannot be critically evaluated, human errors might be left undetected and thus remain uncorrected in the publication record.
-If these errors affect the decision procedure of the analysis, i.e.
+And if undetected errors affect the decision procedure of the analysis, i.e.
 whether a hypothesis is rejected or accepted, these errors might lead to -at best- overconfident, -at worst- false theoretical conclusions.
-The present paper will present evidence that the published literature in experimental linguistics contains a concerning amount of statistical errors, a state of affairs which warrants more rigorous data sharing practices.
+The present paper will present evidence that the published literature in experimental linguistics contains a concerning amount of such statistical errors, a state of affairs which warrants more rigorous data sharing practices.
 
@@
 The null-hypothesis significance testing (NHST) framework is, to date, the most dominant statistical framework that researchers use to test hypotheses in the language sciences [@sondereggera2024advancements].
-These statistical tests are reported in specific formats which usually contain a test statistic, the degrees of freedom of that test (if applicable), and the p-value, representing the probability of observing the data (or more extreme data) given the null hypothesis (i.e.
+NHST tests are commonly reported in specific formats which usually contain the name of the test (e.g.
+F, t, χ2), a test statistic, the degrees of freedom of that test (if applicable), and the p-value, representing the probability of observing the data (or more extreme data) given the null hypothesis (i.e.
 given that the test statistic is zero) (see example 1):
-\
-(1) F(1, 66) = 3.88, p \< .05
-\
+&nbsp;
+
+&nbsp;
+
+```{=tex}
+\begin{equation*}
+  \tikzmarknode{node1}{(1) }
+  \eqnmarkbox[gray]{node2}{F}
+  \eqnmarkbox[orange]{node3}{(1,66)}
+  \tikzmarknode{node4}{=}
+  \eqnmarkbox[purple]{node5}{3.88}
+  \tikzmarknode{node6}{, }
+  \eqnmarkbox[blue]{node7}{p < .05}
+\end{equation*}
+
+\annotate[yshift=1em]{left}{node2}{name of test} 
+\annotate[yshift=-1em]{below,left}{node3}{degrees of freedom} 
+\annotate[yshift=1em]{right}{node5}{test statistic} 
+\annotate[yshift=-1em]{below,right}{node7}{p-value}
+
+```
+&nbsp;
+
 Since data and statistical protocol sharing remains rare across the language sciences [@bochynska2023reproducible], interested readers are left with trusting the authors that the statistical analysis is run and reported correctly.
-Whether that trust is warranted can be assessed, however.
-The three sets of indices in (1) have a clearly defined mathematical relationship and can thus be checked for consistency.
+However, the three sets of indices in (1) have clearly defined mathematical relationships and can thus be easily checked for consistency.
 An F test with the specified degrees of freedom and a test-statistic of 3.88 should result in a p-value of 0.053 which is larger, not smaller, than 0.05.
-Thus, something is wrong with the statistical report in (1).
 Possible reasons for this inconsistency are manifold: It could be a typo of the comparison sign, i.e. the authors meant to use `=` or `>` rather than `<`.
-Alternatively, any of the numbers could be a typo.
+Any of the numbers could contain a typo.
 Sometimes an error might indicate erroneous rounding (e.g. 0.057 being rounded down to 0.05).
-Other times, human error along the data analytical pipeline might have caused an error down the line.
-Without access to data and scripts, it remains unclear to the reader, what has caused the inconsistency. 
-Such inconsistencies can be particularly concerning if the calculated p-value and the reported p-values are not on the same side of the alpha threshold.
-In NHST, p-values below a conventionalized alpha threshold, most commonly 0.05, are interpreted as evidence that the data are sufficiently inconsistent with the null hypothesis (significant).
-p-values above that threshold are considered consistent with the null hypothesis and practically lead to rejecting the alternative hypothesis (not significant).
+Without access to data and scripts, it remains unclear to the reader what has caused this inconsistency.
+Such inconsistencies can be particularly concerning if the calculated p-value (here 0.057) and the reported p-values (here >0.05) are not on the same side of the alpha threshold.
+In NHST, p-values below a conventionalized alpha threshold, most commonly 0.05, are interpreted as evidence that the data are sufficiently inconsistent with the null hypothesis ("significant").
+P-values above that threshold are considered consistent with the null hypothesis and practically lead to rejecting the alternative hypothesis ("non-significant").
 In (1) above, the reported p-value suggest a significant result, the p-value derived from the degrees of freedom and the test statistic suggest a non-significant result.
-In the following we refer to these inconsistencies as "decision inconsistency".
+In the following we refer to these inconsistencies as "decision inconsistencies".
 
-This form of inconsistency assessment can be automatically assessed if statistical tests are reported in an unambiguous format.
-Recently, a series of studies used such automatic assessments to evaluate the prevalence of inconsistent statistical reporting across different disciplines [e.g.
-@nuijten2016prevalence; @nuijten2020statcheck; @green2018statcheck; @gross2021fidelity; @buckley2023estimating].
-For example, looking at over 250000 p-values published in major psychology journals, @nuijten2016prevalence found that around 50% of the articles with statistical results contained at least one inconsistencies and around 12.5% contained at least one “decision inconsistency”.
+The consistency of these values can be automatically assessed if statistical tests are reported in an unambiguous format.
+Recently, a series of studies used such automatic assessments to evaluate the prevalence of inconsistent statistical reporting in psychology [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness; @nuijten2016prevalence; @nuijten2020statcheck; @green2018statcheck; @claesen2023data], medical sciences [@garcia2004incongruence; @van2023comparing], psychiatry [@berle2007inconsistencies], cyber security studies [@gross2021fidelity], technological education research [@buckley2023estimating], and experimental philosophy [@colombo2018statistical].
+For example, looking at over 250 thousand p-values published in major psychology journals, @nuijten2016prevalence found that around 50% of the articles with statistical results contained at least one inconsistency and around 13% contained at least one “decision inconsistency”.
+Other studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% decision consistencies.
 
-To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses over 39000 p-values reported in eight experimental linguistic journals published between 2000 and 2023.
-We further assess whether the rate of inconsistency differs across journals, whether that rate has changed over the course of the last 20 years and whether there is evidence for bias in these statistical-reporting inconsistencies.
+To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in eight experimental linguistic journals published between 2000 and 2023.
+We explore whether the rate of inconsistency differs across journals, whether that rate has changed over the course of the last 20 years and whether there is evidence for bias in these statistical-reporting inconsistencies.
 We discuss the results and offer concrete recommendations for authors, reviewers, and editors to tackle this problem.
@@
 # 3. Method
-All quantitative analyses were conducted using R [] and the r packages [LIST ALL PACKAGES from 01_Analysis.R and add to references]
+
+All quantitative analyses were conducted using @rmanual and the r packages \[LIST ALL PACKAGES from 01_Analysis.R and add to references\]
 
 ## 3.1. Statcheck
-We used the R package statcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
-Statcheck works as follows: After converting pdf or html articles to plain text, statcheck searches for specific strings that correspond to a NHST result, using “regular expressions”.
-That way, statcheck can detect results of t tests, F tests, Z tests, χ2 tests, correlation tests, and Q tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
-it usually misses results reported in tables; and (c) the test result is reported in American Psychological Association style [@APA2020].
-Given these constraints, statcheck is estimated to detect roughly 60% of all reported NHST results [@nuijten2016prevalence] Statcheck uses the reported test statistic and degrees of freedom to recalculate the p-value, compares the reported and recalculated p-values and, if there is a mismatch, reports a comparison as an “inconsistency.” The algorithm takes into account that tests might have been performed as one-tailed by identifying the search strings “one-tailed,” “one-sided,” or “directional”.
-Moreover, statcheck counts p = .000 and p \< .000 as inconsistent as p-values of exactly zero are mathematically impossible and the APA manual [@APA2020] advises to report very small p-values as p \< .001.
-Validity checks suggest that inter rater reliability between manual coding and statcheck is high, i.e. 0.76 for inconsistencies and 0.89 for decision inconsistencies [@nuijten2016prevalence].
-The overall accuracy of statcheck is estimated between 96.2% to 99.9% [@nuijten2017validity].
-We thus consider statcheck a valid, but rough proxy of statistical reporting inconsistencies.
+We used the R package wtatcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
+Statcheck works as follows: After converting articles in pdf or html format to plain text, Statcheck searches for specific strings that correspond to a NHST result using “regular expressions”.
+That way, statcheck can detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, the degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
+Statcheck usually misses reported in tables; and (c) the test result is reported in American Psychological Association style [@APA2020].
+Given these constraints, Statcheck is estimated to detect roughly 60% of all reported NHST results [@nuijten2016prevalence]. 
+Statcheck uses the reported test statistic and degrees of freedom to recalculate the p-value, compares the reported and recalculated p-values and, if there is a mismatch, flags the test as containing an “inconsistency.”
+The algorithm takes into account that tests might have been performed as one-tailed by identifying the search strings “one-tailed,” “one-sided,” or “directional” in the body of the text.
+Moreover, Statcheck considers p = .000 and p \< .000 as inconsistent because p-values of exactly zero are mathematically impossible and the APA manual [@APA2020] advises to report very small p-values as p \< .001.
+Validity checks of Statcheck suggest that inter-rater reliability between manual coding and Statcheck is high, i.e.
+0.76 for inconsistencies and 0.89 for decision inconsistencies [@nuijten2016prevalence].
+The overall accuracy of Statcheck is estimated to be between 96.2% to 99.9% [@nuijten2017validity].
+We thus consider Statcheck a valid proxy of the prevalence of statistical reporting inconsistencies.
 
 ## 3.2. Sample
-Focusing on experimental linguistic research, we started with @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the term “experiment* in title, abstract or keywords.
+Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the term “experiment\* in title, abstract and/or keywords.
 Out of these 100 journals, we selected all journals with at least 10% of articles containing the search string "experiment\*".
 Out of the remaining 37 journals, we selected those journals that urged APA formatting either in the main body of the text or specifically regarding statistics in the author guidelines, resulting in nine remaining journals.
-Moreover, to download the pdfs, the articles had to be either accessible to us through our library license, or open access, resulting in a final list of eight journals: Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA).
+Moreover, to access the articles in .pdf format, the articles had to be either accessible to us through our library license, or open access, resulting in a final list of eight journals: Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA).
 
 We included only original research articles within the publication years of 2000-2023, excluding any book reviews, response articles, commentaries, editorials, corrigenda, errata, advertisements, etc.
-Articles from LAB spanned 2011-2023, while the rest spanned 2000-2023.
-This procedure resulted in `{r} length(files)` research articles.
-All `{r} length(files)` articles were submitted to analysis, using Statcheck v1.5.0.
-157 articles were not analyzable by statcheck, and thus had to be removed from the pool, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt.
+Articles from LAB spanned 2011-2023, while the other journals spanned 2000-2023.
+Statcheck could not parse 157 articles, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt.
+This procedure resulted in `{r} length(files)` research articles which were submitted to analysis.
 
 ## 3.3. Data availability statement
-
 All derived data and corresponding R scripts are publicly available here: LINK.
-The original journal articles cannot be bulk-shared due to distribution restrictions by the publishers.
-NEED TO FIGURE OUT A WAY
 
@@
 ## 4.1 Prevalence of inconsistencies
+
 The results are summarized in @tbl-summary.
 Out of `{r} journal_summary_counts[9,2][[1]]` articles, `{r} journal_summary_counts[9,3][[1]]` articles contained statistical tests that statcheck could assess (`{r} 100 * round(journal_summary_counts[9,3][[1]] / journal_summary_counts[9,2][[1]],2)`%), amounting to `{r} journal_summary_counts[9,4][[1]]` assessable p-values.
-`{r} journal_summary_counts[9,5][[1]]` p-values were flagged as inconsistent (`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`%) and `{r} journal_summary_counts[9,6][[1]]` of which were considered decisions inconsistencies (`{r} 100* round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`%) (see @tbl-summary).
+`{r} journal_summary_counts[9,5][[1]]` p-values were flagged as inconsistent (`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`%) out of which `{r} journal_summary_counts[9,6][[1]]` were considered decision inconsistencies (`{r} 100* round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`%) (see @tbl-summary).
 
@@
 #| label: tbl-summary
-#| tbl-cap: "Summary.
-Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA)."
+#| tbl-cap: "Number of eligible articles, assessable articles and results, inconsistencies and decision inconsistencies across all journals.
+(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA))"
 
@@
 #| fig-cap: "Proportion of articles containing at least one inconsistency / decision inconsistency.
-Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA)."
+(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA))"
 
@@
 #| out-width: 100%
-#| fig-cap: "Proportion of inconsistencies / decision inconsistencies from 2000 to 2023.
-Left: global rates with all journals pooled; Right: rates per journal.
-Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA)."
+#| fig-cap: "Proportion of inconsistencies / decision inconsistencies plotted from 2000 to 2023.
+Left: rates pooled across journals; Right: rates per journal.
+(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA))"
 
@@
 
-When examining reported against recalculated p-values (see @fig-scatter) for all inconsistencies, we can identify certain spatial patterns (only inconsistencies that are based on '=' comparisons were included to make them easier to interpret, n =
-`{r} nrow(xdata[xdata$error == TRUE & xdata$test_comp == "=",])`
-out of 
-`{r} nrow(xdata[xdata$error == TRUE,])`).
-First, the center of density of points is located in the bottom left corner with more values reported closer to the alpha level.
-This is not surprising, since there is a documented publication bias in published quantitative articles, with hypotheses being much more often confirmed (and thus p < .05) than not [@franco2014publication; @sterling1959publication].
+When examining reported against recalculated p-values for all inconsistencies (see @fig-scatter), we can identify certain spatial patterns (only inconsistencies that are based on '=' comparisons were included to make them easier to interpret, n = `{r} nrow(xdata[xdata$error == TRUE & xdata$p_comp == "=",])` out of `{r} nrow(xdata[xdata$error == TRUE,])`).
+First, the center of density of points is located in the bottom left corner with more p-values reported closer to the standard alpha level of 0.05.
+This is not surprising, since the quantitative sciences exhibit a documented publication bias, with hypotheses being more often confirmed (and thus p \< .05) than not [@franco2014publication; @sterling1959publication].
 Second, there is a linear density of points along the diagonal axis which corresponds to numerically small inconsistencies, some of which might be related to simple rounding errors.
@@
 recalculated p-values were as likely to be smaller than the reported p-value as larger).
-The slope of the regression line is flatter than the diagonal axis which means that, on average, reported p-values are lower than their recalculated counter parts. 
-In other words, inconsistencies have a tendency to lead to smaller p-values.
-This potential bias can also be observed in decision inconsistencies. 
-Of all decisions inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant value (p \> 0.05).
+The slope of the regression line is flatter than the diagonal axis which means that, on average, inconsistencies have a tendency to exhibit smaller reported p-values than their recalculated counterparts.
+Similarly, inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p > 0.05 or p < 0.05 ) are not equally prevalent. 
+There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference (e.g.
+p < 0.05).
+So even if we assumed these inconsistencies were merely typos of the comparison sign, inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.
+These biases are also reflected in decisions inconsistencies.
+Of all decisions inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant (p \> 0.05).
 
@@
 
-## 4.2 Manual inspection of decision inconsistencies
-To further understand the nature of decision inconsistencies, we manually inspected all of them (n=655). 
-We (a) evaluated whether statcheck has extracted the information correctly, and (b) assessed whether the text actually suggest an erroneous inferential decision, i.e.
-whether the paper uses a reported significant result to claim a significant effect or a reported non-significant result to claim a null result.
-
 # 5. Discussion and Recommendations
@@
 ## 5.1. Statistical reporting inconsistencies are prevalent
-The present study found a concerning amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[9,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[9,3][[1]]` assessable p-values.
+The present study found a large amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[9,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[9,3][[1]]` assessable p-values.
 `{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`% of all p-values were flagged as inconsistent and `{r} 100 * round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`% were flagged as decision inconsistencies, i.e.
@@
 
-The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency in psychology [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness; @nuijten2016prevalence; @nuijten2020statcheck; @green2018statcheck; @claesen2023data], medical sciences [@garcia2004incongruence,; @van2023comparing], psychiatry [@berle2007inconsistencies], cyber security studies[@gross2021fidelity], technological education research [@buckley2023estimating], and experimental philosophy [@colombo2018statistical].
-These studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% decision consistencies.
+The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent such assessment using Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
+Previous studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% contain at least one decision consistencies.
 
-Even if the prevalence of these inconsistencies could be largely attributed to inconsequential typos or rounding errors (an assumption we cannot test without access to the data), the sheer amount of these errors that have made it through peer-review should concern us.
+Even if the prevalence of these inconsistencies could be largely attributed to inconsequential typos or rounding errors (an assumption we cannot test without access to the data), the sheer amount of these inconsistencies that have made it through peer-review should concern us.
 They are human errors.
-If such a substantial amount of errors is found in plain site, the question naturally arises how many errors that happen during the data analysis itself remain undetected. 
-If the tip of the iceberg above water is so large, how large is the iceberg underneath the surface?
+If such a substantial amount of errors is found in plain site, the question naturally arises as to how many errors during the data analysis itself remain undetected.
+If the tip of the iceberg above water is already so large, how large is the body of the iceberg underneath the surface?
 
-The present examination also looked at whether the rates of inconsistency changed over the last 23 years of publications. 
-Descriptive assessment did not indicate any noticeable trend.
-Moreover, the rates of both inconsistencies and decision inconsistencies seem to remain stable across journals.
+The present examination assessment did not indicate any noticeable trend over the last 23 years of publications nor did the data suggest large differences between journals.
 
 Observed inconsistencies were characterized by reported p-values being on average lower than their recalculated counterparts and the prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant.
-This could indicate a systematic bias in favor of lower p-values in general and a bias towards significant results in particular. 
-Our data do not speak to the causes of these biases, but possible reasons include the following:
-Researchers might intentionally round down p-values because they think a lower p-value is more convincing to reviewers and/or readers.
+These patterns could indicate a systematic bias in favor of lower p-values in general and a bias towards significant results in particular.
+Our data do not speak to the causes of these biases, but possible reasons include the following: Researchers might intentionally round down p-values because they think a lower p-value is more convincing to reviewers and/or readers.
 This practice has been admitted to by 1 in 5 surveyed psychological researchers [@john2012measuring].
-Given that quantitative linguists have been reported to commit questionable research practices (and even fraud) more often than we would like to [@isbell2022misconduct], we cannot exclude the possibility that some of the inconsistencies were intentional.
+Given that a non-trivial amount of quantitative linguists have admitted to commit questionable research practices (and even fraud) [@isbell2022misconduct], we cannot exclude the possibility that some of the inconsistencies were intentional.
 It is our strong belief, however, that the majority of inconsistencies are unintentional.
 
-Researchers might scrutinize non-significant results more than significant results or fail to double check significant results more often because results that confirm their hypothesis feed into their confirmation bias [@nickerson1998confirmation].
+Researchers might scrutinize non-significant results more than significant results or are less likely to double check significant results than non-significant results because results that confirm their hypothesis feed into their confirmation bias [@nickerson1998confirmation].
 For example, @fugelsang2004theory let researchers evaluate data that are either consistent or inconsistent with their prior expectations.
@@
 ## 5.2. Limitations of our study
+
 While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, there are, of course, a number of limitations.
@@
 Thus it is possible that a different selection of journals would have resulted in different results.
-However, given that the inconsistency rates of our study are comparable to similar studies from other disciplines and that the inconsistency rates are both comparable across the eight journals and stable across the most recent 23 years, suggests that our findings are relevant for experimental linguistics at large.
+However, given that the inconsistency rates of our study are comparable to similar studies from other disciplines and that those rates are relatively stable across journals and time, our findings should be considered relevant for experimental linguistics at large.
 
-Second, given the constraints on automatically detecting test statistics, statcheck misses reported values that either diverge from APA reporting standards or are reported in tables.
+Second, given the constraints on automatically detecting test statistics, Statcheck misses reported values that either diverge from APA reporting standards or are reported in tables.
 However, inconsistency rates have been shown to be similar for results in APA format vs. results that diverge from APA formatting [@bakker2011mis; @nuijten2016prevalence].
+
 Third, statcheck slightly overestimates inconsistency rates, because it might not accurately detect corrections for multiple comparisons [@schmidt2017statcheck].
 @nuijten2017validity, however, show that not only were there only a small proportion of flagged inconsistencies related to multiple comparisons, but also that these multiple comparisons themselves were often erroneously reported.
-They conclude that "[a]ny reporting inconsistencies associated with these tests and corrections could not explain the high prevalence of reporting inconsistencies" [@nuijten2017validity, p. 27].
-More elaborate automatic tools for the extraction of statistical information might allow a more detailed and more accurate assessment of statistical reporting in the future [e.g. @kalmbach2023rule].
-But even statcheck is limited and only provides a rough proxy of true inconsistency rate in the published literature, we hope the reader agrees that it is a state of affairs that should be tackled.
+They conclude that "\[a\]ny reporting inconsistencies associated with these tests and corrections could not explain the high prevalence of reporting inconsistencies" [@nuijten2017validity, p. 27].
+
+More elaborate automatic tools for the extraction of statistical information might allow for a more detailed and more accurate assessment of statistical reporting in the future [e.g.
+@kalmbach2023rule].
+Despite its limitations, Statcheck provides a rough proxy of true inconsistency rate in the published literature and we hope the reader agrees that the prevalence of inconsistencies is a state of affairs that should be reflected upon.
 
 ## 5.3. Recommendations for the field
+
 There are concrete actionable steps the field of experimental linguistics can make to reduce statistical reporting inconsistencies.
-In order to avoid simple copy and paste errors related to working in two separate programs for writing the manuscript and conducting the statistical analysis, authors should consider 'literate programming', i.e.
+In order to avoid simple copy-and-paste errors related to working in two separate programs for writing the manuscript and conducting the statistical analysis, authors should consider 'literate programming', i.e.
 an integration of analysis code and prose into a single, dynamic document [@knuth1984literate; @casillas2023opening].
 Several implementations of literate programming are freely available to researchers including common R markdown files (Rmd) and Quarto markdown files (qmd).
-Literate programming can ensure that values derived by the statistical analysis is automatically integrated into the manuscript document, avoiding errors that might happen during a transfer from one program to the other.
+Literate programming can ensure that values derived from the statistical analysis are automatically integrated into the manuscript document, avoiding errors that might happen during a manual transfer from one program to the other.
 
@@
 the anonymized data table that was analyzed) as well as a detailed description of their statistical protocol, ideally in form of reproducible scripts.
-Sharing reproducible analyses with reviewers allows the reviewers the reproduce the authors' analyses, possibly detect errors or even inappropriate statistical choices before publication, thus improving the quality of their work.
-Moreover, publicly sharing their analyses has numerous benefits to the authors themselves beyond error detection: 
-Open data and materials can facilitate collaboration [@boland2017ten], increase efficiency and sustainability [@lowndes2017our], and are cited more often [@colavizza2020citation].
+Sharing reproducible analyses with reviewers allows the reviewers the reproduce the authors' analyses, possibly detect errors or even inappropriate statistical choices before publication, thus improving the quality and robustness of the final product.
+Moreover, publicly sharing their analyses has numerous benefits to the authors themselves beyond error detection: Open data and materials can facilitate collaboration [@boland2017ten], increase efficiency and sustainability [@lowndes2017our], and are cited more often [@colavizza2020citation].
 
 If authors do not share data and scripts, reviewers can check at least the statistical reporting consistency in the manuscript by using tools such as statcheck [@nuijten2023statcheck, http://statcheck.io] or p-checker [@schonbrodt2015p, http://shinyapps.org/apps/p-checker/].
-Reviewers could consider consider requesting data and scripts during peer review.
+Reviewers could consider requesting data and scripts during peer review.
 Such requests might be particularly justified when inconsistencies are apparent.
@@
 Recent evidence suggest experimental linguistics are still characterized by a pluralism of statistical approaches, even when trying to answer the same research question [@coretta_multidimensional_2023].
-Some of these approaches are might be more appropriate than others [@sondereggera2024advancements; @vasishth2023some], so more thorough evaluations of how researchers arrive at their statistical conclusions might elevate their analytical robustness.
+Some of these approaches might be more appropriate than others [@sondereggera2024advancements; @vasishth2023some], so more thorough evaluations of how researchers arrive at their statistical conclusions might elevate their analytical robustness.
 
-Journal editors could explicitly recommend consistency checks with apps such as statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines [Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous].
+Journal editors could explicitly recommend consistency checks with algorhythms such as statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines \[Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous\].
 Editors could also demand, recommend or at least encourage data sharing for publication in their journal. 
@@
 [^1]: http://www.psychologicalscience.org/publications/psychological_science/ps-submissions; accessed on July 15, 2024.
+
 [^2]: https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines; accessed on on July 15, 2024.
+
+Researchers make errors. 
+Researchers have biases.
+This is who we are as humans and there is not much we can do about our nature. 
+Being aware of this fact and how it might affect research might help us to make possibly negative consequences at worst detectable and best preventable.
 

### 2024-07-18 manuscript/manuscript.tex
edits and added conclusion

@@
 
-\title{Statistical Inconsistencies in Experimental Linguistics}
+\title{Statistical Reporting Inconsistencies in Experimental
+Linguistics}
 
@@
 
+\usepackage{annotate-equations}
 \makeatletter
@@
 What we know about human language and its cognitive underpinnings is
-often informed by experimental data, data based on which researchers
-draw theoretically relevant inferences using common statistical
-frameworks. This inference should be transparent in order for other
-researchers to critically evaluate the inference process and potentially
-detect and correct human errors. In the recent years, the quantitative
+often informed by experimental data. Researchers test theoretical
+predictions with these data using statistical tests. Depending on the
+results of such tests, researchers make claims for or against
+theoretical assumptions.\\
+Since they play such an integral part in the reasoning process of
+experimentalists, both the data these tests are based on and the
+computational procedure of the tests itself should be transparent in
+order for other researchers to critically evaluate them. Moreover things
+can go wrong. To err is human. Transparent sharing allows others to
+detect and correct human error. In the recent years, the quantitative
 sciences have seen repeated calls to become more transparent and
@@
 \citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
-\citeproc{ref-roettger2019researcher}{Roettger, 2019}), sharing of
-statistical protocols is still rare across the language sciences
-(\citeproc{ref-bochynska2023reproducible}{Bochynska et al., 2023}). If
-statistical procedures cannot be critically evaluated, human errors
-might be left undetected and thus remain uncorrected in the publication
-record. If these errors affect the decision procedure of the analysis,
-i.e.~whether a hypothesis is rejected or accepted, these errors might
-lead to -at best- overconfident, -at worst- false theoretical
-conclusions. The present paper will present evidence that the published
-literature in experimental linguistics contains a concerning amount of
-statistical errors, a state of affairs which warrants more rigorous data
-sharing practices.
+\citeproc{ref-roettger2019researcher}{Roettger, 2019}). Despite these
+calls, sharing of statistical protocols is still rather rare across the
+language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
+al., 2023}). If statistical procedures cannot be critically evaluated,
+human errors might be left undetected and thus remain uncorrected in the
+publication record. And if undetected errors affect the decision
+procedure of the analysis, i.e.~whether a hypothesis is rejected or
+accepted, these errors might lead to -at best- overconfident, -at worst-
+false theoretical conclusions. The present paper will present evidence
+that the published literature in experimental linguistics contains a
+concerning amount of such statistical errors, a state of affairs which
+warrants more rigorous data sharing practices.
 
@@
 (\citeproc{ref-sondereggera2024advancements}{Sonderegger \& Sóskuthy,
-2024}). These statistical tests are reported in specific formats which
-usually contain a test statistic, the degrees of freedom of that test
-(if applicable), and the p-value, representing the probability of
-observing the data (or more extreme data) given the null hypothesis
-(i.e.~given that the test statistic is zero) (see example 1):\\
-(1) F(1, 66) = 3.88, p \textless{} .05\\
+2024}). NHST tests are commonly reported in specific formats which
+usually contain the name of the test (e.g.~F, t, χ2), a test statistic,
+the degrees of freedom of that test (if applicable), and the p-value,
+representing the probability of observing the data (or more extreme
+data) given the null hypothesis (i.e.~given that the test statistic is
+zero) (see example 1): ~
+
+~
+
+\begin{equation*}
+  \tikzmarknode{node1}{(1) }
+  \eqnmarkbox[gray]{node2}{F}
+  \eqnmarkbox[orange]{node3}{(1,66)}
+  \tikzmarknode{node4}{=}
+  \eqnmarkbox[purple]{node5}{3.88}
+  \tikzmarknode{node6}{, }
+  \eqnmarkbox[blue]{node7}{p < .05}
+\end{equation*}
+
+\annotate[yshift=1em]{left}{node2}{name of test} 
+\annotate[yshift=-1em]{below,left}{node3}{degrees of freedom} 
+\annotate[yshift=1em]{right}{node5}{test statistic} 
+\annotate[yshift=-1em]{below,right}{node7}{p-value}
+
+~
+
 Since data and statistical protocol sharing remains rare across the
@@
 al., 2023}), interested readers are left with trusting the authors that
-the statistical analysis is run and reported correctly. Whether that
-trust is warranted can be assessed, however. The three sets of indices
-in (1) have a clearly defined mathematical relationship and can thus be
-checked for consistency. An F test with the specified degrees of freedom
-and a test-statistic of 3.88 should result in a p-value of 0.053 which
-is larger, not smaller, than 0.05. Thus, something is wrong with the
-statistical report in (1). Possible reasons for this inconsistency are
-manifold: It could be a typo of the comparison sign, i.e.~the authors
-meant to use \texttt{=} or \texttt{\textgreater{}} rather than
-\texttt{\textless{}}. Alternatively, any of the numbers could be a typo.
-Sometimes an error might indicate erroneous rounding (e.g.~0.057 being
-rounded down to 0.05). Other times, human error along the data
-analytical pipeline might have caused an error down the line. Without
-access to data and scripts, it remains unclear to the reader, what has
-caused the inconsistency. Such inconsistencies can be particularly
-concerning if the calculated p-value and the reported p-values are not
-on the same side of the alpha threshold. In NHST, p-values below a
-conventionalized alpha threshold, most commonly 0.05, are interpreted as
-evidence that the data are sufficiently inconsistent with the null
-hypothesis (significant). p-values above that threshold are considered
-consistent with the null hypothesis and practically lead to rejecting
-the alternative hypothesis (not significant). In (1) above, the reported
-p-value suggest a significant result, the p-value derived from the
-degrees of freedom and the test statistic suggest a non-significant
-result. In the following we refer to these inconsistencies as ``decision
-inconsistency''.
-
-This form of inconsistency assessment can be automatically assessed if
+the statistical analysis is run and reported correctly. However, the
+three sets of indices in (1) have clearly defined mathematical
+relationships and can thus be easily checked for consistency. An F test
+with the specified degrees of freedom and a test-statistic of 3.88
+should result in a p-value of 0.053 which is larger, not smaller, than
+0.05. Possible reasons for this inconsistency are manifold: It could be
+a typo of the comparison sign, i.e.~the authors meant to use \texttt{=}
+or \texttt{\textgreater{}} rather than \texttt{\textless{}}. Any of the
+numbers could contain a typo. Sometimes an error might indicate
+erroneous rounding (e.g.~0.057 being rounded down to 0.05). Without
+access to data and scripts, it remains unclear to the reader what has
+caused this inconsistency. Such inconsistencies can be particularly
+concerning if the calculated p-value (here 0.057) and the reported
+p-values (here \textgreater0.05) are not on the same side of the alpha
+threshold. In NHST, p-values below a conventionalized alpha threshold,
+most commonly 0.05, are interpreted as evidence that the data are
+sufficiently inconsistent with the null hypothesis (``significant'').
+P-values above that threshold are considered consistent with the null
+hypothesis and practically lead to rejecting the alternative hypothesis
+(``non-significant''). In (1) above, the reported p-value suggest a
+significant result, the p-value derived from the degrees of freedom and
+the test statistic suggest a non-significant result. In the following we
+refer to these inconsistencies as ``decision inconsistencies''.
+
+The consistency of these values can be automatically assessed if
 statistical tests are reported in an unambiguous format. Recently, a
 series of studies used such automatic assessments to evaluate the
-prevalence of inconsistent statistical reporting across different
-disciplines (\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
+prevalence of inconsistent statistical reporting in psychology
+(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011},
+\citeproc{ref-bakker2014outlier}{2014};
+\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
+\citeproc{ref-claesen2023data}{Claesen et al., 2023};
 \citeproc{ref-green2018statcheck}{Green et al., 2018};
-\citeproc{ref-gross2021fidelity}{Groß, 2021}; e.g.
 \citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016};
-\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020}). For
-example, looking at over 250000 p-values published in major psychology
-journals, Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016})
-found that around 50\% of the articles with statistical results
-contained at least one inconsistencies and around 12.5\% contained at
-least one ``decision inconsistency''.
+\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020};
+\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
+\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}), medical
+sciences (\citeproc{ref-garcia2004incongruence}{Garcı́a-Berthou \&
+Alcaraz, 2004}; \citeproc{ref-van2023comparing}{Van Aert et al., 2023}),
+psychiatry (\citeproc{ref-berle2007inconsistencies}{Berle \& Starcevic,
+2007}), cyber security studies (\citeproc{ref-gross2021fidelity}{Groß,
+2021}), technological education research
+(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023}), and
+experimental philosophy (\citeproc{ref-colombo2018statistical}{Colombo
+et al., 2018}). For example, looking at over 250 thousand p-values
+published in major psychology journals, Nuijten et al.
+(\citeproc{ref-nuijten2016prevalence}{2016}) found that around 50\% of
+the articles with statistical results contained at least one
+inconsistency and around 13\% contained at least one ``decision
+inconsistency''. Other studies report on inconsistency rates between 4\%
+and 14\%, with between 10\% and 63\% of articles containing at least one
+inconsistency and between 3\% and 21\% decision consistencies.
 
@@
 Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
-over 39000 p-values reported in eight experimental linguistic journals
-published between 2000 and 2023. We further assess whether the rate of
-inconsistency differs across journals, whether that rate has changed
-over the course of the last 20 years and whether there is evidence for
-bias in these statistical-reporting inconsistencies. We discuss the
-results and offer concrete recommendations for authors, reviewers, and
-editors to tackle this problem.
+p-values reported in eight experimental linguistic journals published
+between 2000 and 2023. We explore whether the rate of inconsistency
+differs across journals, whether that rate has changed over the course
+of the last 20 years and whether there is evidence for bias in these
+statistical-reporting inconsistencies. We discuss the results and offer
+concrete recommendations for authors, reviewers, and editors to tackle
+this problem.
 
@@
 
-All quantitative analyses were conducted using R {[}{]} and the r
-packages {[}LIST ALL PACKAGES from 01\_Analysis.R and add to
-references{]}
+All quantitative analyses were conducted using R Core Team
+(\citeproc{ref-rmanual}{2021}) and the r packages {[}LIST ALL PACKAGES
+from 01\_Analysis.R and add to references{]}
 
@@
 
-We used the R package statcheck (Version 1.4.1-beta.2,
+We used the R package wtatcheck (Version 1.4.1-beta.2,
 \citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
 automatically detect statistical-reporting inconsistencies. Statcheck
-works as follows: After converting pdf or html articles to plain text,
-statcheck searches for specific strings that correspond to a NHST
-result, using ``regular expressions''. That way, statcheck can detect
-results of t tests, F tests, Z tests, χ2 tests, correlation tests, and Q
-tests as long as the test result fulfills three conditions: (a) the test
-result is reported completely including the test statistic, degrees of
-freedom (if applicable), and the p-value; (b) the test result is in the
-body of the text, i.e.~it usually misses results reported in tables; and
-(c) the test result is reported in American Psychological Association
-style (\citeproc{ref-APA2020}{American Psychological Association,
-2020}). Given these constraints, statcheck is estimated to detect
-roughly 60\% of all reported NHST results
-(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}) Statcheck
+works as follows: After converting articles in pdf or html format to
+plain text, Statcheck searches for specific strings that correspond to a
+NHST result using ``regular expressions''. That way, statcheck can
+detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation
+tests, and Q tests as long as the test result fulfills three conditions:
+(a) the test result is reported completely including the test statistic,
+the degrees of freedom (if applicable), and the p-value; (b) the test
+result is in the body of the text, i.e.~Statcheck usually misses
+reported in tables; and (c) the test result is reported in American
+Psychological Association style (\citeproc{ref-APA2020}{American
+Psychological Association, 2020}). Given these constraints, Statcheck is
+estimated to detect roughly 60\% of all reported NHST results
+(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
 uses the reported test statistic and degrees of freedom to recalculate
 the p-value, compares the reported and recalculated p-values and, if
-there is a mismatch, reports a comparison as an ``inconsistency.'' The
-algorithm takes into account that tests might have been performed as
+there is a mismatch, flags the test as containing an ``inconsistency.''
+The algorithm takes into account that tests might have been performed as
 one-tailed by identifying the search strings ``one-tailed,''
-``one-sided,'' or ``directional''. Moreover, statcheck counts p = .000
-and p \textless{} .000 as inconsistent as p-values of exactly zero are
-mathematically impossible and the APA manual
-(\citeproc{ref-APA2020}{American Psychological Association, 2020})
-advises to report very small p-values as p \textless{} .001. Validity
-checks suggest that inter rater reliability between manual coding and
-statcheck is high, i.e.~0.76 for inconsistencies and 0.89 for decision
-inconsistencies (\citeproc{ref-nuijten2016prevalence}{Nuijten et al.,
-2016}). The overall accuracy of statcheck is estimated between 96.2\% to
+``one-sided,'' or ``directional'' in the body of the text. Moreover,
+Statcheck considers p = .000 and p \textless{} .000 as inconsistent
+because p-values of exactly zero are mathematically impossible and the
+APA manual (\citeproc{ref-APA2020}{American Psychological Association,
+2020}) advises to report very small p-values as p \textless{} .001.
+Validity checks of Statcheck suggest that inter-rater reliability
+between manual coding and Statcheck is high, i.e.~0.76 for
+inconsistencies and 0.89 for decision inconsistencies
+(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
+overall accuracy of Statcheck is estimated to be between 96.2\% to
 99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
-thus consider statcheck a valid, but rough proxy of statistical
+thus consider Statcheck a valid proxy of the prevalence of statistical
 reporting inconsistencies.
@@
 
-Focusing on experimental linguistic research, we started with Kobrock
-and Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
+Focusing on experimental linguistic research, we used Kobrock and
+Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
 departure who list 100 linguistic journals that had at least a hundred
 articles published at the time of assessment (2021) and a high ratio of
-articles containing the term ``experiment* in title, abstract or
+articles containing the term ``experiment* in title, abstract and/or
 keywords. Out of these 100 journals, we selected all journals with at
@@
 statistics in the author guidelines, resulting in nine remaining
-journals. Moreover, to download the pdfs, the articles had to be either
-accessible to us through our library license, or open access, resulting
-in a final list of eight journals: Applied Psycholinguistics (APS),
-Bilingualism: Language and Cognition (BLC), Linguistic Approaches to
-Bilingualism (LAB), Language and Speech (LaS), Language Learning and
-Techology (LLT), Journal of Language and Social Psychology (LSP),
-Journal of Child Language (JCL), and Studies in Second Language
-Acquisition (SLA).
+journals. Moreover, to access the articles in .pdf format, the articles
+had to be either accessible to us through our library license, or open
+access, resulting in a final list of eight journals: Applied
+Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
+Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
+Language Learning and Techology (LLT), Journal of Language and Social
+Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
+Language Acquisition (SLA).
 
@@
 commentaries, editorials, corrigenda, errata, advertisements, etc.
-Articles from LAB spanned 2011-2023, while the rest spanned 2000-2023.
-This procedure resulted in 5804 research articles. All 5804 articles
-were submitted to analysis, using Statcheck v1.5.0. 157 articles were
-not analyzable by statcheck, and thus had to be removed from the pool,
-likely related to issues with rendering the Chi-Squared symbol being
-erroneously converted from .pdf to .txt.
+Articles from LAB spanned 2011-2023, while the other journals spanned
+2000-2023. Statcheck could not parse 157 articles, likely related to
+issues with rendering the Chi-Squared symbol being erroneously converted
+from .pdf to .txt. This procedure resulted in 5804 research articles
+which were submitted to analysis.
 
@@
 All derived data and corresponding R scripts are publicly available
-here: LINK. The original journal articles cannot be bulk-shared due to
-distribution restrictions by the publishers. NEED TO FIGURE OUT A WAY
+here: LINK.
 
@@
 assess (53\%), amounting to 39532 assessable p-values. 5166 p-values
-were flagged as inconsistent (13.1\%) and 655 of which were considered
-decisions inconsistencies (1.7\%) (see Table~\ref{tbl-summary}).
+were flagged as inconsistent (13.1\%) out of which 655 were considered
+decision inconsistencies (1.7\%) (see Table~\ref{tbl-summary}).
 
@@
 
-{\caption{{Summary. Applied Psycholinguistics (APS), Bilingualism:
-Language and Cognition (BLC), Linguistic Approaches to Bilingualism
-(LAB), Language and Speech (LaS), Language Learning and Techology (LLT),
-Journal of Language and Social Psychology (LSP), Journal of Child
-Language (JCL), and Studies in Second Language Acquisition
-(SLA).}{\label{tbl-summary}}}}
+{\caption{{Number of eligible articles, assessable articles and results,
+inconsistencies and decision inconsistencies across all journals.
+(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition
+(BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech
+(LaS), Language Learning and Techology (LLT), Journal of Language and
+Social Psychology (LSP), Journal of Child Language (JCL), and Studies in
+Second Language Acquisition (SLA))}{\label{tbl-summary}}}}
 
@@
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-assessible articles
+assessable articles
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
@@
 \caption{\label{fig-stacked-bar}Proportion of articles containing at
-least one inconsistency / decision inconsistency. Applied
+least one inconsistency / decision inconsistency. (Applied
 Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
@@
 Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
-Language Acquisition (SLA).}
+Language Acquisition (SLA))}
 
@@
 \caption{\label{fig-year-prop}Proportion of inconsistencies / decision
-inconsistencies from 2000 to 2023. Left: global rates with all journals
-pooled; Right: rates per journal. Applied Psycholinguistics (APS),
+inconsistencies plotted from 2000 to 2023. Left: rates pooled across
+journals; Right: rates per journal. (Applied Psycholinguistics (APS),
 Bilingualism: Language and Cognition (BLC), Linguistic Approaches to
@@
 Journal of Child Language (JCL), and Studies in Second Language
-Acquisition (SLA).}
+Acquisition (SLA))}
 
@@
 
-When examining reported against recalculated p-values (see
-Figure~\ref{fig-scatter}) for all inconsistencies, we can identify
-certain spatial patterns (only inconsistencies that are based on `='
-comparisons were included to make them easier to interpret, n = 5154 out
-of 5166). First, the center of density of points is located in the
-bottom left corner with more values reported closer to the alpha level.
-This is not surprising, since there is a documented publication bias in
-published quantitative articles, with hypotheses being much more often
-confirmed (and thus p \textless{} .05) than not
+When examining reported against recalculated p-values for all
+inconsistencies (see Figure~\ref{fig-scatter}), we can identify certain
+spatial patterns (only inconsistencies that are based on `=' comparisons
+were included to make them easier to interpret, n = 3779 out of 5166).
+First, the center of density of points is located in the bottom left
+corner with more p-values reported closer to the standard alpha level of
+0.05. This is not surprising, since the quantitative sciences exhibit a
+documented publication bias, with hypotheses being more often confirmed
+(and thus p \textless{} .05) than not
 (\citeproc{ref-franco2014publication}{Franco et al., 2014};
@@
 reported p-value as larger). The slope of the regression line is flatter
-than the diagonal axis which means that, on average, reported p-values
-are lower than their recalculated counter parts. In other words,
-inconsistencies have a tendency to lead to smaller p-values. This
-potential bias can also be observed in decision inconsistencies. Of all
-decisions inconsistencies (n = 655), 71\% represent cases in which a
-reported significant result (p \textless{} 0.05) is recalculated as
-non-significant value (p \textgreater{} 0.05).
+than the diagonal axis which means that, on average, inconsistencies
+have a tendency to exhibit smaller reported p-values than their
+recalculated counterparts. Similarly, inconsistencies that report the
+p-value as being larger or smaller than a reference value (e.g.~p
+\textgreater{} 0.05 or p \textless{} 0.05 ) are not equally prevalent.
+There were 4.4\% of inconsistencies with p being reported as larger than
+a reference but 7.4\% inconsistencies with p being reported as smaller
+than a reference (e.g.~p \textless{} 0.05). So even if we assumed these
+inconsistencies were merely typos of the comparison sign,
+inconsistencies that erroneously report the p-value to be smaller than a
+reference value are more frequent than inconsistencies that erroneously
+report the p-value to be larger than a reference value. These biases are
+also reflected in decisions inconsistencies. Of all decisions
+inconsistencies (n = 655), 71\% represent cases in which a reported
+significant result (p \textless{} 0.05) is recalculated as
+non-significant (p \textgreater{} 0.05).
 
@@
 
-\subsection{4.2 Manual inspection of decision
-inconsistencies}\label{manual-inspection-of-decision-inconsistencies}
-
-To further understand the nature of decision inconsistencies, we
-manually inspected all of them (n=655). We (a) evaluated whether
-statcheck has extracted the information correctly, and (b) assessed
-whether the text actually suggest an erroneous inferential decision,
-i.e.~whether the paper uses a reported significant result to claim a
-significant effect or a reported non-significant result to claim a null
-result.
-
 \section{5. Discussion and
@@
 
-The present study found a concerning amount of statistical reporting
+The present study found a large amount of statistical reporting
 inconsistencies across a sample of 5804 experimental linguistic
@@
 The present study can be considered a conceptual replication of previous
-studies investigating statistical-reporting inconsistency in psychology
-(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011},
-\citeproc{ref-bakker2014outlier}{2014};
+studies investigating statistical-reporting inconsistency across
+different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
+2011}, \citeproc{ref-bakker2014outlier}{2014};
 \citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
-\citeproc{ref-claesen2023data}{Claesen et al., 2023};
-\citeproc{ref-green2018statcheck}{Green et al., 2018};
-\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016};
-\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020};
 \citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
-\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}), medical
-sciences (\citeproc{ref-garcia2004incongruence}{Garcı́a-Berthou \&
-Alcaraz, 2004},; \citeproc{ref-van2023comparing}{Van Aert et al.,
-2023}), psychiatry (\citeproc{ref-berle2007inconsistencies}{Berle \&
-Starcevic, 2007}), cyber security
-studies(\citeproc{ref-gross2021fidelity}{Groß, 2021}), technological
-education research (\citeproc{ref-buckley2023estimating}{Buckley et al.,
-2023}), and experimental philosophy
-(\citeproc{ref-colombo2018statistical}{Colombo et al., 2018}). These
+\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
+recent such assessment using Statcheck
+(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
+\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
+\citeproc{ref-gross2021fidelity}{Groß, 2021};
+\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Previous
 studies report on inconsistency rates between 4\% and 14\%, with between
 10\% and 63\% of articles containing at least one inconsistency and
-between 3\% and 21\% decision consistencies.
+between 3\% and 21\% contain at least one decision consistencies.
 
@@
 cannot test without access to the data), the sheer amount of these
-errors that have made it through peer-review should concern us. They are
-human errors. If such a substantial amount of errors is found in plain
-site, the question naturally arises how many errors that happen during
-the data analysis itself remain undetected. If the tip of the iceberg
-above water is so large, how large is the iceberg underneath the
-surface?
-
-The present examination also looked at whether the rates of
-inconsistency changed over the last 23 years of publications.
-Descriptive assessment did not indicate any noticeable trend. Moreover,
-the rates of both inconsistencies and decision inconsistencies seem to
-remain stable across journals.
+inconsistencies that have made it through peer-review should concern us.
+They are human errors. If such a substantial amount of errors is found
+in plain site, the question naturally arises as to how many errors
+during the data analysis itself remain undetected. If the tip of the
+iceberg above water is already so large, how large is the body of the
+iceberg underneath the surface?
+
+The present examination assessment did not indicate any noticeable trend
+over the last 23 years of publications nor did the data suggest large
+differences between journals.
 
@@
 of decision inconsistencies was higher for p-values reported as
-significant than for those reported as non-significant. This could
-indicate a systematic bias in favor of lower p-values in general and a
-bias towards significant results in particular. Our data do not speak to
-the causes of these biases, but possible reasons include the following:
-Researchers might intentionally round down p-values because they think a
-lower p-value is more convincing to reviewers and/or readers. This
-practice has been admitted to by 1 in 5 surveyed psychological
-researchers (\citeproc{ref-john2012measuring}{John et al., 2012}). Given
-that quantitative linguists have been reported to commit questionable
-research practices (and even fraud) more often than we would like to
+significant than for those reported as non-significant. These patterns
+could indicate a systematic bias in favor of lower p-values in general
+and a bias towards significant results in particular. Our data do not
+speak to the causes of these biases, but possible reasons include the
+following: Researchers might intentionally round down p-values because
+they think a lower p-value is more convincing to reviewers and/or
+readers. This practice has been admitted to by 1 in 5 surveyed
+psychological researchers (\citeproc{ref-john2012measuring}{John et al.,
+2012}). Given that a non-trivial amount of quantitative linguists have
+admitted to commit questionable research practices (and even fraud)
 (\citeproc{ref-isbell2022misconduct}{Isbell et al., 2022}), we cannot
@@
 Researchers might scrutinize non-significant results more than
-significant results or fail to double check significant results more
-often because results that confirm their hypothesis feed into their
-confirmation bias (\citeproc{ref-nickerson1998confirmation}{Nickerson,
-1998}). For example, Fugelsang et al.
-(\citeproc{ref-fugelsang2004theory}{2004}) let researchers evaluate data
-that are either consistent or inconsistent with their prior
-expectations. They showed that when researchers encounter results that
-disconfirm their expectations, they are likely to blame the methodology
-while results that confirmed their expectations were rarely critically
-scrutinized. Alternatively, the observed bias might be a reflection of
-publication bias (\citeproc{ref-franco2014publication}{Franco et al.,
-2014}; \citeproc{ref-sterling1959publication}{Sterling, 1959}) with
+significant results or are less likely to double check significant
+results than non-significant results because results that confirm their
+hypothesis feed into their confirmation bias
+(\citeproc{ref-nickerson1998confirmation}{Nickerson, 1998}). For
+example, Fugelsang et al. (\citeproc{ref-fugelsang2004theory}{2004}) let
+researchers evaluate data that are either consistent or inconsistent
+with their prior expectations. They showed that when researchers
+encounter results that disconfirm their expectations, they are likely to
+blame the methodology while results that confirmed their expectations
+were rarely critically scrutinized. Alternatively, the observed bias
+might be a reflection of publication bias
+(\citeproc{ref-franco2014publication}{Franco et al., 2014};
+\citeproc{ref-sterling1959publication}{Sterling, 1959}) with
 (erroneously) reported significant p-values being more likely to be
@@
 that the inconsistency rates of our study are comparable to similar
-studies from other disciplines and that the inconsistency rates are both
-comparable across the eight journals and stable across the most recent
-23 years, suggests that our findings are relevant for experimental
-linguistics at large.
+studies from other disciplines and that those rates are relatively
+stable across journals and time, our findings should be considered
+relevant for experimental linguistics at large.
 
 Second, given the constraints on automatically detecting test
-statistics, statcheck misses reported values that either diverge from
+statistics, Statcheck misses reported values that either diverge from
 APA reporting standards or are reported in tables. However,
@@
 (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011};
-\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Third,
-statcheck slightly overestimates inconsistency rates, because it might
-not accurately detect corrections for multiple comparisons
+\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}).
+
+Third, statcheck slightly overestimates inconsistency rates, because it
+might not accurately detect corrections for multiple comparisons
 (\citeproc{ref-schmidt2017statcheck}{Schmidt, 2017}). Nuijten et al.
@@
 inconsistencies'' (\citeproc{ref-nuijten2017validity}{Nuijten et al.,
-2017, p. 27}). More elaborate automatic tools for the extraction of
-statistical information might allow a more detailed and more accurate
-assessment of statistical reporting in the future (e.g.
-\citeproc{ref-kalmbach2023rule}{Kalmbach et al., 2023}). But even
-statcheck is limited and only provides a rough proxy of true
-inconsistency rate in the published literature, we hope the reader
-agrees that it is a state of affairs that should be tackled.
+2017, p. 27}).
+
+More elaborate automatic tools for the extraction of statistical
+information might allow for a more detailed and more accurate assessment
+of statistical reporting in the future (e.g.
+\citeproc{ref-kalmbach2023rule}{Kalmbach et al., 2023}). Despite its
+limitations, Statcheck provides a rough proxy of true inconsistency rate
+in the published literature and we hope the reader agrees that the
+prevalence of inconsistencies is a state of affairs that should be
+reflected upon.
 
@@
 linguistics can make to reduce statistical reporting inconsistencies. In
-order to avoid simple copy and paste errors related to working in two
+order to avoid simple copy-and-paste errors related to working in two
 separate programs for writing the manuscript and conducting the
@@
 common R markdown files (Rmd) and Quarto markdown files (qmd). Literate
-programming can ensure that values derived by the statistical analysis
-is automatically integrated into the manuscript document, avoiding
-errors that might happen during a transfer from one program to the
-other.
+programming can ensure that values derived from the statistical analysis
+are automatically integrated into the manuscript document, avoiding
+errors that might happen during a manual transfer from one program to
+the other.
 
@@
 detect errors or even inappropriate statistical choices before
-publication, thus improving the quality of their work. Moreover,
-publicly sharing their analyses has numerous benefits to the authors
-themselves beyond error detection: Open data and materials can
-facilitate collaboration (\citeproc{ref-boland2017ten}{Boland et al.,
-2017}), increase efficiency and sustainability
-(\citeproc{ref-lowndes2017our}{Lowndes et al., 2017}), and are cited
-more often (\citeproc{ref-colavizza2020citation}{Colavizza et al.,
-2020}).
+publication, thus improving the quality and robustness of the final
+product. Moreover, publicly sharing their analyses has numerous benefits
+to the authors themselves beyond error detection: Open data and
+materials can facilitate collaboration
+(\citeproc{ref-boland2017ten}{Boland et al., 2017}), increase efficiency
+and sustainability (\citeproc{ref-lowndes2017our}{Lowndes et al.,
+2017}), and are cited more often
+(\citeproc{ref-colavizza2020citation}{Colavizza et al., 2020}).
 
@@
 (\citeproc{ref-schonbrodt2015p}{Schönbrodt, 2015},
-http://shinyapps.org/apps/p-checker/). Reviewers could consider consider
+http://shinyapps.org/apps/p-checker/). Reviewers could consider
 requesting data and scripts during peer review. Such requests might be
@@
 research question (\citeproc{ref-coretta_multidimensional_2023}{Coretta
-et al., 2023}). Some of these approaches are might be more appropriate
-than others (\citeproc{ref-sondereggera2024advancements}{Sonderegger \&
+et al., 2023}). Some of these approaches might be more appropriate than
+others (\citeproc{ref-sondereggera2024advancements}{Sonderegger \&
 Sóskuthy, 2024}; \citeproc{ref-vasishth2023some}{Vasishth, 2023}), so
@@
 
-Journal editors could explicitly recommend consistency checks with apps
-such as statcheck during peer review, a practice that has been taken up
-on by several journals from neighboring disciplines (Psychological
-Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
+Journal editors could explicitly recommend consistency checks with
+algorhythms such as statcheck during peer review, a practice that has
+been taken up on by several journals from neighboring disciplines
+{[}Psychological Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
   accessed on July 15, 2024.}, Advances in Methods and Practices in
 Psychological Science\footnote{https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines;
-  accessed on on July 15, 2024.}, Stress \& Health
-\citeproc{ref-barber2017meticulous}{Barber, 2017}). Editors could also
+  accessed on on July 15, 2024.}, Stress \& Health Barber
+(\citeproc{ref-barber2017meticulous}{2017}){]}. Editors could also
 demand, recommend or at least encourage data sharing for publication in
@@
 \citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022}).
+
+Researchers make errors. Researchers have biases. This is who we are as
+humans and there is not much we can do about our nature. Being aware of
+this fact and how it might affect research might help us to make
+possibly negative consequences at worst detectable and best preventable.
 
@@
 
+\bibitem[\citeproctext]{ref-rmanual}
+R Core Team. (2021). \emph{R: A language and environment for statistical
+computing}. R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}
+
 \bibitem[\citeproctext]{ref-roettger2019researcher}

### 2024-07-18 scripts/01_Analysis.R

@@ -22,7 +22,6 @@
 library(rstudioapi)
 library(tidyverse)
 library(ggpubr)
-library(broom)
 library(statcheck)
 library(janitor)
 library(patchwork)
@@ -32,9 +31,10 @@ library(ggstream)
 ## data ########################################################################                        
 
 
-## load in data from location of script (uncomment if you run script outside of manuscript.qmd)
-#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
-#setwd(current_working_dir)
+## load in data from location of script 
+## (uncomment if you run script outside of sourcing it through manuscript.qmd)
+# current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
+# setwd(current_working_dir)
 
 ## list all journal units
 files <- list.files("../Journals", recursive = TRUE, pattern = "\\.pdf")
@@ -98,8 +98,12 @@ journal_summary_counts <- xdata %>%
   adorn_totals(name = "Total") %>% 
   select(journal, n.art, checked_articles, checked_results, errors, gross_errors)
 
-colnames(journal_summary_counts) <- c("Journal", "eligible articles", "assessible articles",
+colnames(journal_summary_counts) <- c("Journal", "eligible articles", "assessable articles",
                                       "assessible results", "inconsistencies", "decision inconsistencies")
+
+# prevalence of errors for different comparison signs
+sign_distribution <- xtabs(~error + p_comp, xdata)
+sign_distribution_decision <- xtabs(~decision_error + p_comp, xdata)
 
 
 ## summary table proportions
@@ -155,7 +159,6 @@ per_article_long <- per_article %>%
 
 
 ## Figure 1 ####################################################################                       
-
 
 ## plot as stacked barplot (figure 1)
 stacked_bar <- 
@@ -323,7 +326,7 @@ xdata <- xdata %>%
 
 xdata_sub <- xdata %>% 
   filter(error == T,
-         test_comp == "=")
+         p_comp == "=")
 
 reportedVcomputed <- ggplot(xdata_sub) + 
   geom_point(data = xdata_sub %>% filter(decision_error == F),
@@ -400,8 +403,92 @@ ggplot(xdata_sub) +
   theme_minimal() 
 
 
-xdata_sub <- xdata %>% 
-  filter(error == T,
-         test_comp == "=")
-
-
[…]

### 2024-07-18 manuscript/manuscript.qmd
typo
added abstract

@@
 ---
-title: "Statistical Reporting Inconsistencies in Experimental Linguistics"
+title: "The (mis)reporting of statistical results in experimental linguistics"
 shorttitle: "Statistical Inconsistencies"
@@
     authorship-agreements: null
-abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum."
+abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals, published between 2000 and 2023.
+Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were inconsistent with their test statistic and degrees of freedom.
+Half of all articles that used null hypothesis signficance testing contained at least one inconsistent p-value.
+One in eight articles contained an inconsistency that may have affected the statistical conclusion.
+The inconsistency rates were stable across journals and publication years.
+We offer actionable steps for authors, reviewers, and editors to remmedy this state-of-affairs."
 keywords: [statistics, statcheck, reproducibility]
@@
 This is who we are as humans and there is not much we can do about our nature. 
-Being aware of this fact and how it might affect research might help us to make possibly negative consequences at worst detectable and best preventable.
+Being aware of this fact and how it might affect research might help us to make possibly negative consequences detectable and preventable.
 

### 2024-07-18 manuscript/manuscript.tex
typo
added abstract

@@
 
-\title{Statistical Reporting Inconsistencies in Experimental
-Linguistics}
+\title{The (mis)reporting of statistical results in experimental
+linguistics}
 
@@
 
-\abstract{Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed
-do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
-ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit
-in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui
-officia deserunt mollit anim id est laborum.}
+\abstract{The present paper investigates the prevalence of statistical
+reporting consistencies across articles in eight experimental
+linguistics journals, published between 2000 and 2023. Using the R
+package ``Statcheck'', we retrieved 39532 statistical test and assessed
+whether p-values were inconsistent with their test statistic and degrees
+of freedom. Half of all articles that used null hypothesis signficance
+testing contained at least one inconsistent p-value. One in eight
+articles contained an inconsistency that may have affected the
+statistical conclusion. The inconsistency rates were stable across
+journals and publication years. We offer actionable steps for authors,
+reviewers, and editors to remmedy this state-of-affairs.}
 
@@
 this fact and how it might affect research might help us to make
-possibly negative consequences at worst detectable and best preventable.
+possibly negative consequences detectable and preventable.
 

### 2024-07-25 manuscript/manuscript.qmd
added credit scores and CoI

@@
 ---
-title: "The (mis)reporting of statistical results in experimental linguistics"
+title: "Statistical Reporting Inconsistencies in Experimental Linguistics"
 shorttitle: "Statistical Inconsistencies"
@@
     related-report: null
-    conflict-of-interest: null
+    conflict-of-interest: The authors have no conflict of interest to declare.
     financial-support: null
     gratitude: null
-    authorship-agreements: null
+    authorship-agreements: Conceptionalization, Methodology, Validation, Formal Analysis, Review & Editing of Manuscript, Data Curation - T.B.R.
+& D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft, Visualization, Supervision - T.B.R.
 abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals, published between 2000 and 2023.
@@
 Without access to data and scripts, it remains unclear to the reader what has caused this inconsistency.
-Such inconsistencies can be particularly concerning if the calculated p-value (here 0.057) and the reported p-values (here >0.05) are not on the same side of the alpha threshold.
+Such inconsistencies can be particularly concerning if the calculated p-value (here 0.057) and the reported p-values (here <0.05) are not on the same side of the alpha threshold.
 In NHST, p-values below a conventionalized alpha threshold, most commonly 0.05, are interpreted as evidence that the data are sufficiently inconsistent with the null hypothesis ("significant").
@@
 ## 3.1. Statcheck
-We used the R package wtatcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
+We used the R package Statcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
 Statcheck works as follows: After converting articles in pdf or html format to plain text, Statcheck searches for specific strings that correspond to a NHST result using “regular expressions”.
@@
 ## 3.2. Sample
-Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the term “experiment\* in title, abstract and/or keywords.
+Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the search string "experiment\*" in title, abstract and/or keywords.
 Out of these 100 journals, we selected all journals with at least 10% of articles containing the search string "experiment\*".
@@
 Statcheck could not parse 157 articles, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt. 
+No assessable statistics were found by Statcheck in JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
 This procedure resulted in `{r} length(files)` research articles which were submitted to analysis.
@@
 
-The proportion of inconsistencies ranged from `{r} round(min(journal_summary_prop[,2]),0)` to `{r} round(max(journal_summary_prop[,2]),0)`% across journals (`{r} round(min(journal_summary_prop[,3]),0)` to `{r} round(max(journal_summary_prop[,3]),0)`% for decisions inconsistencies) (see @fig-stacked-bar).
+The proportion of inconsistencies ranged from `{r} round(min(journal_summary_prop[,2]),0)` to `{r} round(max(journal_summary_prop[,2]),0)`% across journals (`{r} round(min(journal_summary_prop[,3]),0)` to `{r} round(max(journal_summary_prop[,3]),0)`% for decision inconsistencies) (see @fig-stacked-bar).
 These rates appear to be stable across year of publication (see @fig-year-prop).
-On average, `{r} 100*round(gross_per_article[[1]],2)`% of assessable articles contained one or more inconsistencies (journals range from `{r} 100*round(min(per_article[,2]),2)` to `{r} 100*round(max(per_article[,2]),2)`%) and `{r} 100*round(gross_per_article[[2]],2)`% contained one or more decisions inconsistencies (journals range from `{r} 100*round(min(per_article[,3]),2)` to `{r} 100*round(max(per_article[,3]),2)`%).
+On average, `{r} 100*round(gross_per_article[[1]],2)`% of assessable articles contained one or more inconsistencies (journals range from `{r} 100*round(min(per_article[,2]),2)` to `{r} 100*round(max(per_article[,2]),2)`%) and `{r} 100*round(gross_per_article[[2]],2)`% contained one or more decision inconsistencies (journals range from `{r} 100*round(min(per_article[,3]),2)` to `{r} 100*round(max(per_article[,3]),2)`%).
 
@@
 Second, there is a linear density of points along the diagonal axis which corresponds to numerically small inconsistencies, some of which might be related to simple rounding errors.
-However, comparing the diagonal to the black line, which represents a linear model predicting recalculated by reported p-values, we can see a clear divergence of what is expected if inconsistencies were equally likely in both directions (i.e.
-recalculated p-values were as likely to be smaller than the reported p-value as larger).
+However, comparing the diagonal to the black line, which represents a linear model predicting recalculated by reported p-values, we can see a clear divergence of what is expected if inconsistencies were equally likely in both directions.
 The slope of the regression line is flatter than the diagonal axis which means that, on average, inconsistencies have a tendency to exhibit smaller reported p-values than their recalculated counterparts.
@@
 So even if we assumed these inconsistencies were merely typos of the comparison sign, inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.
-These biases are also reflected in decisions inconsistencies.
-Of all decisions inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant (p \> 0.05).
+These biases are also reflected in decision inconsistencies.
+Of all decision inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant (p \> 0.05).
 
@@
 The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent such assessment using Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
-Previous studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% contain at least one decision consistencies.
+Previous studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% contain at least one decision consistency.
 
@@
 If such a substantial amount of errors is found in plain site, the question naturally arises as to how many errors during the data analysis itself remain undetected.
-If the tip of the iceberg above water is already so large, how large is the body of the iceberg underneath the surface?
+If the tip of the iceberg is already so large, what is the volume of the submerged iceberg?
 
@@
 :::
-
-# Appendix
-
-# Title for Appendix

### 2024-07-25 manuscript/manuscript.tex
added credit scores and CoI

@@
 
-\title{The (mis)reporting of statistical results in experimental
-linguistics}
+\title{Statistical Reporting Inconsistencies in Experimental
+Linguistics}
 
@@
 \par{ }
-\par{       }
+\par{   The authors have no conflict of interest to
+declare.   Conceptionalization, Methodology, Validation, Formal
+Analysis, Review \& Editing of Manuscript, Data Curation - T.B.R. \&
+D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft,
+Visualization, Supervision - T.B.R. }
 \par{Correspondence concerning this article should be addressed to Timo
@@
 concerning if the calculated p-value (here 0.057) and the reported
-p-values (here \textgreater0.05) are not on the same side of the alpha
+p-values (here \textless0.05) are not on the same side of the alpha
 threshold. In NHST, p-values below a conventionalized alpha threshold,
@@
 
-We used the R package wtatcheck (Version 1.4.1-beta.2,
+We used the R package Statcheck (Version 1.4.1-beta.2,
 \citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
@@
 articles published at the time of assessment (2021) and a high ratio of
-articles containing the term ``experiment* in title, abstract and/or
-keywords. Out of these 100 journals, we selected all journals with at
-least 10\% of articles containing the search string''experiment*``. Out
-of the remaining 37 journals, we selected those journals that urged APA
-formatting either in the main body of the text or specifically regarding
-statistics in the author guidelines, resulting in nine remaining
-journals. Moreover, to access the articles in .pdf format, the articles
-had to be either accessible to us through our library license, or open
-access, resulting in a final list of eight journals: Applied
-Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
-Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
-Language Learning and Techology (LLT), Journal of Language and Social
-Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
-Language Acquisition (SLA).
+articles containing the search string ``experiment*'' in title, abstract
+and/or keywords. Out of these 100 journals, we selected all journals
+with at least 10\% of articles containing the search string
+``experiment*''. Out of the remaining 37 journals, we selected those
+journals that urged APA formatting either in the main body of the text
+or specifically regarding statistics in the author guidelines, resulting
+in nine remaining journals. Moreover, to access the articles in .pdf
+format, the articles had to be either accessible to us through our
+library license, or open access, resulting in a final list of eight
+journals: Applied Psycholinguistics (APS), Bilingualism: Language and
+Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
+and Speech (LaS), Language Learning and Techology (LLT), Journal of
+Language and Social Psychology (LSP), Journal of Child Language (JCL),
+and Studies in Second Language Acquisition (SLA).
 
@@
 issues with rendering the Chi-Squared symbol being erroneously converted
-from .pdf to .txt. This procedure resulted in 5804 research articles
-which were submitted to analysis.
+from .pdf to .txt. No assessable statistics were found by Statcheck in
+JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
+This procedure resulted in 5804 research articles which were submitted
+to analysis.
 
@@
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-assessible results
+assessable results
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
@@
 The proportion of inconsistencies ranged from 10 to 22\% across journals
-(1 to 7\% for decisions inconsistencies) (see
+(1 to 7\% for decision inconsistencies) (see
 Figure~\ref{fig-stacked-bar}). These rates appear to be stable across
@@
 of assessable articles contained one or more inconsistencies (journals
-range from 43 to 56\%) and 13\% contained one or more decisions
+range from 43 to 56\%) and 13\% contained one or more decision
 inconsistencies (journals range from 11 to 15\%).
@@
 reported p-values, we can see a clear divergence of what is expected if
-inconsistencies were equally likely in both directions
-(i.e.~recalculated p-values were as likely to be smaller than the
-reported p-value as larger). The slope of the regression line is flatter
-than the diagonal axis which means that, on average, inconsistencies
-have a tendency to exhibit smaller reported p-values than their
-recalculated counterparts. Similarly, inconsistencies that report the
-p-value as being larger or smaller than a reference value (e.g.~p
-\textgreater{} 0.05 or p \textless{} 0.05 ) are not equally prevalent.
-There were 4.4\% of inconsistencies with p being reported as larger than
-a reference but 7.4\% inconsistencies with p being reported as smaller
-than a reference (e.g.~p \textless{} 0.05). So even if we assumed these
-inconsistencies were merely typos of the comparison sign,
-inconsistencies that erroneously report the p-value to be smaller than a
-reference value are more frequent than inconsistencies that erroneously
-report the p-value to be larger than a reference value. These biases are
-also reflected in decisions inconsistencies. Of all decisions
-inconsistencies (n = 655), 71\% represent cases in which a reported
-significant result (p \textless{} 0.05) is recalculated as
+inconsistencies were equally likely in both directions. The slope of the
+regression line is flatter than the diagonal axis which means that, on
+average, inconsistencies have a tendency to exhibit smaller reported
+p-values than their recalculated counterparts. Similarly,
+inconsistencies that report the p-value as being larger or smaller than
+a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05 )
+are not equally prevalent. There were 4.4\% of inconsistencies with p
+being reported as larger than a reference but 7.4\% inconsistencies with
+p being reported as smaller than a reference (e.g.~p \textless{} 0.05).
+So even if we assumed these inconsistencies were merely typos of the
+comparison sign, inconsistencies that erroneously report the p-value to
+be smaller than a reference value are more frequent than inconsistencies
+that erroneously report the p-value to be larger than a reference value.
+These biases are also reflected in decision inconsistencies. Of all
+decision inconsistencies (n = 655), 71\% represent cases in which a
+reported significant result (p \textless{} 0.05) is recalculated as
 non-significant (p \textgreater{} 0.05).
@@
 10\% and 63\% of articles containing at least one inconsistency and
-between 3\% and 21\% contain at least one decision consistencies.
+between 3\% and 21\% contain at least one decision consistency.
 
@@
 during the data analysis itself remain undetected. If the tip of the
-iceberg above water is already so large, how large is the body of the
-iceberg underneath the surface?
+iceberg is already so large, what is the volume of the submerged
+iceberg?
 
@@
 Processing}, 2350--2361.
+\url{https://doi.org/10.18653/v1/2022.emnlp-main.150}
 
@@
 Methods}, \emph{43}, 666--678.
+\url{https://doi.org/10.3758/s13428-011-0089-5}
 
@@
 One}, \emph{9}(7), e103360.
+\url{https://doi.org/10.1371/journal.pone.0103360}
 
@@
 together for robust science reporting. \emph{Stress \& Health},
-\emph{33}(2), 89--91.
+\emph{33}(2), 89--91. \url{https://doi.org/10.1002/smi.2756}
 
@@
 \emph{International Journal of Methods in Psychiatric Research},
-\emph{16}(4), 202--207.
+\emph{16}(4), 202--207. \url{https://doi.org/10.1002/mpr.225}
 
@@
 linguistics. \emph{Glossa Psycholinguistics}, \emph{2}(1).
+\url{https://doi.org/10.5070/G6011239}
 
@@
 Library of Science San Francisco, CA USA.
+\url{https://doi.org/10.1371/journal.pcbi.1005278}
 
@@
 Journal of Technology and Design Education}, \emph{33}(4), 1243--1264.
+\url{https://doi.org/10.1007/s10798-022-09787-6}
 
@@
 reported in spanish psychology journals. \emph{Psicothema},
-\emph{25}(3), 408--414.
+\emph{25}(3), 408--414. \url{https://doi.org/10.7334/psicothema2012.207}
 
@@
 Demystifying reproducibility and transparency practices in linguistic
-research}.
+research}. \url{https://doi.org/10.31234/osf.io/spz4w}
 
@@
 molenaar (2011). \emph{Plos One}, \emph{18}(4), e0284243.
+\url{https://doi.org/10.1371/journal.pone.0284243}
 
@@
 to research data. \emph{PloS One}, \emph{15}(4), e0230416.
+\url{https://doi.org/10.1371/journal.pone.0230416}
 
@@
 \emph{PloS One}, \emph{13}(4), e0194360.
+\url{https://doi.org/10.1371/journal.pone.0194360}
 
@@
 \emph{345}(6203), 1502--1505.
+\url{https://doi.org/10.1126/science.1255484}
 
@@
 Exp{é}rimentale}, \emph{58}(2), 86.
+\url{https://doi.org/10.1037/h0085799}
 
@@
 Methodology}, \emph{4}, 1--5.
+\url{https://doi.org/10.1186/1471-2288-4-13}
 
@@
 \emph{Canadian Psychology/Psychologie Canadienne}, \emph{59}(3), 203.
+\url{https://doi.org/10.1037/cap0000139}
 
@@
 Luxembourg, September 26, 2019, Revised Selected Papers 9}, 3--26.
+\url{https://doi.org/10.1007/978-3-030-55958-8_1}
 
@@
 at the journal cognition. \emph{Royal Society Open Science},
-\emph{5}(8), 180448.
+\emph{5}(8), 180448. \url{https://doi.org/10.1098/rsos.180448}
 
@@
 Journal}, \emph{106}(1), 172--195.
+\url{https://doi.org/10.1111/modl.12760}
 
@@
 telling. \emph{Psychological Science}, \emph{23}(5), 524--532.
+\url{https://doi.org/10.1177/0956797611430953}
 
@@
 Information Systems}, 326--338.
+\url{https://doi.org/10.1007/978-3-031-35320-8_23}
 
@@
 in experimental linguistics. \emph{Glossa Psycholinguistics},
-\emph{2}(1), 1--28.
+\emph{2}(1), 1--28. \url{https://doi.org/10.5070/G6011135}
 
@@
 policy. \emph{Journal of Memory and Language}, \emph{125}, 104332.
+\url{https://doi.org/10.1016/j.jml.2022.104332}
 
@@
 \emph{Nature Ecology \& Evolution}, \emph{1}(6), 0160.
+\url{https://doi.org/10.1038/s41559-017-0160}
 
@@
 many guises. \emph{Review of General Psychology}, \emph{2}(2), 175--220.
+\url{https://doi.org/10.1037/1089-2680.2.2.175}
 
@@
 in discovering statistical reporting inconsistencies}.
+\url{https://doi.org/10.31234/osf.io/tcxaj}
 
@@
 in psychology (1985--2013). \emph{Behavior Research Methods}, \emph{48},
-1205--1226.
+1205--1226. \url{https://doi.org/10.5070/G6011135}
 
@@
 research. \emph{Laboratory Phonology}, \emph{10}(1).
+\url{https://doi.org/10.5334/labphon.147}
 
@@
 science. \emph{Perspectives on Psychological Science}, \emph{9}(6),
-652--660.
+652--660. \url{https://doi.org/10.1177/1745691614549257}
 
@@
 Reply to nuijten et al.(2017)}.
+\url{https://doi.org/10.31234/osf.io/hr6qy}
 
@@
 in the 21st century: Quantitative data analysis}.
+\url{https://doi.org/10.31234/osf.io/mc6a9}
 
@@
 \emph{Journal of the American Statistical Association}, \emph{54}(285),
-30--34.
+30--34. \url{https://doi.org/10.2307/2282137}
 
@@
 \emph{Royal Society Open Science}, \emph{10}(8), 202326.
+\url{https://doi.org/10.1098/rsos.202326}
 
@@
 data. \emph{Annual Review of Linguistics}, \emph{9}(1), 273--291.
+\url{https://doi.org/10.1146/annurev-linguistics-031220-010345}
 
@@
 \emph{PloS One}, \emph{9}(12), e114876.
+\url{https://doi.org/10.1371/journal.pone.0114876}
 
@@
 quality of reporting of statistical results. \emph{PloS One},
-\emph{6}(11), e26828.
+\emph{6}(11), e26828. \url{https://doi.org/10.1371/journal.pone.0026828}
 
@@
 
-\appendix
-
-\section{Title for Appendix}\label{title-for-appendix}
-
 \end{document}

### 2024-07-25 scripts/01_Analysis.R

@@ -99,7 +99,7 @@ journal_summary_counts <- xdata %>%
   select(journal, n.art, checked_articles, checked_results, errors, gross_errors)
 
 colnames(journal_summary_counts) <- c("Journal", "eligible articles", "assessable articles",
-                                      "assessible results", "inconsistencies", "decision inconsistencies")
+                                      "assessable results", "inconsistencies", "decision inconsistencies")
 
 # prevalence of errors for different comparison signs
 sign_distribution <- xtabs(~error + p_comp, xdata)

### 2024-07-25 manuscript/manuscript_anonymized.qmd
anonymize

--- 2024-07-25 manuscript/manuscript.qmd
@@
 author:
-  - name: Dara Leonard Jenssen Etemady
-    corresponding: false
-    email: daraetemady@gmail.com
+    - name: anonymous
       affiliations:
-      - name: University of Oslo
-        department: Department of Linguistics & Scandinavian Studies
-        city: Oslo
-        country: Norway
-  - name: Timo B. Roettger
-    corresponding: true
-    orcid: 0000-0000-0000-0001
-    email: timo.roettger@gmail.com
-    affiliations:
-      - name: University of Oslo
-        department: Department of Linguistics & Scandinavian Studies
-        city: Oslo
-        country: Norway
+      - name: anonymous
+  # - name: Dara Leonard Jenssen Etemady
+  #   corresponding: true
+  #   email: daraetemady@gmail.com
+  #   affiliations:
+  #     - name: University of Oslo
+  #       department: Department of Linguistics & Scandinavian Studies
+  #       city: Oslo
+  #       country: Norway
+  # - name: Timo B. Roettger
+  #   corresponding: false
+  #   orcid: 0000-0000-0000-0001
+  #   email: timo.roettger@gmail.com
+  #   affiliations:
+  #     - name: University of Oslo
+  #       department: Department of Linguistics & Scandinavian Studies
+  #       city: Oslo
+  #       country: Norway
 author-note:
@@
     gratitude: null
-    authorship-agreements: Conceptionalization, Methodology, Validation, Formal Analysis, Review & Editing of Manuscript, Data Curation - T.B.R.
+    # authorship-agreements: Conceptionalization, Methodology, Validation, Formal Analysis, Review & Editing of Manuscript, Data Curation - T.B.R.
 & D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft, Visualization, Supervision - T.B.R.
-abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals, published between 2000 and 2023.
-Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were inconsistent with their test statistic and degrees of freedom.
+abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals published between 2000 and 2023.
+Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were consistent with their test statistic and degrees of freedom.
 Half of all articles that used null hypothesis signficance testing contained at least one inconsistent p-value.
 One in eight articles contained an inconsistency that may have affected the statistical conclusion.
-The inconsistency rates were stable across journals and publication years.
-We offer actionable steps for authors, reviewers, and editors to remmedy this state-of-affairs."
-keywords: [statistics, statcheck, reproducibility]
+The inconsistency rates were comparable across journals and seem stable over publication years.
+We discuss possible reasons for this high rate and offer actionable steps for authors, reviewers, and editors to remedy this state-of-affairs."
+keywords: [statistics, statcheck, reproducibility, null hypothesis significance testing]
 bibliography: bibliography.bib
@@
 # 1. Introduction
+
 What we know about human language and its cognitive underpinnings is often informed by experimental data.
-Researchers test theoretical predictions with these data using statistical tests.
-Depending on the results of such tests, researchers make claims for or against theoretical assumptions.\
-Since they play such an integral part in the reasoning process of experimentalists, both the data these tests are based on and the computational procedure of the tests itself should be transparent in order for other researchers to critically evaluate them.
-Moreover things can go wrong.
-To err is human.
+Researchers test theoretical predictions with their data using statistical tests.
+Depending on the results of their tests, researchers make claims for or against theoretical assumptions.\
+Since these tests play such an integral part in the argumentation process of experimentalists, both the data these tests are based on and the computational procedure of the tests itself should be transparent in order for other researchers to critically evaluate them.
+But more than that, things can go wrong.
+Humans are fallible; they make mistakes.
 Transparent sharing allows others to detect and correct human error.
-In the recent years, the quantitative sciences have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
+In the recent years, quantitative linguistics have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
 Despite these calls, sharing of statistical protocols is still rather rare across the language sciences [@bochynska2023reproducible].
@@
 # 2. Statistal reporting inconsistencies
+
 The null-hypothesis significance testing (NHST) framework is, to date, the most dominant statistical framework that researchers use to test hypotheses in the language sciences [@sondereggera2024advancements].
@@
 given that the test statistic is zero) (see example 1):
-&nbsp;
-
-&nbsp;
 
@@
 \annotate[yshift=-1em]{below,right}{node7}{p-value}
+```
 
-```
-&nbsp;
-
-Since data and statistical protocol sharing remains rare across the language sciences [@bochynska2023reproducible], interested readers are left with trusting the authors that the statistical analysis is run and reported correctly.
-However, the three sets of indices in (1) have clearly defined mathematical relationships and can thus be easily checked for consistency.
-An F test with the specified degrees of freedom and a test-statistic of 3.88 should result in a p-value of 0.053 which is larger, not smaller, than 0.05.
-Possible reasons for this inconsistency are manifold: It could be a typo of the comparison sign, i.e. the authors meant to use `=` or `>` rather than `<`.
-Any of the numbers could contain a typo.
-Sometimes an error might indicate erroneous rounding (e.g. 0.057 being rounded down to 0.05).
+Without access to the data and scripts, interested readers are left with trusting the authors that the statistical analysis is run and reported correctly.
+However, the three sets of indices in (1) have clearly defined mathematical relationships and can thus be easily assessed for consistency.
+For example, an F test with the specified degrees of freedom and a test-statistic of 3.88 should result in a p-value of 0.053 which is larger, not smaller, than 0.05.
+Possible reasons for this inconsistency are manifold: It could be a typo of the comparison sign, i.e. the authors meant to use `=` or `<` rather than `>`.
+Additionally, any of the numbers could contain a typo and sometimes an error might indicate erroneous rounding (e.g. 0.057 being rounded down to 0.05).
 Without access to data and scripts, it remains unclear to the reader what has caused this inconsistency.
-Such inconsistencies can be particularly concerning if the calculated p-value (here 0.057) and the reported p-values (here <0.05) are not on the same side of the alpha threshold.
+Such inconsistencies can be particularly concerning if the calculated p-value (here 0.057) and the reported p-values (here \<0.05) are not on the same side of the alpha threshold.
 In NHST, p-values below a conventionalized alpha threshold, most commonly 0.05, are interpreted as evidence that the data are sufficiently inconsistent with the null hypothesis ("significant").
@@
 Recently, a series of studies used such automatic assessments to evaluate the prevalence of inconsistent statistical reporting in psychology [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness; @nuijten2016prevalence; @nuijten2020statcheck; @green2018statcheck; @claesen2023data], medical sciences [@garcia2004incongruence; @van2023comparing], psychiatry [@berle2007inconsistencies], cyber security studies [@gross2021fidelity], technological education research [@buckley2023estimating], and experimental philosophy [@colombo2018statistical].
-For example, looking at over 250 thousand p-values published in major psychology journals, @nuijten2016prevalence found that around 50% of the articles with statistical results contained at least one inconsistency and around 13% contained at least one “decision inconsistency”.
-Other studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% decision consistencies.
+For example, looking at over 250 thousand p-values published in major psychology journals, @nuijten2016prevalence found that around 50% of the articles with statistical results contained at least one inconsistency and around 13% contained at least one decision inconsistency.
+Other studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% contain at least one decision consistency.
+These high inconsistency rates in other disciplines are concerning and have led to a constructive dialog, resulting in recommendations for best practice to either avoid inconsistencies or make them more detectable.
 
 To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in eight experimental linguistic journals published between 2000 and 2023.
-We explore whether the rate of inconsistency differs across journals, whether that rate has changed over the course of the last 20 years and whether there is evidence for bias in these statistical-reporting inconsistencies.
+We explore whether the rate of inconsistency differs across journals, whether their rate has changed over the course of the last 23 years and whether there is evidence for bias in these statistical-reporting inconsistencies.
 We discuss the results and offer concrete recommendations for authors, reviewers, and editors to tackle this problem.
@@
 
-All quantitative analyses were conducted using @rmanual and the r packages \[LIST ALL PACKAGES from 01_Analysis.R and add to references\]
+## 3.1. Data availability statement
 
-## 3.1. Statcheck
+All quantitative analyses were conducted using @rmanual.
+All derived data and corresponding R scripts are publicly available here: [https://github.com/timo-b-roettger/ErrorsLing](https://github.com/timo-b-roettger/ErrorsLing)
+
+## 3.2. Statcheck
+
 We used the R package Statcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
-Statcheck works as follows: After converting articles in pdf or html format to plain text, Statcheck searches for specific strings that correspond to a NHST result using “regular expressions”.
-That way, statcheck can detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, the degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
-Statcheck usually misses reported in tables; and (c) the test result is reported in American Psychological Association style [@APA2020].
+Statcheck works as follows: After converting articles in pdf or html format to plain text, Statcheck searches for specific strings that correspond to a NHST result using regular expressions.
+That way, Statcheck can detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, the degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
+Statcheck usually misses information in tables; and (c) the test result is reported in American Psychological Association style [@APA2020].
 Given these constraints, Statcheck is estimated to detect roughly 60% of all reported NHST results [@nuijten2016prevalence].
-Statcheck uses the reported test statistic and degrees of freedom to recalculate the p-value, compares the reported and recalculated p-values and, if there is a mismatch, flags the test as containing an “inconsistency.”
-The algorithm takes into account that tests might have been performed as one-tailed by identifying the search strings “one-tailed,” “one-sided,” or “directional” in the body of the text.
+Statcheck uses the reported test statistic and degrees of freedom to recalculate the p-value, compares the reported and recalculated p-value and, if there is a mismatch, flags the test as containing an “inconsistency.” The algorithm takes into account that tests might have been performed as one-tailed by identifying the search strings “one-tailed,” “one-sided,” or “directional” in the body of the text.
 Moreover, Statcheck considers p = .000 and p \< .000 as inconsistent because p-values of exactly zero are mathematically impossible and the APA manual [@APA2020] advises to report very small p-values as p \< .001.
@@
 
-## 3.2. Sample
+## 3.3. Sample
+
 Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the search string "experiment\*" in title, abstract and/or keywords.
@@
 This procedure resulted in `{r} length(files)` research articles which were submitted to analysis.
-
-## 3.3. Data availability statement
-All derived data and corresponding R scripts are publicly available here: LINK.
 
@@
 When examining reported against recalculated p-values for all inconsistencies (see @fig-scatter), we can identify certain spatial patterns (only inconsistencies that are based on '=' comparisons were included to make them easier to interpret, n = `{r} nrow(xdata[xdata$error == TRUE & xdata$p_comp == "=",])` out of `{r} nrow(xdata[xdata$error == TRUE,])`).
-First, the center of density of points is located in the bottom left corner with more p-values reported closer to the standard alpha level of 0.05.
+First, the center of density of points is located in the bottom left corner with more p-values reported closer to the most common alpha level of 0.05.
 This is not surprising, since the quantitative sciences exhibit a documented publication bias, with hypotheses being more often confirmed (and thus p \< .05) than not [@franco2014publication; @sterling1959publication].
@@
 The slope of the regression line is flatter than the diagonal axis which means that, on average, inconsistencies have a tendency to exhibit smaller reported p-values than their recalculated counterparts.
-Similarly, inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p > 0.05 or p < 0.05 ) are not equally prevalent. 
+Similarly, inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05 ) are not equally prevalent.
 There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference (e.g.
-p < 0.05).
+p \< 0.05).
 So even if we assumed these inconsistencies were merely typos of the comparison sign, inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.
@@
 ## 5.1. Statistical reporting inconsistencies are prevalent
-The present study found a large amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[9,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[9,3][[1]]` assessable p-values.
+
+The present study found a large amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[9,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[9,4][[1]]` assessable p-values.
 `{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`% of all p-values were flagged as inconsistent and `{r} 100 * round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`% were flagged as decision inconsistencies, i.e.
@@
 The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent such assessment using Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
-Previous studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% contain at least one decision consistency.
+The discovered rates fall in line with these studies who report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency.
 
@@
 If such a substantial amount of errors is found in plain site, the question naturally arises as to how many errors during the data analysis itself remain undetected.
-If the tip of the iceberg is already so large, what is the volume of the submerged iceberg?
+We should ask ourselves, if the tip of the iceberg is already so large, what is the volume of the submerged iceberg?
 
-The present examination assessment did not indicate any noticeable trend over the last 23 years of publications nor did the data suggest large differences between journals.
-
+The present examination did not indicate any noticeable trend over the last 23 years of publications.
 Observed inconsistencies were characterized by reported p-values being on average lower than their recalculated counterparts and the prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant.
-These patterns could indicate a systematic bias in favor of lower p-values in general and a bias towards significant results in particular.
-Our data do not speak to the causes of these biases, but possible reasons include the following: Researchers might intentionally round down p-values because they think a lower p-value is more convincing to reviewers and/or readers.
+These biases have been reported on for other disciplines as well [@nuijten2016prevalence; @bakker2011mis].
+These skewed patterns could indicate a systematic bias in favor of lower p-values in general and a bias towards significant results in particular.
+Our data do not speak to the causes of these biases, but possible reasons include the following: 
+Researchers might intentionally round down p-values because they think lower p-values are more convincing to reviewers and readers.
 This practice has been admitted to by 1 in 5 surveyed psychological researchers [@john2012measuring].
-Given that a non-trivial amount of quantitative linguists have admitted to commit questionable research practices (and even fraud) [@isbell2022misconduct], we cannot exclude the possibility that some of the inconsistencies were intentional.
+Given that a non-trivial amount of quantitative linguists have admitted to commit questionable research practices (and even fraud) [@isbell2022misconduct], we cannot exclude the possibility that some of the inconsistencies in our sample were indeed intentional.
 It is our strong belief, however, that the majority of inconsistencies are unintentional.
@@
 For example, @fugelsang2004theory let researchers evaluate data that are either consistent or inconsistent with their prior expectations.
-They showed that when researchers encounter results that disconfirm their expectations, they are likely to blame the methodology while results that confirmed their expectations were rarely critically scrutinized.
-Alternatively, the observed bias might be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than non-significant ones.
+They showed that when researchers encounter results that are not in line with their expectations, they are likely to blame the methodology while results that confirmed their expectations were rarely critically scrutinized.
+Alternatively, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than non-significant ones.
 
@@
 
-While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, there are, of course, a number of limitations.
-The present assessment and the conclusions we can draw from them are limited.
+While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, the present assessment and the conclusions we can draw from them are limited:
 First, our sample is limited to only a subset of experimental linguistic journals.
@@
 
-There are concrete actionable steps the field of experimental linguistics can make to reduce statistical reporting inconsistencies.
+There are concrete actionable steps the field of experimental linguistics can take to reduce statistical reporting inconsistencies.
 In order to avoid simple copy-and-paste errors related to working in two separate programs for writing the manuscript and conducting the statistical analysis, authors should consider 'literate programming', i.e.
@@
 
-Journal editors could explicitly recommend consistency checks with algorhythms such as statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines \[Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous\].
+Journal editors could explicitly recommend consistency checks with algorhythms such as Statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines (Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous).
 Editors could also demand, recommend or at least encourage data sharing for publication in their journal.
-Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g. @laurinavichyute2022share; @hardwicke2018data].
+Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share].
 

### 2024-07-25 manuscript/manuscript_anonymized.tex
anonymize

--- 2024-07-25 manuscript/manuscript.tex
@@
 
-\authorsnames{Dara Leonard Jenssen Etemady,Timo B. Roettger}
+\author{anonymous}
 
 \affiliation{
-{Department of Linguistics \& Scandinavian Studies, University of Oslo}}
-
-\leftheader{Etemady and Roettger}
+{anonymous}}
+
+\leftheader{anonymous}
 
@@
 reporting consistencies across articles in eight experimental
-linguistics journals, published between 2000 and 2023. Using the R
+linguistics journals published between 2000 and 2023. Using the R
 package ``Statcheck'', we retrieved 39532 statistical test and assessed
-whether p-values were inconsistent with their test statistic and degrees
+whether p-values were consistent with their test statistic and degrees
 of freedom. Half of all articles that used null hypothesis signficance
@@
 articles contained an inconsistency that may have affected the
-statistical conclusion. The inconsistency rates were stable across
-journals and publication years. We offer actionable steps for authors,
-reviewers, and editors to remmedy this state-of-affairs.}
-
-\keywords{statistics, statcheck, reproducibility}
-
-\authornote{\par{\addORCIDlink{Timo B. Roettger}{0000-0000-0000-0001}} 
+statistical conclusion. The inconsistency rates were comparable across
+journals and seem stable over publication years. We discuss possible
+reasons for this high rate and offer actionable steps for authors,
+reviewers, and editors to remedy this state-of-affairs.}
+
+\keywords{statistics, statcheck, reproducibility, null hypothesis
+significance testing}
+
+\authornote{ 
 \par{ }
-\par{   The authors have no conflict of interest to
-declare.   Conceptionalization, Methodology, Validation, Formal
-Analysis, Review \& Editing of Manuscript, Data Curation - T.B.R. \&
-D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft,
-Visualization, Supervision - T.B.R. }
-\par{Correspondence concerning this article should be addressed to Timo
-B. Roettger, Email: timo.roettger@gmail.com}
+\par{   The authors have no conflict of interest to declare.    }
+\par{Correspondence concerning this article should be addressed to }
 }
@@
 often informed by experimental data. Researchers test theoretical
-predictions with these data using statistical tests. Depending on the
-results of such tests, researchers make claims for or against
+predictions with their data using statistical tests. Depending on the
+results of their tests, researchers make claims for or against
 theoretical assumptions.\\
-Since they play such an integral part in the reasoning process of
-experimentalists, both the data these tests are based on and the
-computational procedure of the tests itself should be transparent in
-order for other researchers to critically evaluate them. Moreover things
-can go wrong. To err is human. Transparent sharing allows others to
-detect and correct human error. In the recent years, the quantitative
-sciences have seen repeated calls to become more transparent and
-reproducible by sharing data and statistical protocols
-(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
+Since these tests play such an integral part in the argumentation
+process of experimentalists, both the data these tests are based on and
+the computational procedure of the tests itself should be transparent in
+order for other researchers to critically evaluate them. But more than
+that, things can go wrong. Humans are fallible; they make mistakes.
+Transparent sharing allows others to detect and correct human error. In
+the recent years, quantitative linguistics have seen repeated calls to
+become more transparent and reproducible by sharing data and statistical
+protocols (\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
 \citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
@@
 
-Since data and statistical protocol sharing remains rare across the
-language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
-al., 2023}), interested readers are left with trusting the authors that
-the statistical analysis is run and reported correctly. However, the
-three sets of indices in (1) have clearly defined mathematical
-relationships and can thus be easily checked for consistency. An F test
-with the specified degrees of freedom and a test-statistic of 3.88
-should result in a p-value of 0.053 which is larger, not smaller, than
-0.05. Possible reasons for this inconsistency are manifold: It could be
-a typo of the comparison sign, i.e.~the authors meant to use \texttt{=}
-or \texttt{\textgreater{}} rather than \texttt{\textless{}}. Any of the
-numbers could contain a typo. Sometimes an error might indicate
-erroneous rounding (e.g.~0.057 being rounded down to 0.05). Without
-access to data and scripts, it remains unclear to the reader what has
-caused this inconsistency. Such inconsistencies can be particularly
-concerning if the calculated p-value (here 0.057) and the reported
-p-values (here \textless0.05) are not on the same side of the alpha
-threshold. In NHST, p-values below a conventionalized alpha threshold,
-most commonly 0.05, are interpreted as evidence that the data are
-sufficiently inconsistent with the null hypothesis (``significant'').
-P-values above that threshold are considered consistent with the null
-hypothesis and practically lead to rejecting the alternative hypothesis
+Without access to the data and scripts, interested readers are left with
+trusting the authors that the statistical analysis is run and reported
+correctly. However, the three sets of indices in (1) have clearly
+defined mathematical relationships and can thus be easily assessed for
+consistency. For example, an F test with the specified degrees of
+freedom and a test-statistic of 3.88 should result in a p-value of 0.053
+which is larger, not smaller, than 0.05. Possible reasons for this
+inconsistency are manifold: It could be a typo of the comparison sign,
+i.e.~the authors meant to use \texttt{=} or \texttt{\textless{}} rather
+than \texttt{\textgreater{}}. Additionally, any of the numbers could
+contain a typo and sometimes an error might indicate erroneous rounding
+(e.g.~0.057 being rounded down to 0.05). Without access to data and
+scripts, it remains unclear to the reader what has caused this
+inconsistency. Such inconsistencies can be particularly concerning if
+the calculated p-value (here 0.057) and the reported p-values (here
+\textless0.05) are not on the same side of the alpha threshold. In NHST,
+p-values below a conventionalized alpha threshold, most commonly 0.05,
+are interpreted as evidence that the data are sufficiently inconsistent
+with the null hypothesis (``significant''). P-values above that
+threshold are considered consistent with the null hypothesis and
+practically lead to rejecting the alternative hypothesis
 (``non-significant''). In (1) above, the reported p-value suggest a
@@
 the articles with statistical results contained at least one
-inconsistency and around 13\% contained at least one ``decision
-inconsistency''. Other studies report on inconsistency rates between 4\%
+inconsistency and around 13\% contained at least one decision
+inconsistency. Other studies report on inconsistency rates between 4\%
 and 14\%, with between 10\% and 63\% of articles containing at least one
-inconsistency and between 3\% and 21\% decision consistencies.
+inconsistency and between 3\% and 21\% contain at least one decision
+consistency. These high inconsistency rates in other disciplines are
+concerning and have led to a constructive dialog, resulting in
+recommendations for best practice to either avoid inconsistencies or
+make them more detectable.
 
@@
 between 2000 and 2023. We explore whether the rate of inconsistency
-differs across journals, whether that rate has changed over the course
-of the last 20 years and whether there is evidence for bias in these
+differs across journals, whether their rate has changed over the course
+of the last 23 years and whether there is evidence for bias in these
 statistical-reporting inconsistencies. We discuss the results and offer
@@
 
+\subsection{3.1. Data availability
+statement}\label{data-availability-statement}
+
 All quantitative analyses were conducted using R Core Team
-(\citeproc{ref-rmanual}{2021}) and the r packages {[}LIST ALL PACKAGES
-from 01\_Analysis.R and add to references{]}
-
-\subsection{3.1. Statcheck}\label{statcheck}
+(\citeproc{ref-rmanual}{2021}). All derived data and corresponding R
+scripts are publicly available here:
+\url{https://github.com/timo-b-roettger/ErrorsLing}
+
+\subsection{3.2. Statcheck}\label{statcheck}
 
@@
 plain text, Statcheck searches for specific strings that correspond to a
-NHST result using ``regular expressions''. That way, statcheck can
-detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation
-tests, and Q tests as long as the test result fulfills three conditions:
-(a) the test result is reported completely including the test statistic,
-the degrees of freedom (if applicable), and the p-value; (b) the test
-result is in the body of the text, i.e.~Statcheck usually misses
-reported in tables; and (c) the test result is reported in American
-Psychological Association style (\citeproc{ref-APA2020}{American
-Psychological Association, 2020}). Given these constraints, Statcheck is
-estimated to detect roughly 60\% of all reported NHST results
+NHST result using regular expressions. That way, Statcheck can detect
+results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q
+tests as long as the test result fulfills three conditions: (a) the test
+result is reported completely including the test statistic, the degrees
+of freedom (if applicable), and the p-value; (b) the test result is in
+the body of the text, i.e. Statcheck usually misses information in
+tables; and (c) the test result is reported in American Psychological
+Association style (\citeproc{ref-APA2020}{American Psychological
+Association, 2020}). Given these constraints, Statcheck is estimated to
+detect roughly 60\% of all reported NHST results
 (\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
 uses the reported test statistic and degrees of freedom to recalculate
-the p-value, compares the reported and recalculated p-values and, if
+the p-value, compares the reported and recalculated p-value and, if
 there is a mismatch, flags the test as containing an ``inconsistency.''
@@
 
-\subsection{3.2. Sample}\label{sample}
+\subsection{3.3. Sample}\label{sample}
 
@@
 to analysis.
-
-\subsection{3.3. Data availability
-statement}\label{data-availability-statement}
-
-All derived data and corresponding R scripts are publicly available
-here: LINK.
 
@@
 First, the center of density of points is located in the bottom left
-corner with more p-values reported closer to the standard alpha level of
-0.05. This is not surprising, since the quantitative sciences exhibit a
-documented publication bias, with hypotheses being more often confirmed
-(and thus p \textless{} .05) than not
+corner with more p-values reported closer to the most common alpha level
+of 0.05. This is not surprising, since the quantitative sciences exhibit
+a documented publication bias, with hypotheses being more often
+confirmed (and thus p \textless{} .05) than not
 (\citeproc{ref-franco2014publication}{Franco et al., 2014};
@@
 inconsistencies across a sample of 5804 experimental linguistic
-articles, containing 3059 assessable p-values. 13.1\% of all p-values
+articles, containing 39532 assessable p-values. 13.1\% of all p-values
 were flagged as inconsistent and 1.7\% were flagged as decision
@@
 \citeproc{ref-gross2021fidelity}{Groß, 2021};
-\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Previous
-studies report on inconsistency rates between 4\% and 14\%, with between
-10\% and 63\% of articles containing at least one inconsistency and
-between 3\% and 21\% contain at least one decision consistency.
+\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
+discovered rates fall in line with these studies who report on
+inconsistency rates between 4\% and 14\%, with between 10\% and 63\% of
+articles containing at least one inconsistency.
 
@@
 in plain site, the question naturally arises as to how many errors
-during the data analysis itself remain undetected. If the tip of the
-iceberg is already so large, what is the volume of the submerged
-iceberg?
-
-The present examination assessment did not indicate any noticeable trend
-over the last 23 years of publications nor did the data suggest large
-differences between journals.
-
-Observed inconsistencies were characterized by reported p-values being
-on average lower than their recalculated counterparts and the prevalence
-of decision inconsistencies was higher for p-values reported as
-significant than for those reported as non-significant. These patterns
-could indicate a systematic bias in favor of lower p-values in general
-and a bias towards significant results in particular. Our data do not
-speak to the causes of these biases, but possible reasons include the
-following: Researchers might intentionally round down p-values because
-they think a lower p-value is more convincing to reviewers and/or
-readers. This practice has been admitted to by 1 in 5 surveyed
-psychological researchers (\citeproc{ref-john2012measuring}{John et al.,
-2012}). Given that a non-trivial amount of quantitative linguists have
-admitted to commit questionable research practices (and even fraud)
+during the data analysis itself remain undetected. We should ask
+ourselves, if the tip of the iceberg is already so large, what is the
+volume of the submerged iceberg?
+
+The present examination did not indicate any noticeable trend over the
+last 23 years of publications. Observed inconsistencies were
+characterized by reported p-values being on average lower than their
+recalculated counterparts and the prevalence of decision inconsistencies
+was higher for p-values reported as significant than for those reported
+as non-significant. These biases have been reported on for other
+disciplines as well (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
+2011}; \citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}).
+These skewed patterns could indicate a systematic bias in favor of lower
+p-values in general and a bias towards significant results in
+particular. Our data do not speak to the causes of these biases, but
+possible reasons include the following: Researchers might intentionally
+round down p-values because they think lower p-values are more
+convincing to reviewers and readers. This practice has been admitted to
+by 1 in 5 surveyed psychological researchers
+(\citeproc{ref-john2012measuring}{John et al., 2012}). Given that a
+non-trivial amount of quantitative linguists have admitted to commit
+questionable research practices (and even fraud)
 (\citeproc{ref-isbell2022misconduct}{Isbell et al., 2022}), we cannot
-exclude the possibility that some of the inconsistencies were
-intentional. It is our strong belief, however, that the majority of
-inconsistencies are unintentional.
+exclude the possibility that some of the inconsistencies in our sample
+were indeed intentional. It is our strong belief, however, that the
+majority of inconsistencies are unintentional.
 
@@
 with their prior expectations. They showed that when researchers
-encounter results that disconfirm their expectations, they are likely to
-blame the methodology while results that confirmed their expectations
-were rarely critically scrutinized. Alternatively, the observed bias
-might be a reflection of publication bias
+encounter results that are not in line with their expectations, they are
+likely to blame the methodology while results that confirmed their
+expectations were rarely critically scrutinized. Alternatively, the
+observed bias might merely be a reflection of publication bias
 (\citeproc{ref-franco2014publication}{Franco et al., 2014};
@@
 While we believe our work offers an important contribution to improving
-statistical reporting practices in experimental linguistics, there are,
-of course, a number of limitations. The present assessment and the
-conclusions we can draw from them are limited. First, our sample is
-limited to only a subset of experimental linguistic journals. Our sample
-is based on a crude criterion of what constitutes an experimental
-linguistic journal (see \citeproc{ref-kobrock2023assessing}{Kobrock \&
-Roettger, 2023}) and we restricted ourselves further to (for us)
-accessible journals which explicitly require APA statistical formatting
-in the author guidelines. Thus it is possible that a different selection
-of journals would have resulted in different results. However, given
-that the inconsistency rates of our study are comparable to similar
-studies from other disciplines and that those rates are relatively
-stable across journals and time, our findings should be considered
-relevant for experimental linguistics at large.
+statistical reporting practices in experimental linguistics, the present
+assessment and the conclusions we can draw from them are limited: First,
+our sample is limited to only a subset of experimental linguistic
+journals. Our sample is based on a crude criterion of what constitutes
+an experimental linguistic journal (see
+\citeproc{ref-kobrock2023assessing}{Kobrock \& Roettger, 2023}) and we
+restricted ourselves further to (for us) accessible journals which
+explicitly require APA statistical formatting in the author guidelines.
+Thus it is possible that a different selection of journals would have
+resulted in different results. However, given that the inconsistency
+rates of our study are comparable to similar studies from other
+disciplines and that those rates are relatively stable across journals
+and time, our findings should be considered relevant for experimental
+linguistics at large.
 
@@
 There are concrete actionable steps the field of experimental
-linguistics can make to reduce statistical reporting inconsistencies. In
+linguistics can take to reduce statistical reporting inconsistencies. In
 order to avoid simple copy-and-paste errors related to working in two
@@
 Journal editors could explicitly recommend consistency checks with
-algorhythms such as statcheck during peer review, a practice that has
+algorhythms such as Statcheck during peer review, a practice that has
 been taken up on by several journals from neighboring disciplines
-{[}Psychological Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
+(Psychological Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
   accessed on July 15, 2024.}, Advances in Methods and Practices in
@@
   accessed on on July 15, 2024.}, Stress \& Health Barber
-(\citeproc{ref-barber2017meticulous}{2017}){]}. Editors could also
-demand, recommend or at least encourage data sharing for publication in
-their journal. Data sharing policies have been shown to substantially
-increase the reproducibility of analyses
-(\citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018}; e.g.
+(\citeproc{ref-barber2017meticulous}{2017})). Editors could also demand,
+recommend or at least encourage data sharing for publication in their
+journal. Data sharing policies have been shown to substantially increase
+the reproducibility of analyses (e.g.,
+\citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018};
 \citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022}).

### 2024-07-25 scripts/01_Analysis.R

@@ -1,22 +1,21 @@
 ## title: Statistical reporting inconsistencies in Ling
-## author: Timo B. Roettger
-## contact: timo.b.roettger@gmail.com   
+## author: anonymous
+## contact: anonymous   
 ## date created: 2024-07-11
-## date last edited:2024-07-11
+## date last edited: 2024-07-25
 
 
 ## packages ####################################################################                        
- 
 
 # load in relevant libraries, install if not installed
 # if (!require("pacman")) install.packages("pacman")
 # pacman::p_load(rstudioapi, 
 #                tidyverse, 
 #                ggpubr, 
-#                broom, 
 #                statcheck,
 #                janitor,
-#                patchwork
+#                patchwork,
+#                ggstream
 # )
 
 library(rstudioapi)
@@ -37,6 +36,7 @@ library(ggstream)
 # setwd(current_working_dir)
 
 ## list all journal units
+## only works with access to all article pdfs, skip if you don't have access
 files <- list.files("../Journals", recursive = TRUE, pattern = "\\.pdf")
 
 ## how many articles?
@@ -60,7 +60,6 @@ n_articles = length(files)
 
 ## statcheck ###################################################################                      
 
-
 ## read pdfs in /Journals with statcheck, checks for one tailed tests
 ## can only be run with access to original articles, bulk sharing is not prohibited by the journals
 #stat <- checkPDFdir("Journals", OneTailedTxt = TRUE)
@@ -70,7 +69,6 @@ n_articles = length(files)
 
 
 ## data ########################################################################                        
-
 
 ## load in data
 xdata <- read.csv("../data/statcheck_data.csv")
@@ -104,7 +102,6 @@ colnames(journal_summary_counts) <- c("Journal", "eligible articles", "assessabl
 # prevalence of errors for different comparison signs
 sign_distribution <- xtabs(~error + p_comp, xdata)
 sign_distribution_decision <- xtabs(~decision_error + p_comp, xdata)
-
 
 ## summary table proportions
 journal_summary_prop <- xdata %>% 
@@ -215,7 +212,6 @@ ggsave(filename = "../plots/figure1.png",
 
 ## figure 2 ####################################################################                        
 
-
[…]

### 2024-07-25 manuscript/manuscript.qmd
minor fixes

@@
   - name: Dara Leonard Jenssen Etemady
-    corresponding: false
+    corresponding: true
     email: daraetemady@gmail.com
@@
   - name: Timo B. Roettger
-    corresponding: true
+    corresponding: false
     orcid: 0000-0000-0000-0001
@@
 & D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft, Visualization, Supervision - T.B.R.
-abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals, published between 2000 and 2023.
-Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were inconsistent with their test statistic and degrees of freedom.
-Half of all articles that used null hypothesis signficance testing contained at least one inconsistent p-value.
+abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals published between 2000 and 2023.
+Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were consistent with their test statistic and degrees of freedom.
+Half of all articles that used null hypothesis significance testing contained at least one inconsistent p-value.
 One in eight articles contained an inconsistency that may have affected the statistical conclusion.
-The inconsistency rates were stable across journals and publication years.
-We offer actionable steps for authors, reviewers, and editors to remmedy this state-of-affairs."
-keywords: [statistics, statcheck, reproducibility]
+The inconsistency rates were comparable across journals and seem stable over publication years.
+We discuss possible reasons for this high rate and offer actionable steps for authors, reviewers, and editors to remedy this state-of-affairs."
+keywords: [statistics, statcheck, reproducibility, null hypothesis significance testing]
 bibliography: bibliography.bib

### 2024-07-25 manuscript/manuscript_anonymized.qmd
minor fixes

@@
 Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were consistent with their test statistic and degrees of freedom.
-Half of all articles that used null hypothesis signficance testing contained at least one inconsistent p-value.
+Half of all articles that used null hypothesis significance testing contained at least one inconsistent p-value.
 One in eight articles contained an inconsistency that may have affected the statistical conclusion.
@@
 
-All quantitative analyses were conducted using @rmanual.
-All derived data and corresponding R scripts are publicly available here: [https://github.com/timo-b-roettger/ErrorsLing](https://github.com/timo-b-roettger/ErrorsLing)
+All quantitative analyses were conducted using @rmanual. All derived data and corresponding R scripts are available [here](https://osf.io/gx3ub/?view_only=25957e06b5f449d6820b529eb0ae1937).
 

### 2024-07-29 manuscript/manuscript.tex
residual fixes

@@
 reporting consistencies across articles in eight experimental
-linguistics journals, published between 2000 and 2023. Using the R
+linguistics journals published between 2000 and 2023. Using the R
 package ``Statcheck'', we retrieved 39532 statistical test and assessed
-whether p-values were inconsistent with their test statistic and degrees
-of freedom. Half of all articles that used null hypothesis signficance
+whether p-values were consistent with their test statistic and degrees
+of freedom. Half of all articles that used null hypothesis significance
 testing contained at least one inconsistent p-value. One in eight
 articles contained an inconsistency that may have affected the
-statistical conclusion. The inconsistency rates were stable across
-journals and publication years. We offer actionable steps for authors,
-reviewers, and editors to remmedy this state-of-affairs.}
-
-\keywords{statistics, statcheck, reproducibility}
+statistical conclusion. The inconsistency rates were comparable across
+journals and seem stable over publication years. We discuss possible
+reasons for this high rate and offer actionable steps for authors,
+reviewers, and editors to remedy this state-of-affairs.}
+
+\keywords{statistics, statcheck, reproducibility, null hypothesis
+significance testing}
 
@@
 Visualization, Supervision - T.B.R. }
-\par{Correspondence concerning this article should be addressed to Timo
-B. Roettger, Email: timo.roettger@gmail.com}
+\par{Correspondence concerning this article should be addressed to Dara
+Leonard Jenssen Etemady, Email: daraetemady@gmail.com}
 }

### 2025-05-15 manuscript/manuscript_anonymized.qmd
Revisions and new analysis

@@
 abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals published between 2000 and 2023.
-Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were consistent with their test statistic and degrees of freedom.
-Half of all articles that used null hypothesis significance testing contained at least one inconsistent p-value.
-One in eight articles contained an inconsistency that may have affected the statistical conclusion.
+Using the R package “Statcheck”, we retrieved 85,442 statistical test from 13,065 articles and assessed whether p-values were consistent with their test statistic and degrees of freedom.
+Half of all articles (50%) that used null hypothesis significance testing contained at least one inconsistent p-value.
+More than one in eight articles (13%) contained an inconsistency that may have affected the statistical conclusion.
 The inconsistency rates were comparable across journals and seem stable over publication years.
 We discuss possible reasons for this high rate and offer actionable steps for authors, reviewers, and editors to remedy this state-of-affairs."
-keywords: [statistics, statcheck, reproducibility, null hypothesis significance testing]
+keywords: [statistics, Statcheck, reproducibility, null hypothesis significance testing]
 bibliography: bibliography.bib
@@
 Transparent sharing allows others to detect and correct human error.
-In the recent years, quantitative linguistics have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
+In recent years, quantitative linguistics have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols, often under the banner of "open science" [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
 Despite these calls, sharing of statistical protocols is still rather rare across the language sciences [@bochynska2023reproducible].
@@
 Recently, a series of studies used such automatic assessments to evaluate the prevalence of inconsistent statistical reporting in psychology [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness; @nuijten2016prevalence; @nuijten2020statcheck; @green2018statcheck; @claesen2023data], medical sciences [@garcia2004incongruence; @van2023comparing], psychiatry [@berle2007inconsistencies], cyber security studies [@gross2021fidelity], technological education research [@buckley2023estimating], and experimental philosophy [@colombo2018statistical].
-For example, looking at over 250 thousand p-values published in major psychology journals, @nuijten2016prevalence found that around 50% of the articles with statistical results contained at least one inconsistency and around 13% contained at least one decision inconsistency.
+For example, looking at over 250,000 p-values published in major psychology journals, @nuijten2016prevalence found that around 50% of the articles with statistical results contained at least one inconsistency and around 13% contained at least one decision inconsistency.
 Other studies report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency and between 3% and 21% contain at least one decision consistency.
@@
 
-To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in eight experimental linguistic journals published between 2000 and 2023.
+To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in thirteen experimental linguistic journals published between 2000 and 2023.
 We explore whether the rate of inconsistency differs across journals, whether their rate has changed over the course of the last 23 years and whether there is evidence for bias in these statistical-reporting inconsistencies.
@@
 
-## 3.2. Statcheck
+## 3.2. Sample
 
-We used the R package Statcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
+Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the search string "experiment\*" in title, abstract and/or keywords.
+Out of these 100 journals, we selected all journals with at least 10% of articles containing the search string "experiment\*".
+Out of the remaining 37 journals, we selected those journals that urged APA formatting either in the main body of the text or specifically regarding statistics in the author guidelines, resulting in nine remaining journals.
+Moreover, to access the articles in .pdf format, the articles had to be either accessible to us through our library license, or open access, resulting in a final list of eight journals: Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA).
+An anonymous reviewer raised the justified concern that the resulting sample might not represent the vast majority of work in experimental linguistics. 
+To address this concern, we additionally included those five journals that have published the highest absolute number of experimental articles, regardless of whether they explicitly urge APA formatting or not, resulting in the inclusion of the following psycholinguistic outlets: Journal of Memory and Language (JML), Language Cognition and Neuroscience (LCN, former Language and Cognitive Processes), Journal of Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing Research (SLH), and Brain and Language (BAL).
+
+We included only original research articles within the publication years of 2000-2023, excluding any book reviews, response articles, commentaries, editorials, corrigenda, errata, advertisements, etc.
+
+## 3.3. Statcheck
+
+We used the R package Statcheck [Version 1.5.0, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
 Statcheck works as follows: After converting articles in pdf or html format to plain text, Statcheck searches for specific strings that correspond to a NHST result using regular expressions.
@@
 
-## 3.3. Sample
+Articles from LAB spanned 2011-2023; LCN spanned 2015-2023.
+Statcheck could not parse 291 articles, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt, resulting in some gaps in the coverage.
+This procedure resulted in `{r} length(files)` research articles which were submitted to analysis.
 
-Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the search string "experiment\*" in title, abstract and/or keywords.
-Out of these 100 journals, we selected all journals with at least 10% of articles containing the search string "experiment\*".
-Out of the remaining 37 journals, we selected those journals that urged APA formatting either in the main body of the text or specifically regarding statistics in the author guidelines, resulting in nine remaining journals.
-Moreover, to access the articles in .pdf format, the articles had to be either accessible to us through our library license, or open access, resulting in a final list of eight journals: Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA).
-
-We included only original research articles within the publication years of 2000-2023, excluding any book reviews, response articles, commentaries, editorials, corrigenda, errata, advertisements, etc.
-Articles from LAB spanned 2011-2023, while the other journals spanned 2000-2023.
-Statcheck could not parse 157 articles, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt.
-No assessable statistics were found by Statcheck in JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
-This procedure resulted in `{r} length(files)` research articles which were submitted to analysis.
+## 3.4. Analysis and research questions
+The aims of this study were explicitly exploratory and hypothesis generating, thus analyses remain merely descriptive by reporting on the proportion of articles/tests that are statistically (in)consistent.
+Our investigation was set out to explore the following research questions:
+(1) How prevalent are statistical inconsistencies (and decision inconsistencies) in our sample?
+(2) Do inconsistency rates vary across journals and/or their publication year? 
+And (3) is there evidence for bias, i.e. are processes that result in inconsistencies more likely to produce lower p-values?
 
@@
 
-## 4.1 Prevalence of inconsistencies
+## 4.1 Inconsistencies are highly prevalent
 
 The results are summarized in @tbl-summary.
-Out of `{r} journal_summary_counts[9,2][[1]]` articles, `{r} journal_summary_counts[9,3][[1]]` articles contained statistical tests that statcheck could assess (`{r} 100 * round(journal_summary_counts[9,3][[1]] / journal_summary_counts[9,2][[1]],2)`%), amounting to `{r} journal_summary_counts[9,4][[1]]` assessable p-values.
-`{r} journal_summary_counts[9,5][[1]]` p-values were flagged as inconsistent (`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`%) out of which `{r} journal_summary_counts[9,6][[1]]` were considered decision inconsistencies (`{r} 100* round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`%) (see @tbl-summary).
+Out of `{r} journal_summary_counts[14,2][[1]]` articles, `{r} journal_summary_counts[14,3][[1]]` articles contained statistical tests that Statcheck could assess (`{r} 100 * round(journal_summary_counts[14,3][[1]] / journal_summary_counts[14,2][[1]],2)`%), amounting to `{r} journal_summary_counts[14,4][[1]]` assessable p-values.
+Interestingly, the five journals that did not explicitly encourage APA formatting had virtually identical rates of assessable articles compared to the original sample.
+Overall, `{r} journal_summary_counts[14,5][[1]]` p-values were flagged as inconsistent (`{r} 100 * round(journal_summary_counts[14,5][[1]] / journal_summary_counts[14,4][[1]],3)`%) out of which `{r} journal_summary_counts[14,6][[1]]` were considered decision inconsistencies (`{r} 100* round(journal_summary_counts[14,6][[1]] / journal_summary_counts[14,4][[1]],3)`%) (see @tbl-summary).
 
@@
 #| tbl-cap: "Number of eligible articles, assessable articles and results, inconsistencies and decision inconsistencies across all journals.
-(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA))"
+(Applied Psycholinguistics (APS), Language and Brain (BAL), Bilingualism: Language and Cognition (BLC), Journal of Memory and Language (JML), Journal of Psycholinguistic Research (JPR), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Cognition and Neuroscience (LCN, former Language and Cognitive Processes), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA), Journal Of Speech Language And Hearing Research (SLH))"
 
@@
 ```
+
+## 4.2 Inconsistencies are stable across journals and publication year
 
@@
 #| fig-cap: "Proportion of articles containing at least one inconsistency / decision inconsistency.
-(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA))"
+(Applied Psycholinguistics (APS), Language and Brain (BAL), Bilingualism: Language and Cognition (BLC), Journal of Memory and Language (JML), Journal of Psycholinguistic Research (JPR), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Cognition and Neuroscience (LCN, former Language and Cognitive Processes), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA), Journal Of Speech Language And Hearing Research (SLH))"
 
@@
 #| out-width: 100%
-#| fig-cap: "Proportion of inconsistencies / decision inconsistencies plotted from 2000 to 2023.
-Left: rates pooled across journals; Right: rates per journal.
-(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA))"
+#| fig-cap: "Proportion of inconsistencies / decision inconsistencies across time overall (left panel) and split into journals (right panel).
+(Applied Psycholinguistics (APS), Language and Brain (BAL), Bilingualism: Language and Cognition (BLC), Journal of Memory and Language (JML), Journal of Psycholinguistic Research (JPR), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Cognition and Neuroscience (LCN, former Language and Cognitive Processes), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA), Journal Of Speech Language And Hearing Research (SLH))"
 
@@
 
-When examining reported against recalculated p-values for all inconsistencies (see @fig-scatter), we can identify certain spatial patterns (only inconsistencies that are based on '=' comparisons were included to make them easier to interpret, n = `{r} nrow(xdata[xdata$error == TRUE & xdata$p_comp == "=",])` out of `{r} nrow(xdata[xdata$error == TRUE,])`).
-First, the center of density of points is located in the bottom left corner with more p-values reported closer to the most common alpha level of 0.05.
-This is not surprising, since the quantitative sciences exhibit a documented publication bias, with hypotheses being more often confirmed (and thus p \< .05) than not [@franco2014publication; @sterling1959publication].
-Second, there is a linear density of points along the diagonal axis which corresponds to numerically small inconsistencies, some of which might be related to simple rounding errors.
-However, comparing the diagonal to the black line, which represents a linear model predicting recalculated by reported p-values, we can see a clear divergence of what is expected if inconsistencies were equally likely in both directions.
-The slope of the regression line is flatter than the diagonal axis which means that, on average, inconsistencies have a tendency to exhibit smaller reported p-values than their recalculated counterparts.
-Similarly, inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05 ) are not equally prevalent.
-There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference (e.g.
-p \< 0.05).
-So even if we assumed these inconsistencies were merely typos of the comparison sign, inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.
+## 4.3 Inconsistencies are biased
+
+Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample:
+There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference.
+So even if we assumed these inconsistencies were merely typos of the comparison sign (e.g.
+\< instead of \>), inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.
+
 These biases are also reflected in decision inconsistencies.
-Of all decision inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant (p \> 0.05).
+Of all decision inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant (p \> 0.05), i.e.
+non-significant results are more than twice as likely to be erroneously reported as significant than the other way around.
+The latter pattern, however, seems to be a bias of the past. 
+Reproducing @nuijten2016prevalence, @fig-scatter plots the development of the bias observed for decision inconsistencies over time.
+The prevalence of decision inconsistencies in significant p-values seems to have slightly decreased over the years, while the prevalence of decision inconsistencies in non-significant p-values seems to have slightly increased over the years.
 
@@
 #| out-width: 100%
-#| fig-cap: "Reported vs.
-recalculated p-values for inconsistencies that are based on '=' comparions.
-Black line descriptively indicates the  linear relationship, indicating a bias towards lower reported p-values."
+#| fig-cap: "Percentage of decision inconsistencies falsely reporting significance (black) or non-significance (orange), plotted across publication years."
 
@@
 
-The present study found a large amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[9,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[9,4][[1]]` assessable p-values.
-`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`% of all p-values were flagged as inconsistent and `{r} 100 * round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`% were flagged as decision inconsistencies, i.e.
+The present study found a large amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[14,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[14,4][[1]]` assessable p-values.
+`{r} 100 * round(journal_summary_counts[14,5][[1]] / journal_summary_counts[14,4][[1]],3)`% of all p-values were flagged as inconsistent and `{r} 100 * round(journal_summary_counts[14,6][[1]] / journal_summary_counts[14,4][[1]],3)`% were flagged as decision inconsistencies, i.e.
 the reported p-value is on the opposite side of the alpha threshold than the recalculated p-value.
@@
 
-The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent such assessment using Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
-The discovered rates fall in line with these studies who report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency.
+The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent assessments using Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
+The discovered inconsistency rates fall in line with these studies that report on rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency.
 
@@
 
-The present examination did not indicate any noticeable trend over the last 23 years of publications.
+The present examination did not indicate any noticeable trend over a period of 23 years of publications.
 Observed inconsistencies were characterized by reported p-values being on average lower than their recalculated counterparts and the prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant.
@@
 Our data do not speak to the causes of these biases, but possible reasons include the following:
-Researchers might intentionally round down p-values because they think lower p-values are more convincing to reviewers and readers.
+
+First, researchers might intentionally round down p-values because they think lower p-values are more convincing to reviewers and readers.
 This practice has been admitted to by 1 in 5 surveyed psychological researchers [@john2012measuring].
@@
 
-Researchers might scrutinize non-significant results more than significant results or are less likely to double check significant results than non-significant results because results that confirm their hypothesis feed into their confirmation bias [@nickerson1998confirmation].
+Second, researchers might scrutinize non-significant results more than significant results or are less likely to double check significant results than non-significant results because results that confirm their hypothesis feed into their confirmation bias [@nickerson1998confirmation].
 For example, @fugelsang2004theory let researchers evaluate data that are either consistent or inconsistent with their prior expectations.
 They showed that when researchers encounter results that are not in line with their expectations, they are likely to blame the methodology while results that confirmed their expectations were rarely critically scrutinized.
-Alternatively, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than non-significant ones.
+
+Third, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than non-significant ones.
+Publication bias is a well established pattern in experimental linguistic research with many recent meta analyses discussing possible evidence for it [@isbilen2022statistical; @lehtonen2018bilingualism; @lu2024meta]
+For example, @de2015cognitive showed that studies with results supporting the bilingual-advantage theory were most likely to be published, while studies with results challenging the theory were significantly less likely to be published.
+
+Regardless of possible causes for bias in the processes that generate inconsistencies, the data also suggest that some of these biases has decreased over time, pointing to a positive development.
 
@@
 
-While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, the present assessment and the conclusions we can draw from them are limited:
+While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, the present assessment and the conclusions we can draw from them are limited.
 First, our sample is limited to only a subset of experimental linguistic journals. 
-Our sample is based on a crude criterion of what constitutes an experimental linguistic journal [see @kobrock2023assessing] and we restricted ourselves further to (for us) accessible journals which explicitly require APA statistical formatting in the author guidelines.
-Thus it is possible that a different selection of journals would have resulted in different results.
-However, given that the inconsistency rates of our study are comparable to similar studies from other disciplines and that those rates are relatively stable across journals and time, our findings should be considered relevant for experimental linguistics at large.
+However, given the selection of journals and their standing in the field, and given that the inconsistency rates of our study are not only comparable to similar studies from other disciplines but also relatively stable across journals and time, our findings should be considered relevant for experimental linguistics at large.
 
 Second, given the constraints on automatically detecting test statistics, Statcheck misses reported values that either diverge from APA reporting standards or are reported in tables.
-However, inconsistency rates have been shown to be similar for results in APA format vs. results that diverge from APA formatting [@bakker2011mis; @nuijten2016prevalence].
+However, inconsistency rates in our own sample have been shown to be similar for results in APA format vs. results that diverge from APA formatting [@bakker2011mis; @nuijten2016prevalence].
 
-Third, statcheck slightly overestimates inconsistency rates, because it might not accurately detect corrections for multiple comparisons [@schmidt2017statcheck].
+Third, Statcheck slightly overestimates inconsistency rates, because it might not accurately detect corrections for multiple comparisons [@schmidt2017statcheck].
 @nuijten2017validity, however, show that not only were there only a small proportion of flagged inconsistencies related to multiple comparisons, but also that these multiple comparisons themselves were often erroneously reported.
@@
 @kalmbach2023rule].
-Despite its limitations, Statcheck provides a rough proxy of true inconsistency rate in the published literature and we hope the reader agrees that the prevalence of inconsistencies is a state of affairs that should be reflected upon.
+Despite its limitations, Statcheck provides a rough proxy of true inconsistency rates in the published literature and we hope the reader agrees that the prevalence of inconsistencies is a state of affairs that should be reflected upon.
 
@@
 
-Authors should also consider sharing their derived data (i.e.
+Authors should generally engage in transparent and reproducible practices that can reduce human error or at least make them detectable by sharing their derived data (i.e.
 the anonymized data table that was analyzed) as well as a detailed description of their statistical protocol, ideally in form of reproducible scripts.
@@
 
-If authors do not share data and scripts, reviewers can check at least the statistical reporting consistency in the manuscript by using tools such as statcheck [@nuijten2023statcheck, http://statcheck.io] or p-checker [@schonbrodt2015p, http://shinyapps.org/apps/p-checker/].
+Reviewers can additionally check the statistical reporting consistency in the manuscript by using tools such as Statcheck [@nuijten2023statcheck, http://statcheck.io] or p-checker [@schonbrodt2015p, http://shinyapps.org/apps/p-checker/].
 Reviewers could consider requesting data and scripts during peer review.
@@
 Some of these approaches might be more appropriate than others [@sondereggera2024advancements; @vasishth2023some], so more thorough evaluations of how researchers arrive at their statistical conclusions might elevate their analytical robustness.
+Moreover, a turn towards inferential frameworks that do not focus on binary decision procedures such as the null hypothesis significance testing framework, might alleviate some of the biases we observed in the direction of inconsistencies [@vasishth2018statistical; @cumming2014new].
 
-Journal editors could explicitly recommend consistency checks with algorhythms such as Statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines (Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous).
+Journal editors could explicitly recommend consistency checks with algorithms such as Statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines (Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous).
 Editors could also demand, recommend or at least encourage data sharing for publication in their journal. 
-Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share].
+Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share] and a number of linguistic journals have already implemented such policies, including journals within our sample.
+Having said that, open data policies which for example the Journal of Memory and Language introduced in 2018 did not seem to have affected the proportion of statistical inconsistencies.
+The inconsistency rates for JML (and other journals) were rather stable across time, so open data alone might not resolve the issue without further changes to the research eco-system.
 

### 2025-05-15 manuscript/manuscript_anonymized.tex
minor fixes
Revisions and new analysis

@@
 linguistics journals published between 2000 and 2023. Using the R
-package ``Statcheck'', we retrieved 39532 statistical test and assessed
-whether p-values were consistent with their test statistic and degrees
-of freedom. Half of all articles that used null hypothesis signficance
-testing contained at least one inconsistent p-value. One in eight
-articles contained an inconsistency that may have affected the
-statistical conclusion. The inconsistency rates were comparable across
-journals and seem stable over publication years. We discuss possible
-reasons for this high rate and offer actionable steps for authors,
-reviewers, and editors to remedy this state-of-affairs.}
+package ``Statcheck'', we retrieved 85,442 statistical test from 13,065
+articles and assessed whether p-values were consistent with their test
+statistic and degrees of freedom. Half of all articles (50\%) that used
+null hypothesis significance testing contained at least one inconsistent
+p-value. More than one in eight articles (13\%) contained an
+inconsistency that may have affected the statistical conclusion. The
+inconsistency rates were comparable across journals and seem stable over
+publication years. We discuss possible reasons for this high rate and
+offer actionable steps for authors, reviewers, and editors to remedy
+this state-of-affairs.}
 
@@
 Transparent sharing allows others to detect and correct human error. In
-the recent years, quantitative linguistics have seen repeated calls to
+recent years, quantitative linguistics have seen repeated calls to
 become more transparent and reproducible by sharing data and statistical
-protocols (\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
+protocols, often under the banner of ``open science''
+(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
 \citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
@@
 experimental philosophy (\citeproc{ref-colombo2018statistical}{Colombo
-et al., 2018}). For example, looking at over 250 thousand p-values
-published in major psychology journals, Nuijten et al.
+et al., 2018}). For example, looking at over 250,000 p-values published
+in major psychology journals, Nuijten et al.
 (\citeproc{ref-nuijten2016prevalence}{2016}) found that around 50\% of
@@
 Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
-p-values reported in eight experimental linguistic journals published
+p-values reported in thirteen experimental linguistic journals published
 between 2000 and 2023. We explore whether the rate of inconsistency
@@
 All quantitative analyses were conducted using R Core Team
-(\citeproc{ref-rmanual}{2021}). All derived data and corresponding R
-scripts are publicly available here:
-\url{https://github.com/timo-b-roettger/ErrorsLing}
-
-\subsection{3.2. Statcheck}\label{statcheck}
-
-We used the R package Statcheck (Version 1.4.1-beta.2,
-\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
+(\citeproc{ref-rmanual}{2025}). All derived data and corresponding R
+scripts are available
+\href{https://osf.io/gx3ub/?view_only=25957e06b5f449d6820b529eb0ae1937}{here}.
+
+\subsection{3.2. Sample}\label{sample}
+
+Focusing on experimental linguistic research, we used Kobrock and
+Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
+departure who list 100 linguistic journals that had at least a hundred
+articles published at the time of assessment (2021) and a high ratio of
+articles containing the search string ``experiment*'' in title, abstract
+and/or keywords. Out of these 100 journals, we selected all journals
+with at least 10\% of articles containing the search string
+``experiment*''. Out of the remaining 37 journals, we selected those
+journals that urged APA formatting either in the main body of the text
+or specifically regarding statistics in the author guidelines, resulting
+in nine remaining journals. Moreover, to access the articles in .pdf
+format, the articles had to be either accessible to us through our
+library license, or open access, resulting in a final list of eight
+journals: Applied Psycholinguistics (APS), Bilingualism: Language and
+Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
+and Speech (LaS), Language Learning and Techology (LLT), Journal of
+Language and Social Psychology (LSP), Journal of Child Language (JCL),
+and Studies in Second Language Acquisition (SLA). An anonymous reviewer
+raised the justified concern that the resulting sample might not
+represent the vast majority of work in experimental linguistics. To
+address this concern, we additionally included those five journals that
+have published the highest absolute number of experimental articles,
+regardless of whether they explicitly urge APA formatting or not,
+resulting in the inclusion of the following psycholinguistic outlets:
+Journal of Memory and Language (JML), Language Cognition and
+Neuroscience (LCN, former Language and Cognitive Processes), Journal of
+Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing
+Research (SLH), and Brain and Language (BAL).
+
+We included only original research articles within the publication years
+of 2000-2023, excluding any book reviews, response articles,
+commentaries, editorials, corrigenda, errata, advertisements, etc.
+
+\subsection{3.3. Statcheck}\label{statcheck}
+
+We used the R package Statcheck (Version 1.5.0,
+\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2024}) to
 automatically detect statistical-reporting inconsistencies. Statcheck
@@
 
-\subsection{3.3. Sample}\label{sample}
-
-Focusing on experimental linguistic research, we used Kobrock and
-Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
-departure who list 100 linguistic journals that had at least a hundred
-articles published at the time of assessment (2021) and a high ratio of
-articles containing the search string ``experiment*'' in title, abstract
-and/or keywords. Out of these 100 journals, we selected all journals
-with at least 10\% of articles containing the search string
-``experiment*''. Out of the remaining 37 journals, we selected those
-journals that urged APA formatting either in the main body of the text
-or specifically regarding statistics in the author guidelines, resulting
-in nine remaining journals. Moreover, to access the articles in .pdf
-format, the articles had to be either accessible to us through our
-library license, or open access, resulting in a final list of eight
-journals: Applied Psycholinguistics (APS), Bilingualism: Language and
-Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
-and Speech (LaS), Language Learning and Techology (LLT), Journal of
-Language and Social Psychology (LSP), Journal of Child Language (JCL),
-and Studies in Second Language Acquisition (SLA).
-
-We included only original research articles within the publication years
-of 2000-2023, excluding any book reviews, response articles,
-commentaries, editorials, corrigenda, errata, advertisements, etc.
-Articles from LAB spanned 2011-2023, while the other journals spanned
-2000-2023. Statcheck could not parse 157 articles, likely related to
-issues with rendering the Chi-Squared symbol being erroneously converted
-from .pdf to .txt. No assessable statistics were found by Statcheck in
-JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
-This procedure resulted in 5804 research articles which were submitted
-to analysis.
+Articles from LAB spanned 2011-2023; LCN spanned 2015-2023. Statcheck
+could not parse 291 articles, likely related to issues with rendering
+the Chi-Squared symbol being erroneously converted from .pdf to .txt,
+resulting in some gaps in the coverage. This procedure resulted in 13065
+research articles which were submitted to analysis.
+
+\subsection{3.4. Analysis and research
+questions}\label{analysis-and-research-questions}
+
+The aims of this study were explicitly exploratory and hypothesis
+generating, thus analyses remain merely descriptive by reporting on the
+proportion of articles/tests that are statistically (in)consistent. Our
+investigation was set out to explore the following research questions:
+(1) How prevalent are statistical inconsistencies (and decision
+inconsistencies) in our sample? (2) Do inconsistency rates vary across
+journals and/or their publication year? And (3) is there evidence for
+bias, i.e.~are processes that result in inconsistencies more likely to
+produce lower p-values?
 
@@
 
-\subsection{4.1 Prevalence of
-inconsistencies}\label{prevalence-of-inconsistencies}
-
-The results are summarized in Table~\ref{tbl-summary}. Out of 5804
-articles, 3059 articles contained statistical tests that statcheck could
-assess (53\%), amounting to 39532 assessable p-values. 5166 p-values
-were flagged as inconsistent (13.1\%) out of which 655 were considered
-decision inconsistencies (1.7\%) (see Table~\ref{tbl-summary}).
+\subsection{4.1 Inconsistencies are highly
+prevalent}\label{inconsistencies-are-highly-prevalent}
+
+The results are summarized in Table~\ref{tbl-summary}. Out of 13065
+articles, 6484 articles contained statistical tests that Statcheck could
+assess (50\%), amounting to 85442 assessable p-values. Interestingly,
+the five journals that did not explicitly encourage APA formatting had
+virtually identical rates of assessable articles compared to the
+original sample. Overall, 11279 p-values were flagged as inconsistent
+(13.2\%) out of which 1442 were considered decision inconsistencies
+(1.7\%) (see Table~\ref{tbl-summary}).
 
@@
 inconsistencies and decision inconsistencies across all journals.
-(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition
-(BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech
-(LaS), Language Learning and Techology (LLT), Journal of Language and
-Social Psychology (LSP), Journal of Child Language (JCL), and Studies in
-Second Language Acquisition (SLA))}{\label{tbl-summary}}}}
+(Applied Psycholinguistics (APS), Language and Brain (BAL),
+Bilingualism: Language and Cognition (BLC), Journal of Memory and
+Language (JML), Journal of Psycholinguistic Research (JPR), Linguistic
+Approaches to Bilingualism (LAB), Language and Speech (LaS), Language
+Cognition and Neuroscience (LCN, former Language and Cognitive
+Processes), Language Learning and Techology (LLT), Journal of Language
+and Social Psychology (LSP), Journal of Child Language (JCL), and
+Studies in Second Language Acquisition (SLA), Journal Of Speech Language
+And Hearing Research (SLH))}{\label{tbl-summary}}}}
 
@@
 \endlastfoot
-APS & 953 & 690 & 9570 & 1368 & 170 \\
-BLC & 964 & 610 & 9093 & 1161 & 120 \\
-JCL & 1109 & 529 & 6240 & 750 & 69 \\
-LAB & 471 & 133 & 1719 & 234 & 25 \\
-LLT & 421 & 111 & 919 & 201 & 61 \\
-LSP & 695 & 376 & 4320 & 429 & 60 \\
-LaS & 598 & 363 & 4717 & 552 & 86 \\
-SLA & 593 & 247 & 2954 & 471 & 64 \\
-Total & 5804 & 3059 & 39532 & 5166 & 655 \\
+APS & 953 & 690 & 9570 & 1447 & 198 \\
+BAL & 2253 & 780 & 10344 & 1524 & 198 \\
+BLC & 964 & 610 & 9093 & 1223 & 134 \\
+JCL & 1109 & 529 & 6240 & 779 & 76 \\
+JML & 1507 & 768 & 13236 & 1004 & 132 \\
+JPR & 1137 & 534 & 5576 & 848 & 97 \\
+LAB & 471 & 133 & 1719 & 239 & 25 \\
+LCN & 751 & 397 & 5979 & 909 & 95 \\
+LLT & 421 & 111 & 919 & 210 & 64 \\
+LSP & 695 & 376 & 4320 & 449 & 68 \\
+LaS & 597 & 362 & 4708 & 564 & 88 \\
+SLA & 593 & 247 & 2954 & 476 & 66 \\
+SLH & 1614 & 947 & 10784 & 1607 & 201 \\
+Total & 13065 & 6484 & 85442 & 11279 & 1442 \\
 \end{longtable}
@@
 
-The proportion of inconsistencies ranged from 10 to 22\% across journals
+\subsection{4.2 Inconsistencies are stable across journals and
+publication
+year}\label{inconsistencies-are-stable-across-journals-and-publication-year}
+
+The proportion of inconsistencies ranged from 8 to 23\% across journals
 (1 to 7\% for decision inconsistencies) (see
@@
 of assessable articles contained one or more inconsistencies (journals
-range from 43 to 56\%) and 13\% contained one or more decision
-inconsistencies (journals range from 11 to 15\%).
+range from 45 to 56\%) and 13\% contained one or more decision
+inconsistencies (journals range from 11 to 17\%).
 
@@
 least one inconsistency / decision inconsistency. (Applied
-Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
-Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
-Language Learning and Techology (LLT), Journal of Language and Social
-Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
-Language Acquisition (SLA))}
+Psycholinguistics (APS), Language and Brain (BAL), Bilingualism:
+Language and Cognition (BLC), Journal of Memory and Language (JML),
+Journal of Psycholinguistic Research (JPR), Linguistic Approaches to
+Bilingualism (LAB), Language and Speech (LaS), Language Cognition and
+Neuroscience (LCN, former Language and Cognitive Processes), Language
+Learning and Techology (LLT), Journal of Language and Social Psychology
+(LSP), Journal of Child Language (JCL), and Studies in Second Language
+Acquisition (SLA), Journal Of Speech Language And Hearing Research
+(SLH))}
 
@@
 \caption{\label{fig-year-prop}Proportion of inconsistencies / decision
-inconsistencies plotted from 2000 to 2023. Left: rates pooled across
-journals; Right: rates per journal. (Applied Psycholinguistics (APS),
-Bilingualism: Language and Cognition (BLC), Linguistic Approaches to
-Bilingualism (LAB), Language and Speech (LaS), Language Learning and
-Techology (LLT), Journal of Language and Social Psychology (LSP),
-Journal of Child Language (JCL), and Studies in Second Language
-Acquisition (SLA))}
+inconsistencies across time overall (left panel) and split into journals
+(right panel). (Applied Psycholinguistics (APS), Language and Brain
+(BAL), Bilingualism: Language and Cognition (BLC), Journal of Memory and
+Language (JML), Journal of Psycholinguistic Research (JPR), Linguistic
+Approaches to Bilingualism (LAB), Language and Speech (LaS), Language
+Cognition and Neuroscience (LCN, former Language and Cognitive
+Processes), Language Learning and Techology (LLT), Journal of Language
+and Social Psychology (LSP), Journal of Child Language (JCL), and
+Studies in Second Language Acquisition (SLA), Journal Of Speech Language
+And Hearing Research (SLH))}
 
@@
 
-When examining reported against recalculated p-values for all
-inconsistencies (see Figure~\ref{fig-scatter}), we can identify certain
-spatial patterns (only inconsistencies that are based on `=' comparisons
-were included to make them easier to interpret, n = 3779 out of 5166).
-First, the center of density of points is located in the bottom left
-corner with more p-values reported closer to the most common alpha level
-of 0.05. This is not surprising, since the quantitative sciences exhibit
-a documented publication bias, with hypotheses being more often
-confirmed (and thus p \textless{} .05) than not
-(\citeproc{ref-franco2014publication}{Franco et al., 2014};
-\citeproc{ref-sterling1959publication}{Sterling, 1959}). Second, there
-is a linear density of points along the diagonal axis which corresponds
-to numerically small inconsistencies, some of which might be related to
-simple rounding errors. However, comparing the diagonal to the black
-line, which represents a linear model predicting recalculated by
-reported p-values, we can see a clear divergence of what is expected if
-inconsistencies were equally likely in both directions. The slope of the
-regression line is flatter than the diagonal axis which means that, on
-average, inconsistencies have a tendency to exhibit smaller reported
-p-values than their recalculated counterparts. Similarly,
-inconsistencies that report the p-value as being larger or smaller than
-a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05 )
-are not equally prevalent. There were 4.4\% of inconsistencies with p
-being reported as larger than a reference but 7.4\% inconsistencies with
-p being reported as smaller than a reference (e.g.~p \textless{} 0.05).
+\subsection{4.3 Inconsistencies are
+biased}\label{inconsistencies-are-biased}
+
+Inconsistencies that report the p-value as being larger or smaller than
+a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05,
+respectively) are not equally prevalent in the sample: There were 4.4\%
+of inconsistencies with p being reported as larger than a reference but
+7.1\% inconsistencies with p being reported as smaller than a reference.
 So even if we assumed these inconsistencies were merely typos of the
-comparison sign, inconsistencies that erroneously report the p-value to
-be smaller than a reference value are more frequent than inconsistencies
-that erroneously report the p-value to be larger than a reference value.
+comparison sign (e.g.~\textless{} instead of \textgreater),
+inconsistencies that erroneously report the p-value to be smaller than a
+reference value are more frequent than inconsistencies that erroneously
+report the p-value to be larger than a reference value.
+
 These biases are also reflected in decision inconsistencies. Of all
-decision inconsistencies (n = 655), 71\% represent cases in which a
+decision inconsistencies (n = 1442), 72\% represent cases in which a
 reported significant result (p \textless{} 0.05) is recalculated as
-non-significant (p \textgreater{} 0.05).
+non-significant (p \textgreater{} 0.05), i.e.~non-significant results
+are more than twice as likely to be erroneously reported as significant
+than the other way around. The latter pattern, however, seems to be a
+bias of the past. Reproducing Nuijten et al.
+(\citeproc{ref-nuijten2016prevalence}{2016}), Figure~\ref{fig-scatter}
+plots the development of the bias observed for decision inconsistencies
+over time. The prevalence of decision inconsistencies in significant
+p-values seems to have slightly decreased over the years, while the
+prevalence of decision inconsistencies in non-significant p-values seems
+to have slightly increased over the years.
 
@@
 
-\caption{\label{fig-scatter}Reported vs.~recalculated p-values for
-inconsistencies that are based on `=' comparions. Black line
-descriptively indicates the linear relationship, indicating a bias
-towards lower reported p-values.}
+\caption{\label{fig-scatter}Percentage of decision inconsistencies
+falsely reporting significance (black) or non-significance (orange),
+plotted across publication years.}
 
@@
 The present study found a large amount of statistical reporting
-inconsistencies across a sample of 5804 experimental linguistic
-articles, containing 39532 assessable p-values. 13.1\% of all p-values
+inconsistencies across a sample of 13065 experimental linguistic
+articles, containing 85442 assessable p-values. 13.2\% of all p-values
 were flagged as inconsistent and 1.7\% were flagged as decision
@@
 \citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
-recent such assessment using Statcheck
+recent assessments using Statcheck
 (\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
@@
 \citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
-discovered rates fall in line with these studies who report on
-inconsistency rates between 4\% and 14\%, with between 10\% and 63\% of
+discovered inconsistency rates fall in line with these studies that
+report on rates between 4\% and 14\%, with between 10\% and 63\% of
 articles containing at least one inconsistency.
@@
 
-The present examination did not indicate any noticeable trend over the
-last 23 years of publications. Observed inconsistencies were
+The present examination did not indicate any noticeable trend over a
+period of 23 years of publications. Observed inconsistencies were
 characterized by reported p-values being on average lower than their
@@
 particular. Our data do not speak to the causes of these biases, but
-possible reasons include the following: Researchers might intentionally
-round down p-values because they think lower p-values are more
-convincing to reviewers and readers. This practice has been admitted to
-by 1 in 5 surveyed psychological researchers
-(\citeproc{ref-john2012measuring}{John et al., 2012}). Given that a
-non-trivial amount of quantitative linguists have admitted to commit
-questionable research practices (and even fraud)
+possible reasons include the following:
+
+First, researchers might intentionally round down p-values because they
+think lower p-values are more convincing to reviewers and readers. This
+practice has been admitted to by 1 in 5 surveyed psychological
+researchers (\citeproc{ref-john2012measuring}{John et al., 2012}). Given
+that a non-trivial amount of quantitative linguists have admitted to
+commit questionable research practices (and even fraud)
 (\citeproc{ref-isbell2022misconduct}{Isbell et al., 2022}), we cannot
@@
 
-Researchers might scrutinize non-significant results more than
+Second, researchers might scrutinize non-significant results more than
 significant results or are less likely to double check significant
@@
 likely to blame the methodology while results that confirmed their
-expectations were rarely critically scrutinized. Alternatively, the
-observed bias might merely be a reflection of publication bias
-(\citeproc{ref-franco2014publication}{Franco et al., 2014};
+expectations were rarely critically scrutinized.
+
+Third, the observed bias might merely be a reflection of publication
+bias (\citeproc{ref-franco2014publication}{Franco et al., 2014};
 \citeproc{ref-sterling1959publication}{Sterling, 1959}) with
 (erroneously) reported significant p-values being more likely to be
-published than non-significant ones.
+published than non-significant ones. Publication bias is a well
+established pattern in experimental linguistic research with many recent
+meta analyses discussing possible evidence for it
+(\citeproc{ref-isbilen2022statistical}{Isbilen \& Christiansen, 2022};
+\citeproc{ref-lehtonen2018bilingualism}{Lehtonen et al., 2018};
+\citeproc{ref-lu2024meta}{Lu et al., 2024}) For example, De Bruin et al.
+(\citeproc{ref-de2015cognitive}{2015}) showed that studies with results
+supporting the bilingual-advantage theory were most likely to be
+published, while studies with results challenging the theory were
+significantly less likely to be published.
+
+Regardless of possible causes for bias in the processes that generate
+inconsistencies, the data also suggest that some of these biases has
+decreased over time, pointing to a positive development.
 
@@
 statistical reporting practices in experimental linguistics, the present
-assessment and the conclusions we can draw from them are limited: First,
+assessment and the conclusions we can draw from them are limited. First,
 our sample is limited to only a subset of experimental linguistic
-journals. Our sample is based on a crude criterion of what constitutes
-an experimental linguistic journal (see
-\citeproc{ref-kobrock2023assessing}{Kobrock \& Roettger, 2023}) and we
-restricted ourselves further to (for us) accessible journals which
-explicitly require APA statistical formatting in the author guidelines.
-Thus it is possible that a different selection of journals would have
-resulted in different results. However, given that the inconsistency
-rates of our study are comparable to similar studies from other
-disciplines and that those rates are relatively stable across journals
-and time, our findings should be considered relevant for experimental
-linguistics at large.
+journals. However, given the selection of journals and their standing in
+the field, and given that the inconsistency rates of our study are not
+only comparable to similar studies from other disciplines but also
+relatively stable across journals and time, our findings should be
+considered relevant for experimental linguistics at large.
 
@@
 APA reporting standards or are reported in tables. However,
-inconsistency rates have been shown to be similar for results in APA
-format vs.~results that diverge from APA formatting
+inconsistency rates in our own sample have been shown to be similar for
+results in APA format vs.~results that diverge from APA formatting
 (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011};
@@
 
-Third, statcheck slightly overestimates inconsistency rates, because it
+Third, Statcheck slightly overestimates inconsistency rates, because it
 might not accurately detect corrections for multiple comparisons
@@
 \citeproc{ref-kalmbach2023rule}{Kalmbach et al., 2023}). Despite its
-limitations, Statcheck provides a rough proxy of true inconsistency rate
-in the published literature and we hope the reader agrees that the
+limitations, Statcheck provides a rough proxy of true inconsistency
+rates in the published literature and we hope the reader agrees that the
 prevalence of inconsistencies is a state of affairs that should be
@@
 
-Authors should also consider sharing their derived data (i.e.~the
-anonymized data table that was analyzed) as well as a detailed
-description of their statistical protocol, ideally in form of
-reproducible scripts. Sharing reproducible analyses with reviewers
-allows the reviewers the reproduce the authors' analyses, possibly
-detect errors or even inappropriate statistical choices before
-publication, thus improving the quality and robustness of the final
-product. Moreover, publicly sharing their analyses has numerous benefits
-to the authors themselves beyond error detection: Open data and
-materials can facilitate collaboration
+Authors should generally engage in transparent and reproducible
+practices that can reduce human error or at least make them detectable
+by sharing their derived data (i.e.~the anonymized data table that was
+analyzed) as well as a detailed description of their statistical
+protocol, ideally in form of reproducible scripts. Sharing reproducible
+analyses with reviewers allows the reviewers the reproduce the authors'
+analyses, possibly detect errors or even inappropriate statistical
+choices before publication, thus improving the quality and robustness of
+the final product. Moreover, publicly sharing their analyses has
+numerous benefits to the authors themselves beyond error detection: Open
+data and materials can facilitate collaboration
 (\citeproc{ref-boland2017ten}{Boland et al., 2017}), increase efficiency
@@
 
-If authors do not share data and scripts, reviewers can check at least
-the statistical reporting consistency in the manuscript by using tools
-such as statcheck (\citeproc{ref-nuijten2023statcheck}{Nuijten \&
-Epskamp, 2023}, http://statcheck.io) or p-checker
+Reviewers can additionally check the statistical reporting consistency
+in the manuscript by using tools such as Statcheck
+(\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2024},
+http://statcheck.io) or p-checker
 (\citeproc{ref-schonbrodt2015p}{Schönbrodt, 2015},
@@
 more thorough evaluations of how researchers arrive at their statistical
-conclusions might elevate their analytical robustness.
+conclusions might elevate their analytical robustness. Moreover, a turn
+towards inferential frameworks that do not focus on binary decision
+procedures such as the null hypothesis significance testing framework,
+might alleviate some of the biases we observed in the direction of
+inconsistencies (\citeproc{ref-cumming2014new}{Cumming, 2014};
+\citeproc{ref-vasishth2018statistical}{Vasishth et al., 2018}).
 
 Journal editors could explicitly recommend consistency checks with
-algorhythms such as Statcheck during peer review, a practice that has
+algorithms such as Statcheck during peer review, a practice that has
 been taken up on by several journals from neighboring disciplines
@@
 \citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018};
-\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022}).
+\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022})
+and a number of linguistic journals have already implemented such
+policies, including journals within our sample. Having said that, open
+data policies which for example the Journal of Memory and Language
+introduced in 2018 did not seem to have affected the proportion of
+statistical inconsistencies. The inconsistency rates for JML (and other
+journals) were rather stable across time, so open data alone might not
+resolve the issue without further changes to the research eco-system.
 
@@
 
+\bibitem[\citeproctext]{ref-cumming2014new}
+Cumming, G. (2014). The new statistics: Why and how. \emph{Psychological
+Science}, \emph{25}(1), 7--29.
+
+\bibitem[\citeproctext]{ref-de2015cognitive}
+De Bruin, A., Treccani, B., \& Della Sala, S. (2015). Cognitive
+advantage in bilingualism: An example of publication bias?
+\emph{Psychological Science}, \emph{26}(1), 99--107.
+
 \bibitem[\citeproctext]{ref-franco2014publication}
@@
 
+\bibitem[\citeproctext]{ref-isbilen2022statistical}
+Isbilen, E. S., \& Christiansen, M. H. (2022). Statistical learning of
+language: A meta-analysis into 25 years of research. \emph{Cognitive
+Science}, \emph{46}(9), e13198.
+
 \bibitem[\citeproctext]{ref-john2012measuring}
@@
 
+\bibitem[\citeproctext]{ref-lehtonen2018bilingualism}
+Lehtonen, M., Soveri, A., Laine, A., Järvenpää, J., De Bruin, A., \&
+Antfolk, J. (2018). Is bilingualism associated with enhanced executive
+functioning in adults? A meta-analytic review. \emph{Psychological
+Bulletin}, \emph{144}(4), 394.
+
 \bibitem[\citeproctext]{ref-lowndes2017our}
@@
 
+\bibitem[\citeproctext]{ref-lu2024meta}
+Lu, J., Frank, M., Degen, J., Lu, J., Frank, M. C., \& Degen, J. (2024).
+A meta-analysis of syntactic satiation in extraction from islands.
+\emph{Glossa Psycholinguistics}, \emph{3}(1).
+
 \bibitem[\citeproctext]{ref-nickerson1998confirmation}
@@
 \bibitem[\citeproctext]{ref-nuijten2023statcheck}
-Nuijten, M. B., \& Epskamp, S. (2023). \emph{Statcheck: Extract
-statistics from articles and recompute p-values(1.4. 1-beta. 2){[}r{]}}.
+Nuijten, M. B., \& Epskamp, S. (2024). \emph{Statcheck: Extract
+statistics from articles and recompute p-values(1.5.0){[}r{]}}.
 
@@
 \bibitem[\citeproctext]{ref-rmanual}
-R Core Team. (2021). \emph{R: A language and environment for statistical
+R Core Team. (2025). \emph{R: A language and environment for statistical
 computing}. R Foundation for Statistical Computing.
@@
 
+\bibitem[\citeproctext]{ref-vasishth2018statistical}
+Vasishth, S., Mertzen, D., Jäger, L. A., \& Gelman, A. (2018). The
+statistical significance filter leads to overoptimistic expectations of
+replicability. \emph{Journal of Memory and Language}, \emph{103},
+151--175.
+
 \bibitem[\citeproctext]{ref-veldkamp2014statistical}

### 2025-05-15 scripts/01_Analysis.R

@@ -32,16 +32,16 @@ library(ggstream)
 
 ## load in data from location of script 
 ## (uncomment if you run script outside of sourcing it through manuscript.qmd)
-# current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
-# setwd(current_working_dir)
+#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
+#setwd(current_working_dir)
 
 ## list all journal units
 ## only works with access to all article pdfs, skip if you don't have access
-files <- list.files("../Journals", recursive = TRUE, pattern = "\\.pdf")
+files <- list.files("../Journals", recursive = TRUE, pattern = "\\.txt")
 
 ## how many articles?
 length(files)
-# 5804
+# 13065
 
 ## summary table
 files_wrangled <- as.data.frame(files) %>%
@@ -65,13 +65,13 @@ n_articles = length(files)
 #stat <- checkPDFdir("Journals", OneTailedTxt = TRUE)
 
 ## write statcheck data to csv
-#write.csv(stat, "../data/statcheck_data.csv", row.names = FALSE)
+#write.csv(stat, "../data/statcheck_data_second_round.csv", row.names = FALSE)
 
 
 ## data ########################################################################                        
 
 ## load in data
-xdata <- read.csv("../data/statcheck_data.csv")
+xdata <- read_csv("../data/statcheck_revised_sample.csv")
 
 ## split strings into year and journal, drop NAs
 xdata <- xdata %>% 
@@ -80,6 +80,11 @@ xdata <- xdata %>%
   drop_na(reported_p) %>%
   drop_na(computed_p) %>%
   filter(year != 2024)
+
+## remove LCN before before 2015 
+xdata <- xdata |> 
+  filter(! (journal == "LCN" & year < 2015))
+
 
 ## summary tables ##############################################################                       
 
@@ -98,6 +103,14 @@ journal_summary_counts <- xdata %>%
 
 colnames(journal_summary_counts) <- c("Journal", "eligible articles", "assessable articles",
                                       "assessable results", "inconsistencies", "decision inconsistencies")
+
+# less accessibility in journals without explicit APA guidelines?
+journal_summary_counts |> 
+  mutate(sample = ifelse(Journal %in% c("JML", "LCN", "JPR", "SLH", "BAL"), "new", "old"),
+         accessibility = `assessable articles` / `eligible articles`) |> 
+  group_by(sample) |> 
+  summarise(mean_assess = mean(accessibility))
[…]

### 2025-05-20 manuscript/manuscript.qmd
ready for resubmission

@@
 shorttitle: "Statistical Inconsistencies"
+number-sections: true
 author:
@@
     corresponding: false
-    orcid: 0000-0000-0000-0001
-    email: timo.roettger@gmail.com
+    orcid: 0000-0003-1400-2739
+    email: timo.roettger@iln.uio.no
     affiliations:
@@
     gratitude: null
-    authorship-agreements: Conceptionalization, Methodology, Validation, Formal Analysis, Review & Editing of Manuscript, Data Curation - T.B.R.
-& D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft, Visualization, Supervision - T.B.R.
-abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals published between 2000 and 2023.
-Using the R package “Statcheck”, we retrieved 39532 statistical test and assessed whether p-values were consistent with their test statistic and degrees of freedom.
-Half of all articles that used null hypothesis significance testing contained at least one inconsistent p-value.
-One in eight articles contained an inconsistency that may have affected the statistical conclusion.
+    authorship-agreements: Conceptionalization, Methodology, Validation, Formal Analysis, Review & Editing of Manuscript, Data Curation - TBR.
+& DLJE; Software, Investigation - DLJE; Writing of Original Draft, Visualization, Supervision - TBR.
+abstract: "The present paper investigates the prevalence of statistical reporting inconsistencies across articles in thirteen experimental linguistics journals published between 2000 and 2023.
+Using the R package “Statcheck”, we retrieved 85,442 statistical test from 13,065 articles and assessed whether p-values were consistent with their test statistic and degrees of freedom.
+Half of all articles (50%) that used null hypothesis significance testing contained at least one inconsistent p-value.
+More than one in eight articles (13%) contained an inconsistency that may have affected the statistical conclusion.
 The inconsistency rates were comparable across journals and seem stable over publication years.
 We discuss possible reasons for this high rate and offer actionable steps for authors, reviewers, and editors to remedy this state-of-affairs."
-keywords: [statistics, statcheck, reproducibility, null hypothesis significance testing]
+keywords: [statistics, Statcheck, reproducibility, null hypothesis significance testing]
 bibliography: bibliography.bib
@@
 
-# 1. Introduction
+# Introduction
 
@@
 Depending on the results of their tests, researchers make claims for or against theoretical assumptions.\
-Since these tests play such an integral part in the argumentation process of experimentalists, both the data these tests are based on and the computational procedure of the tests itself should be transparent in order for other researchers to critically evaluate them.
-But more than that, things can go wrong.
-Humans are fallible; they make mistakes.
+Since these tests play such an integral part in the argumentation process of experimentalists, both the data these tests are based on and the computational procedure of the tests itself should be error-free.
+But humans are fallible; they make mistakes.
+We cannot avoid making errors, but we can make them at least detectable.
 Transparent sharing allows others to detect and correct human error.
-In the recent years, quantitative linguistics have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
+
+In recent years, quantitative linguistics have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols, often under the banner of "open science" [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
 Despite these calls, sharing of statistical protocols is still rather rare across the language sciences [@bochynska2023reproducible].
@@
 whether a hypothesis is rejected or accepted, these errors might lead to -at best- overconfident, -at worst- false theoretical conclusions.
-The present paper will present evidence that the published literature in experimental linguistics contains a concerning amount of such statistical errors, a state of affairs which warrants more rigorous data sharing practices.
+The present paper will present evidence that the published literature in experimental linguistics contains a concerning amount of such statistical inconsitencies, a state of affairs which warrants more rigorous data sharing practices.
 
-# 2. Statistal reporting inconsistencies
+# Statistal reporting inconsistencies
 
@@
 
-To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in eight experimental linguistic journals published between 2000 and 2023.
-We explore whether the rate of inconsistency differs across journals, whether their rate has changed over the course of the last 23 years and whether there is evidence for bias in these statistical-reporting inconsistencies.
+To assess the prevalence of statistical reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in thirteen experimental linguistic journals published between 2000 and 2023.
+We explore whether the inconsistency rates differ across journals, whether they have changed over the course of the last 23 years and whether there is evidence for bias in these inconsistencies.
 We discuss the results and offer concrete recommendations for authors, reviewers, and editors to tackle this problem.
@@
 
-# 3. Method
+# Method
 
-## 3.1. Data availability statement
+## Data availability statement
 
-All quantitative analyses were conducted using @rmanual.
-All derived data and corresponding R scripts are publicly available here: [https://github.com/timo-b-roettger/ErrorsLing](https://github.com/timo-b-roettger/ErrorsLing)
+All quantitative analyses were conducted using @rmanual. All derived data and corresponding R scripts are available [here](https://osf.io/gx3ub/?view_only=25957e06b5f449d6820b529eb0ae1937).
 
-## 3.2. Statcheck
+## Sample
 
-We used the R package Statcheck [Version 1.4.1-beta.2, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
+Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the search string "experiment\*" in title, abstract and/or keywords.
+Out of these 100 journals, we selected all journals with at least 10% of articles containing the search string "experiment\*".
+Out of the remaining 37 journals, we selected those journals that urged APA formatting either in the main body of the text or specifically regarding statistics in the author guidelines, resulting in nine remaining journals.
+Moreover, to access the articles in .pdf format, the articles had to be either accessible to us through our library license, or open access, resulting in a final list of eight journals: Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA).
+An anonymous reviewer raised the justified concern that the resulting sample might not represent the vast majority of work in experimental linguistics. 
+To address this concern, we additionally included those five journals that have published the highest absolute number of experimental articles [according to @kobrock2023assessing], regardless of whether they explicitly urge APA formatting or not, resulting in the inclusion of the following psycholinguistic outlets: Journal of Memory and Language (JML), Language, Cognition and Neuroscience (LCN, former Language and Cognitive Processes), Journal of Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing Research (SLH), and Brain and Language (BAL).
+
+We included only original research articles within the publication years of 2000-2023, excluding any book reviews, response articles, commentaries, editorials, corrigenda, errata, advertisements, etc.
+
+## Statcheck
+
+We used the R package Statcheck [Version 1.5.0, @nuijten2023statcheck] to automatically detect statistical reporting inconsistencies.
 Statcheck works as follows: After converting articles in pdf or html format to plain text, Statcheck searches for specific strings that correspond to a NHST result using regular expressions.
-That way, Statcheck can detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, the degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
+That way, Statcheck can detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q-tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, the degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
 Statcheck usually misses information in tables; and (c) the test result is reported in American Psychological Association style [@APA2020].
@@
 
-## 3.3. Sample
-
-Focusing on experimental linguistic research, we used @kobrock2023assessing as a point of departure who list 100 linguistic journals that had at least a hundred articles published at the time of assessment (2021) and a high ratio of articles containing the search string "experiment\*" in title, abstract and/or keywords.
-Out of these 100 journals, we selected all journals with at least 10% of articles containing the search string "experiment\*".
-Out of the remaining 37 journals, we selected those journals that urged APA formatting either in the main body of the text or specifically regarding statistics in the author guidelines, resulting in nine remaining journals.
-Moreover, to access the articles in .pdf format, the articles had to be either accessible to us through our library license, or open access, resulting in a final list of eight journals: Applied Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS), Language Learning and Techology (LLT), Journal of Language and Social Psychology (LSP), Journal of Child Language (JCL), and Studies in Second Language Acquisition (SLA).
-
-We included only original research articles within the publication years of 2000-2023, excluding any book reviews, response articles, commentaries, editorials, corrigenda, errata, advertisements, etc.
-Articles from LAB spanned 2011-2023, while the other journals spanned 2000-2023.
-Statcheck could not parse 157 articles, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt.
-No assessable statistics were found by Statcheck in JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
+Articles from Linguistic Approaches to Bilingualism spanned 2011-2023; articles from Language, Cognition and Neuroscience spanned 2015-2023.
+Statcheck could not parse 291 articles, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt, resulting in some gaps in the coverage.
 This procedure resulted in `{r} length(files)` research articles which were submitted to analysis.
 
-# 4. Results
+## Analysis and research questions
+The aims of this study were explicitly exploratory and hypothesis generating, thus analyses remain merely descriptive by reporting on the proportion of articles/tests that are statistically (in)consistent.
+Our investigation was set out to explore the following research questions:
+(1) How prevalent are statistical inconsistencies (and decision inconsistencies) in our sample? (see @sec-prevalent)
+(2) Do inconsistency rates vary across journals and/or their publication year? (see @sec-stable)
+And (3) is there evidence for bias, i.e. are processes that result in inconsistencies more likely to produce lower p-values? (see @sec-bias)
 
-## 4.1 Prevalence of inconsistencies
+# Results
+
+## Inconsistencies are highly prevalent {#sec-prevalent}
 
 The results are summarized in @tbl-summary.
-Out of `{r} journal_summary_counts[9,2][[1]]` articles, `{r} journal_summary_counts[9,3][[1]]` articles contained statistical tests that statcheck could assess (`{r} 100 * round(journal_summary_counts[9,3][[1]] / journal_summary_counts[9,2][[1]],2)`%), amounting to `{r} journal_summary_counts[9,4][[1]]` assessable p-values.
-`{r} journal_summary_counts[9,5][[1]]` p-values were flagged as inconsistent (`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`%) out of which `{r} journal_summary_counts[9,6][[1]]` were considered decision inconsistencies (`{r} 100* round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`%) (see @tbl-summary).
+Out of `{r} journal_summary_counts[14,2][[1]]` articles, `{r} journal_summary_counts[14,3][[1]]` articles contained statistical tests that Statcheck could assess (`{r} 100 * round(journal_summary_counts[14,3][[1]] / journal_summary_counts[14,2][[1]],2)`%), amounting to `{r} journal_summary_counts[14,4][[1]]` assessable p-values.
+Interestingly, the five journals that did not explicitly encourage APA formatting had virtually identical rates of assessable articles compared to the original sample.
+Overall, `{r} journal_summary_counts[14,5][[1]]` p-values were flagged as inconsistent (`{r} 100 * round(journal_summary_counts[14,5][[1]] / journal_summary_counts[14,4][[1]],3)`%) out of which `{r} journal_summary_counts[14,6][[1]]` were considered decision inconsistencies (`{r} 100* round(journal_summary_counts[14,6][[1]] / journal_summary_counts[14,4][[1]],3)`%) (see @tbl-summary).
 
@@
 ```
+
+## Inconsistencies are stable across journals and publication year {#sec-stable}
 
@@
 
-When examining reported against recalculated p-values for all inconsistencies (see @fig-scatter), we can identify certain spatial patterns (only inconsistencies that are based on '=' comparisons were included to make them easier to interpret, n = `{r} nrow(xdata[xdata$error == TRUE & xdata$p_comp == "=",])` out of `{r} nrow(xdata[xdata$error == TRUE,])`).
-First, the center of density of points is located in the bottom left corner with more p-values reported closer to the most common alpha level of 0.05.
-This is not surprising, since the quantitative sciences exhibit a documented publication bias, with hypotheses being more often confirmed (and thus p \< .05) than not [@franco2014publication; @sterling1959publication].
-Second, there is a linear density of points along the diagonal axis which corresponds to numerically small inconsistencies, some of which might be related to simple rounding errors.
-However, comparing the diagonal to the black line, which represents a linear model predicting recalculated by reported p-values, we can see a clear divergence of what is expected if inconsistencies were equally likely in both directions.
-The slope of the regression line is flatter than the diagonal axis which means that, on average, inconsistencies have a tendency to exhibit smaller reported p-values than their recalculated counterparts.
-Similarly, inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05 ) are not equally prevalent.
-There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference (e.g.
-p \< 0.05).
-So even if we assumed these inconsistencies were merely typos of the comparison sign, inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.
+## Inconsistencies appear biased but bias decreases over time {#sec-bias}
+
+If inconsistencies were bias-free, we would expect different types of inconsistencies to be equally frequent. 
+However, this is not the case.
+Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample:
+There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference.
+So even if we assumed these inconsistencies were merely typos of the comparison sign (e.g.
+\< instead of \>), inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.
+
 These biases are also reflected in decision inconsistencies.
-Of all decision inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant (p \> 0.05).
+Of all decision inconsistencies (n = `{r} nrow(xdata[xdata$decision_error == TRUE,])`), `{r} 100 * round(nrow(xdata[xdata$decision_error == TRUE & xdata$computed_p >= .05,]) / nrow(xdata[xdata$decision_error == TRUE,]),2)`% represent cases in which a reported significant result (p \< 0.05) is recalculated as non-significant (p \> 0.05), i.e.
+non-significant results are more than twice as likely to be erroneously reported as significant than the other way around.
+The latter pattern, however, seems to have decreased over time. 
+Reproducing @nuijten2016prevalence, @fig-scatter plots the development of the bias observed for decision inconsistencies over time.
+The prevalence of decision inconsistencies in significant p-values seems to have slightly decreased over the years, while the prevalence of decision inconsistencies in non-significant p-values seems to have slightly increased over the years.
 
@@
 #| label: fig-scatter
-#| out-width: 100%
-#| fig-cap: "Reported vs.
-recalculated p-values for inconsistencies that are based on '=' comparions.
-Black line descriptively indicates the  linear relationship, indicating a bias towards lower reported p-values."
+#| out-width: 80%
+#| fig-cap: "Percentage of decision inconsistencies falsely reporting significance (black) or non-significance (orange), plotted across publication years."
 
@@
 
-# 5. Discussion and Recommendations
+# Discussion and Recommendations
 
-## 5.1. Statistical reporting inconsistencies are prevalent
+## Statistical reporting inconsistencies are prevalent
 
-The present study found a large amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[9,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[9,4][[1]]` assessable p-values.
-`{r} 100 * round(journal_summary_counts[9,5][[1]] / journal_summary_counts[9,4][[1]],3)`% of all p-values were flagged as inconsistent and `{r} 100 * round(journal_summary_counts[9,6][[1]] / journal_summary_counts[9,4][[1]],3)`% were flagged as decision inconsistencies, i.e.
+The present study found a large amount of statistical reporting inconsistencies across a sample of `{r} journal_summary_counts[14,2][[1]]` experimental linguistic articles, containing `{r} journal_summary_counts[14,4][[1]]` assessable p-values.
+`{r} 100 * round(journal_summary_counts[14,5][[1]] / journal_summary_counts[14,4][[1]],3)`% of all p-values were flagged as inconsistent and `{r} 100 * round(journal_summary_counts[14,6][[1]] / journal_summary_counts[14,4][[1]],3)`% were flagged as decision inconsistencies, i.e.
 the reported p-value is on the opposite side of the alpha threshold than the recalculated p-value.
 On average, `{r} 100*round(gross_per_article[[1]],2)`% of assessable articles contained at least one inconsistency and `{r} 100*round(gross_per_article[[2]],2)`% contained at least one decision inconsistency.
+The present examination did not indicate any noticeable trend for inconsistency rates across journals or publication year.
 
-The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent such assessment using Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
-The discovered rates fall in line with these studies who report on inconsistency rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency.
+The present study can be considered a conceptual replication of previous studies investigating statistical reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent assessments using the automatic tool Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
+The discovered inconsistency rates fall in line with these studies that report on rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency. 
+Moreover, the observed rates of inconsistencies and decision inconsistencies are virtually identical to rates reported on by @nuijten2016prevalence on the psychological science literature.
 
@@
 
-The present examination did not indicate any noticeable trend over the last 23 years of publications.
-Observed inconsistencies were characterized by reported p-values being on average lower than their recalculated counterparts and the prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant.
+Our results suggest biases as well. 
+The prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant.
 These biases have been reported on for other disciplines as well [@nuijten2016prevalence; @bakker2011mis].
@@
 
-Researchers might scrutinize non-significant results more than significant results or are less likely to double check significant results than non-significant results because results that confirm their hypothesis feed into their confirmation bias [@nickerson1998confirmation].
+Second, researchers might scrutinize non-significant results more than significant results or are less likely to double check significant results than non-significant results because results that confirm their hypothesis feed into their confirmation bias [@nickerson1998confirmation].
 For example, @fugelsang2004theory let researchers evaluate data that are either consistent or inconsistent with their prior expectations.
 They showed that when researchers encounter results that are not in line with their expectations, they are likely to blame the methodology while results that confirmed their expectations were rarely critically scrutinized.
-Alternatively, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than non-significant ones.
 
-## 5.2. Limitations of our study
+Third, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than  (erroneously) reported non-significant ones.
+Publication bias is a well established pattern in experimental linguistic research with many recent meta analyses discussing possible evidence for it [@isbilen2022statistical; @lehtonen2018bilingualism; @lu2024meta]
+For example, @de2015cognitive showed that studies with results supporting the bilingual-advantage theory were most likely to be published, while studies with results challenging the theory were significantly less likely to be published.
 
-While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, the present assessment and the conclusions we can draw from them are limited:
+Regardless of what possibly causes biases in the processes that generate inconsistencies, our data also suggest a positive development. 
+The over proportional occurrence of decision inconsistencies for p-values reported as significant has decreased over time.
+
+## Limitations of our study
+
+While we believe our work offers an important contribution to improving statistical reporting practices in experimental linguistics, the present assessment and the conclusions we can draw from them are limited.
 First, our sample is limited to only a subset of experimental linguistic journals. 
-Our sample is based on a crude criterion of what constitutes an experimental linguistic journal [see @kobrock2023assessing] and we restricted ourselves further to (for us) accessible journals which explicitly require APA statistical formatting in the author guidelines.
-Thus it is possible that a different selection of journals would have resulted in different results.
-However, given that the inconsistency rates of our study are comparable to similar studies from other disciplines and that those rates are relatively stable across journals and time, our findings should be considered relevant for experimental linguistics at large.
+However, given the selection of journals and their standing in the field, and given that the inconsistency rates of our study are not only comparable to similar studies from other disciplines but also relatively stable across journals and time, our findings should be considered relevant for experimental linguistics at large.
 
 Second, given the constraints on automatically detecting test statistics, Statcheck misses reported values that either diverge from APA reporting standards or are reported in tables.
-However, inconsistency rates have been shown to be similar for results in APA format vs. results that diverge from APA formatting [@bakker2011mis; @nuijten2016prevalence].
+However, inconsistency rates in our own sample have been shown to be similar for results in APA format vs. results that diverge from APA formatting [@bakker2011mis; @nuijten2016prevalence].
 
-Third, statcheck slightly overestimates inconsistency rates, because it might not accurately detect corrections for multiple comparisons [@schmidt2017statcheck].
+Third, Statcheck slightly overestimates inconsistency rates, because it might not accurately detect corrections for multiple comparisons [@schmidt2017statcheck].
 @nuijten2017validity, however, show that not only were there only a small proportion of flagged inconsistencies related to multiple comparisons, but also that these multiple comparisons themselves were often erroneously reported.
@@
 @kalmbach2023rule].
-Despite its limitations, Statcheck provides a rough proxy of true inconsistency rate in the published literature and we hope the reader agrees that the prevalence of inconsistencies is a state of affairs that should be reflected upon.
+Despite its limitations, Statcheck provides a rough proxy of true inconsistency rates in the published literature and we hope the reader agrees that the prevalence of inconsistencies is a state of affairs that should be reflected upon.
 
-## 5.3. Recommendations for the field
+## Recommendations for the field
 
@@
 Some of these approaches might be more appropriate than others [@sondereggera2024advancements; @vasishth2023some], so more thorough evaluations of how researchers arrive at their statistical conclusions might elevate their analytical robustness.
+Moreover, a turn towards inferential frameworks that do not focus on binary decision procedures such as the null hypothesis significance testing framework, might alleviate some of the biases we observed in the direction of inconsistencies [@vasishth2018statistical; @cumming2014new].
 
-Journal editors could explicitly recommend consistency checks with algorhythms such as Statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines (Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous).
+Journal editors could explicitly recommend consistency checks with algorithms such as Statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines (Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous).
 Editors could also demand, recommend or at least encourage data sharing for publication in their journal. 
-Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share].
+Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share] and a number of linguistic journals have already implemented such policies, including journals within our sample.
+Having said that, open data policies which for example the Journal of Memory and Language introduced in 2018 did not seem to have affected the proportion of statistical inconsistencies after their introduction.
+The inconsistency rates for JML (and other journals) were rather stable across time, so open data alone might not resolve the issue without further changes to the research eco-system.
 

### 2025-05-20 manuscript/manuscript.tex
ready for resubmission

@@
 \abstract{The present paper investigates the prevalence of statistical
-reporting consistencies across articles in eight experimental
+reporting inconsistencies across articles in thirteen experimental
 linguistics journals published between 2000 and 2023. Using the R
-package ``Statcheck'', we retrieved 39532 statistical test and assessed
-whether p-values were consistent with their test statistic and degrees
-of freedom. Half of all articles that used null hypothesis significance
-testing contained at least one inconsistent p-value. One in eight
-articles contained an inconsistency that may have affected the
-statistical conclusion. The inconsistency rates were comparable across
-journals and seem stable over publication years. We discuss possible
-reasons for this high rate and offer actionable steps for authors,
-reviewers, and editors to remedy this state-of-affairs.}
-
-\keywords{statistics, statcheck, reproducibility, null hypothesis
+package ``Statcheck'', we retrieved 85,442 statistical test from 13,065
+articles and assessed whether p-values were consistent with their test
+statistic and degrees of freedom. Half of all articles (50\%) that used
+null hypothesis significance testing contained at least one inconsistent
+p-value. More than one in eight articles (13\%) contained an
+inconsistency that may have affected the statistical conclusion. The
+inconsistency rates were comparable across journals and seem stable over
+publication years. We discuss possible reasons for this high rate and
+offer actionable steps for authors, reviewers, and editors to remedy
+this state-of-affairs.}
+
+\keywords{statistics, Statcheck, reproducibility, null hypothesis
 significance testing}
 
-\authornote{\par{\addORCIDlink{Timo B. Roettger}{0000-0000-0000-0001}} 
+\authornote{\par{\addORCIDlink{Timo B. Roettger}{0000-0003-1400-2739}} 
 \par{ }
@@
 declare.   Conceptionalization, Methodology, Validation, Formal
-Analysis, Review \& Editing of Manuscript, Data Curation - T.B.R. \&
-D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft,
-Visualization, Supervision - T.B.R. }
+Analysis, Review \& Editing of Manuscript, Data Curation - TBR. \& DLJE;
+Software, Investigation - DLJE; Writing of Original Draft,
+Visualization, Supervision - TBR. }
 \par{Correspondence concerning this article should be addressed to Dara
@@
 
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 
@@
 
-\section{1. Introduction}\label{introduction}
+\section{Introduction}\label{introduction}
 
@@
 process of experimentalists, both the data these tests are based on and
-the computational procedure of the tests itself should be transparent in
-order for other researchers to critically evaluate them. But more than
-that, things can go wrong. Humans are fallible; they make mistakes.
-Transparent sharing allows others to detect and correct human error. In
-the recent years, quantitative linguistics have seen repeated calls to
+the computational procedure of the tests itself should be error-free.
+But humans are fallible; they make mistakes. We cannot avoid making
+errors, but we can make them at least detectable. Transparent sharing
+allows others to detect and correct human error.
+
+In recent years, quantitative linguistics have seen repeated calls to
 become more transparent and reproducible by sharing data and statistical
-protocols (\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
+protocols, often under the banner of ``open science''
+(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
 \citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
@@
 that the published literature in experimental linguistics contains a
-concerning amount of such statistical errors, a state of affairs which
-warrants more rigorous data sharing practices.
-
-\section{2. Statistal reporting
+concerning amount of such statistical inconsitencies, a state of affairs
+which warrants more rigorous data sharing practices.
+
+\section{Statistal reporting
 inconsistencies}\label{statistal-reporting-inconsistencies}
@@
 
-To assess the prevalence of statistical-reporting inconsistencies in
+To assess the prevalence of statistical reporting inconsistencies in
 experimental linguistics, the present paper conceptually replicates
 Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
-p-values reported in eight experimental linguistic journals published
-between 2000 and 2023. We explore whether the rate of inconsistency
-differs across journals, whether their rate has changed over the course
-of the last 23 years and whether there is evidence for bias in these
-statistical-reporting inconsistencies. We discuss the results and offer
-concrete recommendations for authors, reviewers, and editors to tackle
-this problem.
-
-\section{3. Method}\label{method}
-
-\subsection{3.1. Data availability
+p-values reported in thirteen experimental linguistic journals published
+between 2000 and 2023. We explore whether the inconsistency rates differ
+across journals, whether they have changed over the course of the last
+23 years and whether there is evidence for bias in these
+inconsistencies. We discuss the results and offer concrete
+recommendations for authors, reviewers, and editors to tackle this
+problem.
+
+\section{Method}\label{method}
+
+\subsection{Data availability
 statement}\label{data-availability-statement}
@@
 All quantitative analyses were conducted using R Core Team
-(\citeproc{ref-rmanual}{2021}). All derived data and corresponding R
-scripts are publicly available here:
-\url{https://github.com/timo-b-roettger/ErrorsLing}
-
-\subsection{3.2. Statcheck}\label{statcheck}
-
-We used the R package Statcheck (Version 1.4.1-beta.2,
-\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
-automatically detect statistical-reporting inconsistencies. Statcheck
+(\citeproc{ref-rmanual}{2025}). All derived data and corresponding R
+scripts are available
+\href{https://osf.io/gx3ub/?view_only=25957e06b5f449d6820b529eb0ae1937}{here}.
+
+\subsection{Sample}\label{sample}
+
+Focusing on experimental linguistic research, we used Kobrock and
+Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
+departure who list 100 linguistic journals that had at least a hundred
+articles published at the time of assessment (2021) and a high ratio of
+articles containing the search string ``experiment*'' in title, abstract
+and/or keywords. Out of these 100 journals, we selected all journals
+with at least 10\% of articles containing the search string
+``experiment*''. Out of the remaining 37 journals, we selected those
+journals that urged APA formatting either in the main body of the text
+or specifically regarding statistics in the author guidelines, resulting
+in nine remaining journals. Moreover, to access the articles in .pdf
+format, the articles had to be either accessible to us through our
+library license, or open access, resulting in a final list of eight
+journals: Applied Psycholinguistics (APS), Bilingualism: Language and
+Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
+and Speech (LaS), Language Learning and Techology (LLT), Journal of
+Language and Social Psychology (LSP), Journal of Child Language (JCL),
+and Studies in Second Language Acquisition (SLA). An anonymous reviewer
+raised the justified concern that the resulting sample might not
+represent the vast majority of work in experimental linguistics. To
+address this concern, we additionally included those five journals that
+have published the highest absolute number of experimental articles
+(according to \citeproc{ref-kobrock2023assessing}{Kobrock \& Roettger,
+2023}), regardless of whether they explicitly urge APA formatting or
+not, resulting in the inclusion of the following psycholinguistic
+outlets: Journal of Memory and Language (JML), Language, Cognition and
+Neuroscience (LCN, former Language and Cognitive Processes), Journal of
+Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing
+Research (SLH), and Brain and Language (BAL).
+
+We included only original research articles within the publication years
+of 2000-2023, excluding any book reviews, response articles,
+commentaries, editorials, corrigenda, errata, advertisements, etc.
+
+\subsection{Statcheck}\label{statcheck}
+
+We used the R package Statcheck (Version 1.5.0,
+\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2024}) to
+automatically detect statistical reporting inconsistencies. Statcheck
 works as follows: After converting articles in pdf or html format to
@@
 NHST result using regular expressions. That way, Statcheck can detect
-results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q
-tests as long as the test result fulfills three conditions: (a) the test
-result is reported completely including the test statistic, the degrees
-of freedom (if applicable), and the p-value; (b) the test result is in
-the body of the text, i.e. Statcheck usually misses information in
+results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and
+Q-tests as long as the test result fulfills three conditions: (a) the
+test result is reported completely including the test statistic, the
+degrees of freedom (if applicable), and the p-value; (b) the test result
+is in the body of the text, i.e. Statcheck usually misses information in
 tables; and (c) the test result is reported in American Psychological
@@
 
-\subsection{3.3. Sample}\label{sample}
-
-Focusing on experimental linguistic research, we used Kobrock and
-Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
-departure who list 100 linguistic journals that had at least a hundred
-articles published at the time of assessment (2021) and a high ratio of
-articles containing the search string ``experiment*'' in title, abstract
-and/or keywords. Out of these 100 journals, we selected all journals
-with at least 10\% of articles containing the search string
-``experiment*''. Out of the remaining 37 journals, we selected those
-journals that urged APA formatting either in the main body of the text
-or specifically regarding statistics in the author guidelines, resulting
-in nine remaining journals. Moreover, to access the articles in .pdf
-format, the articles had to be either accessible to us through our
-library license, or open access, resulting in a final list of eight
-journals: Applied Psycholinguistics (APS), Bilingualism: Language and
-Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
-and Speech (LaS), Language Learning and Techology (LLT), Journal of
-Language and Social Psychology (LSP), Journal of Child Language (JCL),
-and Studies in Second Language Acquisition (SLA).
-
-We included only original research articles within the publication years
-of 2000-2023, excluding any book reviews, response articles,
-commentaries, editorials, corrigenda, errata, advertisements, etc.
-Articles from LAB spanned 2011-2023, while the other journals spanned
-2000-2023. Statcheck could not parse 157 articles, likely related to
-issues with rendering the Chi-Squared symbol being erroneously converted
-from .pdf to .txt. No assessable statistics were found by Statcheck in
-JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
-This procedure resulted in 5804 research articles which were submitted
-to analysis.
-
-\section{4. Results}\label{results}
-
-\subsection{4.1 Prevalence of
-inconsistencies}\label{prevalence-of-inconsistencies}
-
-The results are summarized in Table~\ref{tbl-summary}. Out of 5804
-articles, 3059 articles contained statistical tests that statcheck could
-assess (53\%), amounting to 39532 assessable p-values. 5166 p-values
-were flagged as inconsistent (13.1\%) out of which 655 were considered
-decision inconsistencies (1.7\%) (see Table~\ref{tbl-summary}).
+Articles from Linguistic Approaches to Bilingualism spanned 2011-2023;
+articles from Language, Cognition and Neuroscience spanned 2015-2023.
+Statcheck could not parse 291 articles, likely related to issues with
+rendering the Chi-Squared symbol being erroneously converted from .pdf
+to .txt, resulting in some gaps in the coverage. This procedure resulted
+in 13065 research articles which were submitted to analysis.
+
+\subsection{Analysis and research
+questions}\label{analysis-and-research-questions}
+
+The aims of this study were explicitly exploratory and hypothesis
+generating, thus analyses remain merely descriptive by reporting on the
+proportion of articles/tests that are statistically (in)consistent. Our
+investigation was set out to explore the following research questions:
+(1) How prevalent are statistical inconsistencies (and decision
+inconsistencies) in our sample? (see Section~\ref{sec-prevalent}) (2) Do
+inconsistency rates vary across journals and/or their publication year?
+(see Section~\ref{sec-stable}) And (3) is there evidence for bias,
+i.e.~are processes that result in inconsistencies more likely to produce
+lower p-values? (see Section~\ref{sec-bias})
+
+\section{Results}\label{results}
+
+\subsection{Inconsistencies are highly prevalent}\label{sec-prevalent}
+
+The results are summarized in Table~\ref{tbl-summary}. Out of 13065
+articles, 6484 articles contained statistical tests that Statcheck could
+assess (50\%), amounting to 85442 assessable p-values. Interestingly,
+the five journals that did not explicitly encourage APA formatting had
+virtually identical rates of assessable articles compared to the
+original sample. Overall, 11279 p-values were flagged as inconsistent
+(13.2\%) out of which 1442 were considered decision inconsistencies
+(1.7\%) (see Table~\ref{tbl-summary}).
 
@@
 
-The proportion of inconsistencies ranged from 10 to 22\% across journals
+\subsection{Inconsistencies are stable across journals and publication
+year}\label{sec-stable}
+
+The proportion of inconsistencies ranged from 8 to 23\% across journals
 (1 to 7\% for decision inconsistencies) (see
@@
 
-When examining reported against recalculated p-values for all
-inconsistencies (see Figure~\ref{fig-scatter}), we can identify certain
-spatial patterns (only inconsistencies that are based on `=' comparisons
-were included to make them easier to interpret, n = 3779 out of 5166).
-First, the center of density of points is located in the bottom left
-corner with more p-values reported closer to the most common alpha level
-of 0.05. This is not surprising, since the quantitative sciences exhibit
-a documented publication bias, with hypotheses being more often
-confirmed (and thus p \textless{} .05) than not
-(\citeproc{ref-franco2014publication}{Franco et al., 2014};
-\citeproc{ref-sterling1959publication}{Sterling, 1959}). Second, there
-is a linear density of points along the diagonal axis which corresponds
-to numerically small inconsistencies, some of which might be related to
-simple rounding errors. However, comparing the diagonal to the black
-line, which represents a linear model predicting recalculated by
-reported p-values, we can see a clear divergence of what is expected if
-inconsistencies were equally likely in both directions. The slope of the
-regression line is flatter than the diagonal axis which means that, on
-average, inconsistencies have a tendency to exhibit smaller reported
-p-values than their recalculated counterparts. Similarly,
-inconsistencies that report the p-value as being larger or smaller than
-a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05 )
-are not equally prevalent. There were 4.4\% of inconsistencies with p
-being reported as larger than a reference but 7.4\% inconsistencies with
-p being reported as smaller than a reference (e.g.~p \textless{} 0.05).
+\subsection{Inconsistencies appear biased but bias decreases over
+time}\label{sec-bias}
+
+If inconsistencies were bias-free, we would expect different types of
+inconsistencies to be equally frequent. However, this is not the case.
+Inconsistencies that report the p-value as being larger or smaller than
+a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05,
+respectively) are not equally prevalent in the sample: There were 4.4\%
+of inconsistencies with p being reported as larger than a reference but
+7.1\% inconsistencies with p being reported as smaller than a reference.
 So even if we assumed these inconsistencies were merely typos of the
-comparison sign, inconsistencies that erroneously report the p-value to
-be smaller than a reference value are more frequent than inconsistencies
-that erroneously report the p-value to be larger than a reference value.
+comparison sign (e.g.~\textless{} instead of \textgreater),
+inconsistencies that erroneously report the p-value to be smaller than a
+reference value are more frequent than inconsistencies that erroneously
+report the p-value to be larger than a reference value.
+
 These biases are also reflected in decision inconsistencies. Of all
-decision inconsistencies (n = 655), 71\% represent cases in which a
+decision inconsistencies (n = 1442), 72\% represent cases in which a
 reported significant result (p \textless{} 0.05) is recalculated as
-non-significant (p \textgreater{} 0.05).
+non-significant (p \textgreater{} 0.05), i.e.~non-significant results
+are more than twice as likely to be erroneously reported as significant
+than the other way around. The latter pattern, however, seems to have
+decreased over time. Reproducing Nuijten et al.
+(\citeproc{ref-nuijten2016prevalence}{2016}), Figure~\ref{fig-scatter}
+plots the development of the bias observed for decision inconsistencies
+over time. The prevalence of decision inconsistencies in significant
+p-values seems to have slightly decreased over the years, while the
+prevalence of decision inconsistencies in non-significant p-values seems
+to have slightly increased over the years.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{../plots/figure3.png}
+\includegraphics[width=0.8\textwidth,height=\textheight]{../plots/figure3.png}
 
@@
 
-\section{5. Discussion and
+\section{Discussion and
 Recommendations}\label{discussion-and-recommendations}
 
-\subsection{5.1. Statistical reporting inconsistencies are
+\subsection{Statistical reporting inconsistencies are
 prevalent}\label{statistical-reporting-inconsistencies-are-prevalent}
@@
 assessable articles contained at least one inconsistency and 13\%
-contained at least one decision inconsistency.
+contained at least one decision inconsistency. The present examination
+did not indicate any noticeable trend for inconsistency rates across
+journals or publication year.
 
 The present study can be considered a conceptual replication of previous
-studies investigating statistical-reporting inconsistency across
+studies investigating statistical reporting inconsistency across
 different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
@@
 \citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
-recent such assessment using Statcheck
+recent assessments using the automatic tool Statcheck
 (\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
@@
 \citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
-discovered rates fall in line with these studies who report on
-inconsistency rates between 4\% and 14\%, with between 10\% and 63\% of
-articles containing at least one inconsistency.
+discovered inconsistency rates fall in line with these studies that
+report on rates between 4\% and 14\%, with between 10\% and 63\% of
+articles containing at least one inconsistency. Moreover, the observed
+rates of inconsistencies and decision inconsistencies are virtually
+identical to rates reported on by Nuijten et al.
+(\citeproc{ref-nuijten2016prevalence}{2016}) on the psychological
+science literature.
 
@@
 
-The present examination did not indicate any noticeable trend over the
-last 23 years of publications. Observed inconsistencies were
-characterized by reported p-values being on average lower than their
-recalculated counterparts and the prevalence of decision inconsistencies
-was higher for p-values reported as significant than for those reported
-as non-significant. These biases have been reported on for other
-disciplines as well (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
-2011}; \citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}).
-These skewed patterns could indicate a systematic bias in favor of lower
-p-values in general and a bias towards significant results in
+Our results suggest biases as well. The prevalence of decision
+inconsistencies was higher for p-values reported as significant than for
+those reported as non-significant. These biases have been reported on
+for other disciplines as well (\citeproc{ref-bakker2011mis}{Bakker \&
+Wicherts, 2011}; \citeproc{ref-nuijten2016prevalence}{Nuijten et al.,
+2016}). These skewed patterns could indicate a systematic bias in favor
+of lower p-values in general and a bias towards significant results in
 particular. Our data do not speak to the causes of these biases, but
-possible reasons include the following: Researchers might intentionally
-round down p-values because they think lower p-values are more
-convincing to reviewers and readers. This practice has been admitted to
-by 1 in 5 surveyed psychological researchers
-(\citeproc{ref-john2012measuring}{John et al., 2012}). Given that a
-non-trivial amount of quantitative linguists have admitted to commit
-questionable research practices (and even fraud)
+possible reasons include the following:
+
+First, researchers might intentionally round down p-values because they
+think lower p-values are more convincing to reviewers and readers. This
+practice has been admitted to by 1 in 5 surveyed psychological
+researchers (\citeproc{ref-john2012measuring}{John et al., 2012}). Given
+that a non-trivial amount of quantitative linguists have admitted to
+commit questionable research practices (and even fraud)
 (\citeproc{ref-isbell2022misconduct}{Isbell et al., 2022}), we cannot
@@
 likely to blame the methodology while results that confirmed their
-expectations were rarely critically scrutinized. Alternatively, the
-observed bias might merely be a reflection of publication bias
-(\citeproc{ref-franco2014publication}{Franco et al., 2014};
+expectations were rarely critically scrutinized.
+
+Third, the observed bias might merely be a reflection of publication
+bias (\citeproc{ref-franco2014publication}{Franco et al., 2014};
 \citeproc{ref-sterling1959publication}{Sterling, 1959}) with
 (erroneously) reported significant p-values being more likely to be
-published than non-significant ones.
-
-\subsection{5.2. Limitations of our
-study}\label{limitations-of-our-study}
+published than (erroneously) reported non-significant ones. Publication
+bias is a well established pattern in experimental linguistic research
+with many recent meta analyses discussing possible evidence for it
+(\citeproc{ref-isbilen2022statistical}{Isbilen \& Christiansen, 2022};
+\citeproc{ref-lehtonen2018bilingualism}{Lehtonen et al., 2018};
+\citeproc{ref-lu2024meta}{Lu et al., 2024}) For example, De Bruin et al.
+(\citeproc{ref-de2015cognitive}{2015}) showed that studies with results
+supporting the bilingual-advantage theory were most likely to be
+published, while studies with results challenging the theory were
+significantly less likely to be published.
+
+Regardless of what possibly causes biases in the processes that generate
+inconsistencies, our data also suggest a positive development. The over
+proportional occurrence of decision inconsistencies for p-values
+reported as significant has decreased over time.
+
+\subsection{Limitations of our study}\label{limitations-of-our-study}
 
@@
 
-\subsection{5.3. Recommendations for the
+\subsection{Recommendations for the
 field}\label{recommendations-for-the-field}
@@
 \citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018};
-\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022}).
+\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022})
+and a number of linguistic journals have already implemented such
+policies, including journals within our sample. Having said that, open
+data policies which for example the Journal of Memory and Language
+introduced in 2018 did not seem to have affected the proportion of
+statistical inconsistencies after their introduction. The inconsistency
+rates for JML (and other journals) were rather stable across time, so
+open data alone might not resolve the issue without further changes to
+the research eco-system.
 

### 2025-05-20 manuscript/manuscript_anonymized.qmd
ready for resubmission

@@
 shorttitle: "Statistical Inconsistencies"
+number-sections: true
 author:
-    - name: anonymous
-      affiliations:
-      - name: anonymous
+  - name: "anonymous"
+    affiliations: "anonymous"
+  - name: "anonymous"
   # - name: Dara Leonard Jenssen Etemady
@@
   #       country: Norway
-author-note:
-  status-changes: 
-    affiliation-change: null
-    deceased: null
-  disclosures:
-    study-registration: null
-    data-sharing: null
-    related-report: null
-    conflict-of-interest: The authors have no conflict of interest to declare.
-    financial-support: null
-    gratitude: null
+# author-note:
+#   status-changes: 
+#     affiliation-change: null
+#     deceased: null
+#   disclosures:
+#     study-registration: null
+#     data-sharing: null
+#     related-report: null
+#     conflict-of-interest: The authors have no conflict of interest to declare.
+#     financial-support: null
+#     gratitude: null
     # authorship-agreements: Conceptionalization, Methodology, Validation, Formal Analysis, Review & Editing of Manuscript, Data Curation - T.B.R.
 & D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft, Visualization, Supervision - T.B.R.
-abstract: "The present paper investigates the prevalence of statistical reporting consistencies across articles in eight experimental linguistics journals published between 2000 and 2023.
+abstract: "The present paper investigates the prevalence of statistical reporting inconsistencies across articles in thirteen experimental linguistics journals published between 2000 and 2023.
 Using the R package “Statcheck”, we retrieved 85,442 statistical test from 13,065 articles and assessed whether p-values were consistent with their test statistic and degrees of freedom.
@@
 Depending on the results of their tests, researchers make claims for or against theoretical assumptions.\
-Since these tests play such an integral part in the argumentation process of experimentalists, both the data these tests are based on and the computational procedure of the tests itself should be transparent in order for other researchers to critically evaluate them.
-But more than that, things can go wrong.
-Humans are fallible; they make mistakes.
+Since these tests play such an integral part in the argumentation process of experimentalists, both the data these tests are based on and the computational procedure of the tests itself should be error-free.
+But humans are fallible; they make mistakes.
+We cannot avoid making errors, but we can make them at least detectable.
 Transparent sharing allows others to detect and correct human error.
+
 In recent years, quantitative linguistics have seen repeated calls to become more transparent and reproducible by sharing data and statistical protocols, often under the banner of "open science" [@roettger2019researcher; @arvan2022reproducibility; @laurinavichyute2022share].
@@
 
-To assess the prevalence of statistical-reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in thirteen experimental linguistic journals published between 2000 and 2023.
-We explore whether the rate of inconsistency differs across journals, whether their rate has changed over the course of the last 23 years and whether there is evidence for bias in these statistical-reporting inconsistencies.
+To assess the prevalence of statistical reporting inconsistencies in experimental linguistics, the present paper conceptually replicates @nuijten2016prevalence and assesses p-values reported in thirteen experimental linguistic journals published between 2000 and 2023.
+We explore whether the inconsistency rates differ across journals, whether they have changed over the course of the last 23 years and whether there is evidence for bias in these inconsistencies.
 We discuss the results and offer concrete recommendations for authors, reviewers, and editors to tackle this problem.
@@
 
-# 3. Method
+# Method
 
-## 3.1. Data availability statement
+## Data availability statement
 
@@
 
-## 3.2. Sample
+## Sample
 
@@
 An anonymous reviewer raised the justified concern that the resulting sample might not represent the vast majority of work in experimental linguistics. 
-To address this concern, we additionally included those five journals that have published the highest absolute number of experimental articles, regardless of whether they explicitly urge APA formatting or not, resulting in the inclusion of the following psycholinguistic outlets: Journal of Memory and Language (JML), Language Cognition and Neuroscience (LCN, former Language and Cognitive Processes), Journal of Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing Research (SLH), and Brain and Language (BAL).
+To address this concern, we additionally included those five journals that have published the highest absolute number of experimental articles [according to @kobrock2023assessing], regardless of whether they explicitly urge APA formatting or not, resulting in the inclusion of the following psycholinguistic outlets: Journal of Memory and Language (JML), Language, Cognition and Neuroscience (LCN, former Language and Cognitive Processes), Journal of Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing Research (SLH), and Brain and Language (BAL).
 
@@
 
-## 3.3. Statcheck
+## Statcheck
 
-We used the R package Statcheck [Version 1.5.0, @nuijten2023statcheck] to automatically detect statistical-reporting inconsistencies.
+We used the R package Statcheck [Version 1.5.0, @nuijten2023statcheck] to automatically detect statistical reporting inconsistencies.
 Statcheck works as follows: After converting articles in pdf or html format to plain text, Statcheck searches for specific strings that correspond to a NHST result using regular expressions.
-That way, Statcheck can detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, the degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
+That way, Statcheck can detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q-tests as long as the test result fulfills three conditions: (a) the test result is reported completely including the test statistic, the degrees of freedom (if applicable), and the p-value; (b) the test result is in the body of the text, i.e.
 Statcheck usually misses information in tables; and (c) the test result is reported in American Psychological Association style [@APA2020].
@@
 
-Articles from LAB spanned 2011-2023; LCN spanned 2015-2023.
+Articles from Linguistic Approaches to Bilingualism spanned 2011-2023; articles from Language, Cognition and Neuroscience spanned 2015-2023.
 Statcheck could not parse 291 articles, likely related to issues with rendering the Chi-Squared symbol being erroneously converted from .pdf to .txt, resulting in some gaps in the coverage.
@@
 
-## 3.4. Analysis and research questions
+## Analysis and research questions
 The aims of this study were explicitly exploratory and hypothesis generating, thus analyses remain merely descriptive by reporting on the proportion of articles/tests that are statistically (in)consistent.
 Our investigation was set out to explore the following research questions:
-(1) How prevalent are statistical inconsistencies (and decision inconsistencies) in our sample?
-(2) Do inconsistency rates vary across journals and/or their publication year? 
-And (3) is there evidence for bias, i.e. are processes that result in inconsistencies more likely to produce lower p-values?
+(1) How prevalent are statistical inconsistencies (and decision inconsistencies) in our sample? (see @sec-prevalent)
+(2) Do inconsistency rates vary across journals and/or their publication year? (see @sec-stable)
+And (3) is there evidence for bias, i.e. are processes that result in inconsistencies more likely to produce lower p-values? (see @sec-bias)
 
-# 4. Results
+# Results
 
-## 4.1 Inconsistencies are highly prevalent
+## Inconsistencies are highly prevalent {#sec-prevalent}
 
@@
 
-## 4.2 Inconsistencies are stable across journals and publication year
+## Inconsistencies are stable across journals and publication year {#sec-stable}
 
@@
 
-## 4.3 Inconsistencies are biased
+## Inconsistencies appear biased but bias decreases over time {#sec-bias}
 
+If inconsistencies were bias-free, we would expect different types of inconsistencies to be equally frequent. 
+However, this is not the case.
 Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample:
@@
 non-significant results are more than twice as likely to be erroneously reported as significant than the other way around.
-The latter pattern, however, seems to be a bias of the past. 
+The latter pattern, however, seems to have decreased over time. 
 Reproducing @nuijten2016prevalence, @fig-scatter plots the development of the bias observed for decision inconsistencies over time.
@@
 #| label: fig-scatter
-#| out-width: 100%
+#| out-width: 80%
 #| fig-cap: "Percentage of decision inconsistencies falsely reporting significance (black) or non-significance (orange), plotted across publication years."
@@
 
-# 5. Discussion and Recommendations
+# Discussion and Recommendations
 
-## 5.1. Statistical reporting inconsistencies are prevalent
+## Statistical reporting inconsistencies are prevalent
 
@@
 On average, `{r} 100*round(gross_per_article[[1]],2)`% of assessable articles contained at least one inconsistency and `{r} 100*round(gross_per_article[[2]],2)`% contained at least one decision inconsistency.
+The present examination did not indicate any noticeable trend for inconsistency rates across journals or publication year.
 
-The present study can be considered a conceptual replication of previous studies investigating statistical-reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent assessments using Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
+The present study can be considered a conceptual replication of previous studies investigating statistical reporting inconsistency across different disciplines [@bakker2011mis; @bakker2014outlier; @caperos2013consistency; @veldkamp2014statistical; @wicherts2011willingness] and most recent assessments using the automatic tool Statcheck [@nuijten2016prevalence; @buckley2023estimating; @colombo2018statistical; @gross2021fidelity].
 The discovered inconsistency rates fall in line with these studies that report on rates between 4% and 14%, with between 10% and 63% of articles containing at least one inconsistency. 
+Moreover, the observed rates of inconsistencies and decision inconsistencies are virtually identical to rates reported on by @nuijten2016prevalence on the psychological science literature.
 
@@
 
-The present examination did not indicate any noticeable trend over a period of 23 years of publications.
-Observed inconsistencies were characterized by reported p-values being on average lower than their recalculated counterparts and the prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant.
+Our results suggest biases as well. 
+The prevalence of decision inconsistencies was higher for p-values reported as significant than for those reported as non-significant.
 These biases have been reported on for other disciplines as well [@nuijten2016prevalence; @bakker2011mis].
@@
 
-Third, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than non-significant ones.
+Third, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than  (erroneously) reported non-significant ones.
 Publication bias is a well established pattern in experimental linguistic research with many recent meta analyses discussing possible evidence for it [@isbilen2022statistical; @lehtonen2018bilingualism; @lu2024meta]
@@
 
-Regardless of possible causes for bias in the processes that generate inconsistencies, the data also suggest that some of these biases has decreased over time, pointing to a positive development.
+Regardless of what possibly causes biases in the processes that generate inconsistencies, our data also suggest a positive development. 
+The over proportional occurrence of decision inconsistencies for p-values reported as significant has decreased over time.
 
-## 5.2. Limitations of our study
+## Limitations of our study
 
@@
 
-## 5.3. Recommendations for the field
+## Recommendations for the field
 
@@
 Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share] and a number of linguistic journals have already implemented such policies, including journals within our sample.
-Having said that, open data policies which for example the Journal of Memory and Language introduced in 2018 did not seem to have affected the proportion of statistical inconsistencies.
+Having said that, open data policies which for example the Journal of Memory and Language introduced in 2018 did not seem to have affected the proportion of statistical inconsistencies after their introduction.
 The inconsistency rates for JML (and other journals) were rather stable across time, so open data alone might not resolve the issue without further changes to the research eco-system.

### 2025-05-20 manuscript/manuscript_anonymized.tex
ready for resubmission

@@
 
-\author{anonymous}
+\authorsnames{anonymous,anonymous}
 
@@
 
-\leftheader{anonymous}
+\leftheader{anonymous and anonymous}
 
 \abstract{The present paper investigates the prevalence of statistical
-reporting consistencies across articles in eight experimental
+reporting inconsistencies across articles in thirteen experimental
 linguistics journals published between 2000 and 2023. Using the R
@@
 
-\keywords{statistics, statcheck, reproducibility, null hypothesis
+\keywords{statistics, Statcheck, reproducibility, null hypothesis
 significance testing}
@@
 \authornote{
+
 \par{       }
-\par{   The authors have no conflict of interest to declare.    }
 \par{Correspondence concerning this article should be addressed to }
@@
 process of experimentalists, both the data these tests are based on and
-the computational procedure of the tests itself should be transparent in
-order for other researchers to critically evaluate them. But more than
-that, things can go wrong. Humans are fallible; they make mistakes.
-Transparent sharing allows others to detect and correct human error. In
-recent years, quantitative linguistics have seen repeated calls to
+the computational procedure of the tests itself should be error-free.
+But humans are fallible; they make mistakes. We cannot avoid making
+errors, but we can make them at least detectable. Transparent sharing
+allows others to detect and correct human error.
+
+In recent years, quantitative linguistics have seen repeated calls to
 become more transparent and reproducible by sharing data and statistical
@@
 
-To assess the prevalence of statistical-reporting inconsistencies in
+To assess the prevalence of statistical reporting inconsistencies in
 experimental linguistics, the present paper conceptually replicates
@@
 p-values reported in thirteen experimental linguistic journals published
-between 2000 and 2023. We explore whether the rate of inconsistency
-differs across journals, whether their rate has changed over the course
-of the last 23 years and whether there is evidence for bias in these
-statistical-reporting inconsistencies. We discuss the results and offer
-concrete recommendations for authors, reviewers, and editors to tackle
-this problem.
-
-\section{3. Method}\label{method}
-
-\subsection{3.1. Data availability
+between 2000 and 2023. We explore whether the inconsistency rates differ
+across journals, whether they have changed over the course of the last
+23 years and whether there is evidence for bias in these
+inconsistencies. We discuss the results and offer concrete
+recommendations for authors, reviewers, and editors to tackle this
+problem.
+
+\section{Method}\label{method}
+
+\subsection{Data availability
 statement}\label{data-availability-statement}
@@
 
-\subsection{3.2. Sample}\label{sample}
+\subsection{Sample}\label{sample}
 
@@
 address this concern, we additionally included those five journals that
-have published the highest absolute number of experimental articles,
-regardless of whether they explicitly urge APA formatting or not,
-resulting in the inclusion of the following psycholinguistic outlets:
-Journal of Memory and Language (JML), Language Cognition and
+have published the highest absolute number of experimental articles
+(according to \citeproc{ref-kobrock2023assessing}{Kobrock \& Roettger,
+2023}), regardless of whether they explicitly urge APA formatting or
+not, resulting in the inclusion of the following psycholinguistic
+outlets: Journal of Memory and Language (JML), Language, Cognition and
 Neuroscience (LCN, former Language and Cognitive Processes), Journal of
@@
 
-\subsection{3.3. Statcheck}\label{statcheck}
+\subsection{Statcheck}\label{statcheck}
 
@@
 \citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2024}) to
-automatically detect statistical-reporting inconsistencies. Statcheck
+automatically detect statistical reporting inconsistencies. Statcheck
 works as follows: After converting articles in pdf or html format to
@@
 
-Articles from LAB spanned 2011-2023; LCN spanned 2015-2023. Statcheck
-could not parse 291 articles, likely related to issues with rendering
-the Chi-Squared symbol being erroneously converted from .pdf to .txt,
-resulting in some gaps in the coverage. This procedure resulted in 13065
-research articles which were submitted to analysis.
-
-\subsection{3.4. Analysis and research
+Articles from Linguistic Approaches to Bilingualism spanned 2011-2023;
+articles from Language, Cognition and Neuroscience spanned 2015-2023.
+Statcheck could not parse 291 articles, likely related to issues with
+rendering the Chi-Squared symbol being erroneously converted from .pdf
+to .txt, resulting in some gaps in the coverage. This procedure resulted
+in 13065 research articles which were submitted to analysis.
+
+\subsection{Analysis and research
 questions}\label{analysis-and-research-questions}
@@
 (1) How prevalent are statistical inconsistencies (and decision
-inconsistencies) in our sample? (2) Do inconsistency rates vary across
-journals and/or their publication year? And (3) is there evidence for
-bias, i.e.~are processes that result in inconsistencies more likely to
-produce lower p-values?
-
-\section{4. Results}\label{results}
-
-\subsection{4.1 Inconsistencies are highly
-prevalent}\label{inconsistencies-are-highly-prevalent}
+inconsistencies) in our sample? (see Section~\ref{sec-prevalent}) (2) Do
+inconsistency rates vary across journals and/or their publication year?
+(see Section~\ref{sec-stable}) And (3) is there evidence for bias,
+i.e.~are processes that result in inconsistencies more likely to produce
+lower p-values? (see Section~\ref{sec-bias})
+
+\section{Results}\label{results}
+
+\subsection{Inconsistencies are highly prevalent}\label{sec-prevalent}
 
@@
 
-\subsection{4.2 Inconsistencies are stable across journals and
-publication
-year}\label{inconsistencies-are-stable-across-journals-and-publication-year}
+\subsection{Inconsistencies are stable across journals and publication
+year}\label{sec-stable}
 
@@
 
-\subsection{4.3 Inconsistencies are
-biased}\label{inconsistencies-are-biased}
-
+\subsection{Inconsistencies appear biased but bias decreases over
+time}\label{sec-bias}
+
+If inconsistencies were bias-free, we would expect different types of
+inconsistencies to be equally frequent. However, this is not the case.
 Inconsistencies that report the p-value as being larger or smaller than
@@
 are more than twice as likely to be erroneously reported as significant
-than the other way around. The latter pattern, however, seems to be a
-bias of the past. Reproducing Nuijten et al.
+than the other way around. The latter pattern, however, seems to have
+decreased over time. Reproducing Nuijten et al.
 (\citeproc{ref-nuijten2016prevalence}{2016}), Figure~\ref{fig-scatter}
@@
 \citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
-recent assessments using Statcheck
+recent assessments using the automatic tool Statcheck
 (\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
@@
 report on rates between 4\% and 14\%, with between 10\% and 63\% of
-articles containing at least one inconsistency.
+articles containing at least one inconsistency. Moreover, the observed
+rates of inconsistencies and decision inconsistencies are virtually
+identical to rates reported on by Nuijten et al.
+(\citeproc{ref-nuijten2016prevalence}{2016}) on the psychological
+science literature.
 
@@
 
-The present examination did not indicate any noticeable trend over a
-period of 23 years of publications. Observed inconsistencies were
-characterized by reported p-values being on average lower than their
-recalculated counterparts and the prevalence of decision inconsistencies
-was higher for p-values reported as significant than for those reported
-as non-significant. These biases have been reported on for other
-disciplines as well (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
-2011}; \citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}).
-These skewed patterns could indicate a systematic bias in favor of lower
-p-values in general and a bias towards significant results in
+Our results suggest biases as well. The prevalence of decision
+inconsistencies was higher for p-values reported as significant than for
+those reported as non-significant. These biases have been reported on
+for other disciplines as well (\citeproc{ref-bakker2011mis}{Bakker \&
+Wicherts, 2011}; \citeproc{ref-nuijten2016prevalence}{Nuijten et al.,
+2016}). These skewed patterns could indicate a systematic bias in favor
+of lower p-values in general and a bias towards significant results in
 particular. Our data do not speak to the causes of these biases, but
@@
 (erroneously) reported significant p-values being more likely to be
-published than non-significant ones. Publication bias is a well
-established pattern in experimental linguistic research with many recent
-meta analyses discussing possible evidence for it
+published than (erroneously) reported non-significant ones. Publication
+bias is a well established pattern in experimental linguistic research
+with many recent meta analyses discussing possible evidence for it
 (\citeproc{ref-isbilen2022statistical}{Isbilen \& Christiansen, 2022};
@@
 
-Regardless of possible causes for bias in the processes that generate
-inconsistencies, the data also suggest that some of these biases has
-decreased over time, pointing to a positive development.
-
-\subsection{5.2. Limitations of our
-study}\label{limitations-of-our-study}
+Regardless of what possibly causes biases in the processes that generate
+inconsistencies, our data also suggest a positive development. The over
+proportional occurrence of decision inconsistencies for p-values
+reported as significant has decreased over time.
+
+\subsection{Limitations of our study}\label{limitations-of-our-study}
 
@@
 introduced in 2018 did not seem to have affected the proportion of
-statistical inconsistencies. The inconsistency rates for JML (and other
-journals) were rather stable across time, so open data alone might not
-resolve the issue without further changes to the research eco-system.
+statistical inconsistencies after their introduction. The inconsistency
+rates for JML (and other journals) were rather stable across time, so
+open data alone might not resolve the issue without further changes to
+the research eco-system.
 

### 2025-05-20 scripts/01_Analysis.R

@@ -298,7 +298,8 @@ year_journal_plot <-
    scale_y_continuous(breaks = c(0,.5,1),
                       labels = c("0%", "25%", "50%")
                       ) +
-  scale_x_continuous(breaks =  c(2000, 2023)) +
+  scale_x_continuous(breaks =  c(2000, 2023),
+                     labels = c("'00", "'23")) +
   theme_minimal() + 
     theme(legend.position = "none",
           panel.spacing.x = unit(1, "lines"))
@@ -345,42 +346,44 @@ ggplot(xdata_sub_agg |> filter(direction != "NA"),
            y = prop_gross,
            shape = direction)) + 
   geom_path(aes(group = direction, color = direction),
-            linewidth = 3,
+            linewidth = 2,
             lineend = "round") +
   # legend
   annotate("point", 
            x = 2015,
-           y = 3.5,
+           y = 4.5,
            size = 5,
+           color = "white",
            fill = "black",
            pch = 21) +
   annotate("point", 
            x = 2015,
-           y = 3.25,
+           y = 4.0,
            size = 5,
            color = "white",
            fill = "#ee6002",
            pch = 21) +
   annotate("text", 
            x = 2016,
-           y = 3.5,
+           y = 4.5,
            hjust = 0,
            color = "black",
            label = "falsely reported as significant") +
   annotate("text", 
            x = 2016,
-           y = 3.25,
+           y = 4.0,
            hjust = 0,
            color = "black",
            label = "falsely reported as not significant") +
   scale_color_manual(values = c("#ee6002", "black")) +
-  scale_y_continuous(breaks = c(0,0.5,1,2,3),
-                     labels = c("0%", "0.5%", "1%", "2%", "3%")
+  scale_y_continuous(limits = c(0,5),
+                     breaks = c(0,0.5,1,2,3,4,5),
+                     labels = c("0%", "0.5%", "1%", "2%", "3%", "4%", "5%")
   ) +
   scale_x_continuous(limits = c(2000,2024),
                      breaks =  c(2000, 2005, 2010, 2015, 2020, 2023)) +
   labs(x = "\nyear of publication",
-       y = "percentage of decision errors\n") +
+       y = "percentage of decision inconsistencies\n") +
   theme_minimal() + 
   theme(legend.position = "none",
         panel.spacing.x = unit(1, "lines"))
@@ -388,8 +391,8 @@ ggplot(xdata_sub_agg |> filter(direction != "NA"),
[…]

### 2025-07-17 manuscript/manuscript.qmd
fixed minor revision and added script for subset sample

@@
 Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample:
-There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference.
+There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistent tests with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% of inconsistent tests with p being reported as smaller than a reference.
 So even if we assumed these inconsistencies were merely typos of the comparison sign (e.g.
@@
 #| label: fig-scatter
-#| out-width: 80%
+#| out-width: 100%
 #| fig-cap: "Percentage of decision inconsistencies falsely reporting significance (black) or non-significance (orange), plotted across publication years."
@@
 They are human errors.
-If such a substantial amount of errors is found in plain site, the question naturally arises as to how many errors during the data analysis itself remain undetected.
+If such a substantial amount of errors is found in plain sight, the question naturally arises as to how many errors during the data analysis itself remain undetected.
 We should ask ourselves, if the tip of the iceberg is already so large, what is the volume of the submerged iceberg?

### 2025-07-17 scripts/01_Analysis.R

@@ -1,6 +1,6 @@
 ## title: Statistical reporting inconsistencies in Ling
-## author: anonymous
-## contact: anonymous   
+## author: Timo Roettger & Dara Etemady
+## contact: timo.b.roettger@gmail.com   
 ## date created: 2024-07-11
 ## date last edited: 2024-07-25
 
@@ -197,6 +197,26 @@ stacked_bar <-
                 label = paste0(round(prop_has_gross * 100, 0), "%")),
             vjust = 0,
             color = "white") +
+  # legend
+  annotate("rect",
+           xmin = 0.8,
+           xmax = 5,
+           ymin = 82,
+           ymax = 99,
+           fill = "white") + 
+  annotate("rect",
+           xmin = c(0.9, 0.9),
+           xmax = c(1.3,1.3),
+           ymin = c(85,92),
+           ymax = c(89,96),
+           fill = c("#ee6002","#6002ee")) + 
+  annotate("text",
+           label = c("inconsistency", "decision inconsistency"),
+           x = c(1.5, 1.5),
+           y = c(94, 87),
+           color = c("black","black"),
+           size = 4,
+           hjust = 0) + 
   scale_fill_manual(name = "",
                     values = c("prop_has_error" = "#6002ee", "prop_has_gross" = "#ee6002"),
                     labels = c("inconsistency", "decision inconsistency")) +
@@ -265,16 +285,36 @@ year_plot <-
               type = "proportion")  +
   scale_fill_manual(values = c("lightgrey","#6002ee","#ee6002"),
                     name = "") +
-  annotate("label",
+  # legend
+  annotate("rect",
+           xmin = 2000.25,
+           xmax = 2017,
+           ymin = 0.82,
+           ymax = 0.99,
+           fill = "white") + 
+  annotate("rect",
+           xmin = c(2000.5, 2000.5),
+           xmax = c(2001.5,2001.5),
+           ymin = c(0.85,0.92),
+           ymax = c(0.89,0.96),
+           fill = c("#ee6002","#6002ee")) + 
+  annotate("text",
            label = c("inconsistency", "decision inconsistency"),
-           x = c(2001, 2001),
-           y = c(0.95, 0.86),
-           fill = c("#6002ee","#ee6002"),
-           color = c("white", "white"),
-           fontface = "bold",
-           label.padding = unit(0.5, "lines"),
-           size = 3,
+           x = c(2002, 2002),
+           y = c(0.94, 0.87),
+           color = c("black","black"),
+           size = 4,
[…]

### 2025-07-17 scripts/02_sampling_for_manual_inspection.R

@@ -1,0 +1,112 @@
+## title: Statistical reporting inconsistencies in Ling - Sampling for manual check
+## author: Timo Roettger
+## contact: timo.b.roettger@gmail.com   
+## date created: 2024-07-11
+## date last edited: 2024-07-25
+
+
+## packages ####################################################################                        
+
+# load in relevant libraries, install if not installed
+# if (!require("pacman")) install.packages("pacman")
+# pacman::p_load(rstudioapi, 
+#                tidyverse, 
+#                ggpubr
+# )
+
+## data ########################################################################                        
+
+## load in data from location of script 
+## (uncomment if you run script outside of sourcing it through manuscript.qmd)
+#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
+#setwd(current_working_dir)
+
+## load in data
+xdata <- read_csv("../data/statcheck_revised_sample.csv")
+
+## split strings into year and journal, drop NAs
+xdata <- xdata %>% 
+  mutate(year = strtoi(substring(source, 1, 4)),
+         journal = substring(source, 9, 11)) %>%
+  drop_na(reported_p) %>%
+  drop_na(computed_p) %>%
+  filter(year != 2024)
+
+## remove LCN before before 2015 
+xdata <- xdata |> 
+  filter(! (journal == "LCN" & year < 2015))
+
+# wrangle for sampling
+xdata <- xdata |> 
+  filter(decision_error == TRUE)
+
+# check distribution
+table(xdata$journal, xdata$year)
+
+# many journals have gaps, so we only sample across years
+
+# randomly sample x% of decision inconsistencies
+# max one per article
+# evenly across years
+
+sample_prop = 0.15
+sample_size = nrow(xdata) * sample_prop
+# 216
+sample_size_per_year = sample_size / length(unique(xdata$year))
+# 9
+  
+# set random seed
+set.seed(999)
+
+# first create subset with one test per article
+xdata_sample <- xdata |> 
+  group_by(source) %>% 
+  sample_n(1)
+
+# check if possible
+table(xdata_sample$year)
+
+# yes, proceed
+
+# second create subset with the desired tests per year
+xdata_sample <- xdata_sample |> 
+  group_by(year) %>% 
+  sample_n(sample_size_per_year)
+
+# check if duplicate articles
+duplicated(xdata_sample$source)
+# no
+
+# save list
+write_csv(xdata_sample, "../data/subsample/subsample.csv")
+
+
+# extract files and store them into folder
[…]

### 2025-07-19 scripts/01_Analysis.R
changed subsample specifications

@@ -32,8 +32,8 @@ library(ggstream)
 
 ## load in data from location of script 
 ## (uncomment if you run script outside of sourcing it through manuscript.qmd)
-#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
-#setwd(current_working_dir)
+current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
+setwd(current_working_dir)
 
 ## list all journal units
 ## only works with access to all article pdfs, skip if you don't have access
@@ -166,7 +166,6 @@ gross_per_article <- xdata %>%
 ## per_article into long format for plot
 per_article_long <- per_article %>% 
   pivot_longer(!journal, names_to = "type", values_to = "proportion")
-
 
 ## Figure 1 ####################################################################                       
 

### 2025-07-19 scripts/02_sampling_for_manual_inspection.R

@@ -45,6 +45,14 @@ table(xdata$journal, xdata$year)
 
 # many journals have gaps, so we only sample across years
 
+# take into account comparison sign. Decision errors with = are more difficult to interprete
+table(xdata$p_comp)
+table(xdata$p_comp, xdata$year)
+
+# should focus only on comparisons that are not equal
+xdata_without_equal <- xdata |> 
+  filter(p_comp != "=")
+
 # randomly sample x% of decision inconsistencies
 # max one per article
 # evenly across years
@@ -53,33 +61,43 @@ sample_prop = 0.15
 sample_size = nrow(xdata) * sample_prop
 # 216
 sample_size_per_year = sample_size / length(unique(xdata$year))
-# 9
+# desired sample per year 9
   
 # set random seed
-set.seed(999)
+set.seed(9191)
 
 # first create subset with one test per article
-xdata_sample <- xdata |> 
+xdata_sample <- xdata_without_equal |> 
   group_by(source) %>% 
   sample_n(1)
 
 # check if possible
 table(xdata_sample$year)
 
-# yes, proceed
+# except for 2021, yes
 
 # second create subset with the desired tests per year
 xdata_sample <- xdata_sample |> 
   group_by(year) %>% 
-  sample_n(sample_size_per_year)
+  # sample the desire number if possible, else sample all from the year
+  sample_n(min(n(), sample_size_per_year))
 
 # check if duplicate articles
 duplicated(xdata_sample$source)
-# no
+# no, proceed
+
+# check
+table(xdata_sample$year)
 
 # save list
 write_csv(xdata_sample, "../data/subsample/subsample.csv")
 
+# save simpler list for manual work
+xdata_sample |> 
+  select(source, p_comp, computed_p, raw, one_tailed_in_txt) |> 
+  mutate(sentence = "copy paste sentence",
+         comment = "add comments") |> 
+  write_csv("../data/subsample/subsample_for_manual_annotation.csv")
 
 # extract files and store them into folder
 # folder from which to extract
@@ -107,6 +125,42 @@ for (file in all_files) {
     file.copy(from = file, to = file.path(dest_dir, filename), overwrite = TRUE)
   }
 }
+
+---------
+
+# closer look at potential impact of the decision errors  
+
+## check all articles that contain at least one decision error
+x_gross <- xdata |> 
+  filter(decision_error == TRUE)
+
[…]

### 2025-07-19 manuscript/manuscript.qmd
prepared script for manual assessment

@@
 
-```{=tex}
 \begin{equation*}
@@
   \eqnmarkbox[blue]{node7}{p < .05}
-\end{equation*}
-
-\annotate[yshift=1em]{left}{node2}{name of test} 
-\annotate[yshift=-1em]{below,left}{node3}{degrees of freedom} 
-\annotate[yshift=1em]{right}{node5}{test statistic} 
-\annotate[yshift=-1em]{below,right}{node7}{p-value}
-```
+\end{equation*} \annotate[yshift=1em]{left}{node2}{name of test} \annotate[yshift=-1em]{below,left}{node3}{degrees of freedom} \annotate[yshift=1em]{right}{node5}{test statistic} \annotate[yshift=-1em]{below,right}{node7}{p-value}
 
@@
 
+suppressWarnings(suppressMessages(suppressPackageStartupMessages({
+source("../scripts/02_sampling_for_manual_inspection.R")
+})))
+
 knitr::opts_chunk$set(dpi = 300)
@@
 
-All quantitative analyses were conducted using @rmanual. All derived data and corresponding R scripts are available [here](https://osf.io/gx3ub/?view_only=25957e06b5f449d6820b529eb0ae1937).
+All quantitative analyses were conducted using @rmanual.
+All derived data and corresponding R scripts are available [here](https://osf.io/gx3ub/?view_only=25957e06b5f449d6820b529eb0ae1937).
 
@@
 ## Analysis and research questions
+
 The aims of this study were explicitly exploratory and hypothesis generating, thus analyses remain merely descriptive by reporting on the proportion of articles/tests that are statistically (in)consistent.
-Our investigation was set out to explore the following research questions:
-(1) How prevalent are statistical inconsistencies (and decision inconsistencies) in our sample? (see @sec-prevalent)
-(2) Do inconsistency rates vary across journals and/or their publication year? (see @sec-stable)
-And (3) is there evidence for bias, i.e. are processes that result in inconsistencies more likely to produce lower p-values? (see @sec-bias)
+Our investigation was set out to explore the following research questions: (1) How prevalent are statistical inconsistencies (and decision inconsistencies) in our sample?
+(see @sec-prevalent) (2) Do inconsistency rates vary across journals and/or their publication year?
+(see @sec-stable) (3) is there evidence for bias, i.e. are processes that result in inconsistencies more likely to produce lower p-values?
+(see @sec-bias) And (4) do decision inconsistencies matter, i.e. are they merely typos or do they lead to erroneous theoretical interpretations?
+(see @sec-matter)
 
@@
 However, this is not the case.
-Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample:
-There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistent tests with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% of inconsistent tests with p being reported as smaller than a reference.
+Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g.
+p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample: There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistent tests with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% of inconsistent tests with p being reported as smaller than a reference.
 So even if we assumed these inconsistencies were merely typos of the comparison sign (e.g.
@@
 ```
+
+## Do decision inconsistencies matter? {#sec-matter}
+
+Now the question arises, do the inconsistencies theoretically matter.
+This is a difficult question to empirically address because experimental behavioral research is often characterized by weak links between theory and predictions [@scheel2021hypothesis], making it difficult to even extract hypotheses and whether hypotheses have been corroborated or not from published articles [@scheel2022most].
+To gain some insight, we approximate an answer to this questions as follows: First, we extracted a random subsample (15%, n = `{r} nrow(xdata_sample)`) of articles which contain at least one decision inconsistencies, equally distributed across publication years.
+We restricted the subsample to only those decision inconsistencies that involve either a larger (\>) or small (\<) comparison of the p-value to allow unambiguous interpretation of the results.
+We then manually extracted the relevant sentence which the test statistic is embedded in and assessed whether the authors interpreted the reported p-value correctly (a true decision inconsistency) or incorrectly (the decision inconsistency is likely a typo).
+This ratio will let us approximate the true decision inconsistency rate.
+We consider true decision inconsistencies theoretically relevant.
+E.g. if an author claims that two groups are different (or not) based on an erroneously reported significant test, we consider this claim to potentially have theoretical consequences.
+Out of the `{r} nrow(xdata_sample)` assessed cases, we coded ## (##%) as genuine typos, e.g.
+the authors interpreted the results in line with the computed p-value and not the erroneously reported p-value.
+We coded ## (##%) cases as unclear, i.e. even after assessing the manuscript in detail, we could not interpret how the p-value was interpreted. 
+This leaves ## (##%) of tests being coded as true decision inconsistencies that potentially have theoretical consequences.
 
@@
 Third, the observed bias might merely be a reflection of publication bias [@franco2014publication; @sterling1959publication] with (erroneously) reported significant p-values being more likely to be published than (erroneously) reported non-significant ones.
-Publication bias is a well established pattern in experimental linguistic research with many recent meta analyses discussing possible evidence for it [@isbilen2022statistical; @lehtonen2018bilingualism; @lu2024meta]
-For example, @de2015cognitive showed that studies with results supporting the bilingual-advantage theory were most likely to be published, while studies with results challenging the theory were significantly less likely to be published.
+Publication bias is a well established pattern in experimental linguistic research with many recent meta analyses discussing possible evidence for it [@isbilen2022statistical; @lehtonen2018bilingualism; @lu2024meta] For example, @de2015cognitive showed that studies with results supporting the bilingual-advantage theory were most likely to be published, while studies with results challenging the theory were significantly less likely to be published.
 
@@
 Journal editors could explicitly recommend consistency checks with algorithms such as Statcheck during peer review, a practice that has been taken up on by several journals from neighboring disciplines (Psychological Science[^1], Advances in Methods and Practices in Psychological Science[^2], Stress & Health @barber2017meticulous).
-Editors could also demand, recommend or at least encourage data sharing for publication in their journal. 
-Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share] and a number of linguistic journals have already implemented such policies, including journals within our sample.
-Having said that, open data policies which for example the Journal of Memory and Language introduced in 2018 did not seem to have affected the proportion of statistical inconsistencies after their introduction.
-The inconsistency rates for JML (and other journals) were rather stable across time, so open data alone might not resolve the issue without further changes to the research eco-system.
+Editors could also demand, recommend or at least encourage data sharing for publication in their journal
+.
+Data sharing policies have been shown to substantially increase the reproducibility of analyses [e.g., @hardwicke2018data; @laurinavichyute2022share] and a number of linguistic journals have already implemented such policies, including journals within our sample
+.
+Having said that, open data policies which for example the Journal of Memory and Language introduced in 2018 did not seem to have affected the proportion of statistical inconsistencies after their introduction
+. The inconsistency rates for JML (and other journals) were rather stable across time, so open data alone might not resolve the issue without further changes to the research eco-system
+.
 

### 2025-07-19 scripts/02_sampling_for_manual_inspection.R

@@ -18,8 +18,8 @@
 
 ## load in data from location of script 
 ## (uncomment if you run script outside of sourcing it through manuscript.qmd)
-#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
-#setwd(current_working_dir)
+# current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
+# setwd(current_working_dir)
 
 ## load in data
 xdata <- read_csv("../data/statcheck_revised_sample.csv")
@@ -37,20 +37,20 @@ xdata <- xdata |>
   filter(! (journal == "LCN" & year < 2015))
 
 # wrangle for sampling
-xdata <- xdata |> 
+xdata_de <- xdata |> 
   filter(decision_error == TRUE)
 
 # check distribution
-table(xdata$journal, xdata$year)
+table(xdata_de$journal, xdata_de$year)
 
 # many journals have gaps, so we only sample across years
 
 # take into account comparison sign. Decision errors with = are more difficult to interprete
-table(xdata$p_comp)
-table(xdata$p_comp, xdata$year)
+table(xdata_de$p_comp)
+table(xdata_de$p_comp, xdata_de$year)
 
 # should focus only on comparisons that are not equal
-xdata_without_equal <- xdata |> 
+xdata_without_equal <- xdata_de |> 
   filter(p_comp != "=")
 
 # randomly sample x% of decision inconsistencies
@@ -58,10 +58,10 @@ xdata_without_equal <- xdata |>
 # evenly across years
 
 sample_prop = 0.15
-sample_size = nrow(xdata) * sample_prop
+sample_size = nrow(xdata_de) * sample_prop
 # 216
 sample_size_per_year = sample_size / length(unique(xdata$year))
-# desired sample per year 9
+# desired sample per year ~9
   
 # set random seed
 set.seed(9191)
@@ -74,7 +74,7 @@ xdata_sample <- xdata_without_equal |>
 # check if possible
 table(xdata_sample$year)
 
-# except for 2021, yes
+# except for 2004 & 2021, yes
 
 # second create subset with the desired tests per year
 xdata_sample <- xdata_sample |> 
@@ -90,14 +90,15 @@ duplicated(xdata_sample$source)
 table(xdata_sample$year)
 
 # save list
-write_csv(xdata_sample, "../data/subsample/subsample.csv")
-
-# save simpler list for manual work
-xdata_sample |> 
-  select(source, p_comp, computed_p, raw, one_tailed_in_txt) |> 
[…]

### 2025-07-20 scripts/02_sampling_for_manual_inspection.R
double checking one-sided tests

@@ -9,8 +9,8 @@
 
 # load in relevant libraries, install if not installed
 # if (!require("pacman")) install.packages("pacman")
-# pacman::p_load(rstudioapi, 
-#                tidyverse, 
+# pacman::p_load(rstudioapi,
+#                tidyverse,
 #                ggpubr
 # )
 
@@ -18,8 +18,8 @@
 
 ## load in data from location of script 
 ## (uncomment if you run script outside of sourcing it through manuscript.qmd)
-# current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
-# setwd(current_working_dir)
+#current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
+#setwd(current_working_dir)
 
 ## load in data
 xdata <- read_csv("../data/statcheck_revised_sample.csv")
@@ -93,12 +93,12 @@ table(xdata_sample$year)
 # write_csv(xdata_sample, "../data/subsample/subsample.csv")
 # 
 # # save simpler list for manual work
-# xdata_sample |> 
-#   select(source, p_comp, computed_p, raw, one_tailed_in_txt) |> 
-#   mutate(sentence = "copy paste sentence",
-#          testing_what = "specify what type of test (test on assumptions, main effect, interaction)",
-#          comment = "add comments") |> 
-#   write_csv("../data/subsample/subsample_for_manual_annotation.csv")
+xdata_sample |>
+  select(source, p_comp, computed_p, raw, one_tailed_in_txt) |>
+  mutate(sentence = "copy paste sentence",
+         reported_as = "significant/not significant/unclear",
+         comment = "add comments") |>
+  write_csv("../data/subsample/subsample_for_manual_annotation.csv")
 
 # extract files and store them into folder
 # folder from which to extract

### 2025-07-20 scripts/09_double_check_one_sided.R

@@ -1,0 +1,15 @@
+# double checking one sided t-test
+check <- xdata |> 
+  filter(decision_error == TRUE, 
+         test_type %in% c("t","F")) |> 
+  mutate(real_p = ifelse(test_type == "t", 
+                            2 * (1 - pt(q=test_value, df=df2)),
+                            pf(test_value, df1, df2)),
+         real_p_prop = real_p / computed_p)
+
+## all of them are along the [1,1] axis meaning they are not halfing the p-value for one-sided tests
+ggplot(check, aes(x = real_p,
+                  y = computed_p,
+                  color = one_tailed_in_txt)) + 
+  geom_point() +
+  facet_grid(one_tailed_in_txt ~ test_type)
